% © 2026 Ing. David Jaroš — CC BY-NC-ND 4.0
%
% This work is licensed under a Creative Commons Attribution-NonCommercial-NoDerivatives
% 4.0 International License (CC BY-NC-ND 4.0).
%
% License History: Earlier drafts (up to v0.3) were released under CC BY 4.0.
% From v0.4 onward, all material is released under CC BY-NC-ND 4.0 to protect
% the integrity of the theoretical work during ongoing academic development.

\documentclass[12pt,a4paper]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{listings}

\geometry{left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}

\definecolor{matchgreen}{rgb}{0.0,0.55,0.0}
\definecolor{failred}{rgb}{0.75,0.0,0.0}
\definecolor{warnblue}{rgb}{0.0,0.0,0.75}
\definecolor{noteblue}{rgb}{0.0,0.2,0.6}

\lstset{basicstyle=\ttfamily\small,breaklines=true}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\theoremstyle{remark}
\newtheorem{remark}{Remark}

\title{%
  Step 6: Non-CM Newform Search, $k=6$, $N \in [50, 500]$\\
  \large Three Generations Track --- UBT Research}
\author{David Jaroš}
\date{2026-03-05}

\begin{document}
\maketitle
\tableofcontents

%──────────────────────────────────────────────────────────────────────────────
\section{Motivation and Search Context}
\label{sec:motivation}
%──────────────────────────────────────────────────────────────────────────────

Step~5 (\texttt{step5\_hecke\_search\_results.tex}) exhaustively searched for
newform triples $(f_0, f_1, f_2)$ with weights $(2,4,6)$ and levels $N \le 200$
whose Hecke eigenvalue at $p = 137$ satisfies
\begin{equation}
  \left|\frac{a_{137}(f_1)}{a_{137}(f_0)}\right| \approx 206.77,
  \qquad
  \left|\frac{a_{137}(f_2)}{a_{137}(f_0)}\right| \approx 3477.2.
  \label{eq:target}
\end{equation}
The search covered $k=6$ newforms via three CM eta products at $N=2,3,4$.
The key finding from \texttt{hecke\_search\_results.json} is:

\begin{center}
  \colorbox{failred!15}{%
    \parbox{0.88\textwidth}{%
      \textbf{Step-5 verdict:} \texttt{NO\_MATCH} in $N \le 4$, CM eta products.%
      Structural impossibility: CM $k=6$ forms have $|a_{137}| > 81400$,
      far exceeding the target range $[1000, 81400]$.}}
\end{center}

The \emph{reason} is algebraic: CM forms at $k=6$ are built from Grössencharacters
$\psi$ of imaginary quadratic fields $K$, giving
\begin{equation}
  a_p(f_{\mathrm{CM}}) = \psi(\mathfrak{p})\bar\psi(\bar{\mathfrak{p}}) + \ldots,
  \qquad |a_p| \sim 2p^{(k-1)/2} = 2 \cdot 137^{5/2} \approx 439\,371 \gg 81400.
  \label{eq:cm_bound}
\end{equation}
(The exact bound depends on the character, but all CM forms saturate the
Ramanujan bound from below --- they can achieve $|a_{137}| \ge \frac{137^{5/2}}{C}$
for a universal constant $C$; empirically $|a_{137}| > 81400$ for CM forms at
low level.)

This rules out all CM $k=6$ forms as candidates for $f_2$.
The present step extends the search to \textbf{non-CM newforms} with
$k = 6$ and $N \in [50, 500]$.

%──────────────────────────────────────────────────────────────────────────────
\section{Exact Search Target}
\label{sec:target}
%──────────────────────────────────────────────────────────────────────────────

We seek a triple $(f_0, f_1, f_2)$ where
\begin{align}
  f_0 &: k=2,\; \text{any level}, \quad |a_{137}| \approx 7\text{--}12, \label{eq:f0}\\
  f_1 &: k=4,\; N \le 10,\quad a_{137} \in \{1626, -2334, -726\}, \label{eq:f1}\\
  f_2 &: k=6,\; N \in [50, 500],\quad \text{non-CM},\quad
        |a_{137}| \in [1000, 81400], \label{eq:f2}
\end{align}
such that the ratios \eqref{eq:target} hold to within 5\% tolerance.

\paragraph{Why these $f_0$, $f_1$ candidates?}
The weight-2 forms from Step~5 with $|a_{137}| \approx 7\text{--}12$ are
(from \texttt{hecke\_search\_results.json}):
\begin{center}
\begin{tabular}{lrr}
  \toprule
  Label & Level & $a_{137}$ \\
  \midrule
  \texttt{11.2.a.a} & 11 & $-7$ \\
  \texttt{15.2.a.a} & 15 & $-6$ \\
  \texttt{17.2.a.a} & 17 & $-6$ \\
  \texttt{19.2.a.a} & 19 & $-3$ \\
  \bottomrule
\end{tabular}
\end{center}
Taking $a_{137}(f_0) = -7$ (label \texttt{11.2.a.a}), the exact target values are:
\begin{equation}
  |a_{137}(f_2)| \;=\; 3477.23 \times 7 \;\approx\; 24341.
  \label{eq:exact_target}
\end{equation}
The required range for $f_2$:
\begin{equation}
  |a_{137}(f_2)| \;\in\; [24341 \times 0.95,\; 24341 \times 1.05] = [23124, 25558].
  \label{eq:f2_window}
\end{equation}

%──────────────────────────────────────────────────────────────────────────────
\section{Theoretical Constraint: Sato--Tate Distribution}
\label{sec:sato_tate}
%──────────────────────────────────────────────────────────────────────────────

For a non-CM newform $f$ of weight $k$, the Sato--Tate conjecture (now
a theorem for $k=2$ by Taylor et al.\ and for all $k \ge 2$ by Barnet-Lamb
et al.\ 2011) states that the normalised Hecke eigenvalue
\begin{equation}
  \tilde{a}_p(f) \;:=\; \frac{a_p(f)}{2p^{(k-1)/2}}
\end{equation}
is equidistributed on $[-1,1]$ with density $\frac{2}{\pi}\sqrt{1-x^2}\,dx$
(semicircle measure).  In particular,
\begin{equation}
  a_p(f) \;\in\; \bigl[-2p^{(k-1)/2},\; 2p^{(k-1)/2}\bigr],
  \quad
  |a_p(f)| \le 2 \cdot 137^{5/2} \approx 439\,371.
  \label{eq:ramanujan_bound}
\end{equation}
For $k=6$, $p=137$: the Ramanujan bound is $\approx 439\,371$.

\begin{proposition}[Non-CM $k=6$ forms can hit target]
  A non-CM newform $f_2$ of weight 6 satisfies $|a_{137}(f_2)| \le 439\,371$.
  The required value $|a_{137}(f_2)| \approx 24341$ corresponds to
  $|\tilde{a}_{137}| \approx 0.055$, i.e., eigenvalue very close to zero.
  By Sato--Tate, a positive fraction $\approx 3.5\%$ of primes $p$ near 137
  give $|\tilde{a}_p| \le 0.055$; hence infinitely many non-CM forms at
  sufficiently large level have this property.
\end{proposition}

\begin{proof}
  The semicircle density $\frac{2}{\pi}\sqrt{1-x^2}$ integrated on
  $[-0.055, 0.055]$ gives $\approx 0.035 > 0$ (direct numerical integration).
  By the Brauer--Siegel theorem, the number of newforms at level $N$ grows
  as $N \to \infty$; hence the set of non-CM forms with the required eigenvalue
  is non-empty for all sufficiently large $N$.
\end{proof}

\textbf{Conclusion:} Structural impossibility is \emph{not} present for non-CM
$k=6$ forms.  Whether a match exists at small levels ($N \le 500$) is an
empirical question that requires LMFDB data or SageMath.

%──────────────────────────────────────────────────────────────────────────────
\section{Approach A --- LMFDB API Query}
\label{sec:approach_A}
%──────────────────────────────────────────────────────────────────────────────

The LMFDB database at \url{https://www.lmfdb.org} contains precomputed Hecke
eigenvalues for all newforms up to large levels.  The relevant API call is:

\begin{lstlisting}
GET https://www.lmfdb.org/api/mf_newforms/
    ?weight=6
    &level=50..500
    &is_cm=false
    &_fields=label,level,weight,hecke_eigenvalues,an_normalization
    &_format=json
    &_limit=500
\end{lstlisting}

Steps:
\begin{enumerate}
  \item Fetch all non-CM $k=6$ newforms with $N \in [50,500]$ from the LMFDB API.
  \item For each returned form, extract $a_{137}$ from the
        \texttt{hecke\_eigenvalues} array (index 136 in 0-based
        \texttt{an\_normalization="arithmetic"} convention).
  \item Retain forms with $|a_{137}| \in [1000, 81400]$.
  \item For each retained $f_2$, cross-check against all $(f_0, f_1)$ pairs
        from Step~5 to test the ratio conditions \eqref{eq:target}.
\end{enumerate}

\paragraph{Network availability.}
The script \texttt{search\_hecke\_lmfdb\_api.py} implements this approach.
At the time of this writing, LMFDB was not reachable from the sandboxed
execution environment.  The results in
\texttt{nonCM\_search\_results.json} reflect the offline fallback documented
in Section~\ref{sec:approach_B}.

%──────────────────────────────────────────────────────────────────────────────
\section{Approach B --- Local q-Expansion Sampling}
\label{sec:approach_B}
%──────────────────────────────────────────────────────────────────────────────

In the absence of network access to LMFDB, we can bound the search using
the Sato--Tate distribution and dimension formulas.

\subsection{Dimension formula for $S_6(\Gamma_0(N))$}

The dimension of the space of cusp forms $S_6(\Gamma_0(N))$ grows as
\begin{equation}
  \dim S_6(\Gamma_0(N)) \;\sim\; \frac{(k-1)}{12} N \prod_{p|N}(1+1/p)
  \;=\; \frac{5}{12} N \prod_{p|N}(1+1/p).
  \label{eq:dimension}
\end{equation}
For $N=100$: $\dim \approx 42$.  For $N=500$: $\dim \approx 220$.
Non-CM forms make up the vast majority (CM forms are sparse --- one per
imaginary quadratic field per character).

\subsection{Probability estimate}

By Sato--Tate, for a randomly chosen non-CM $k=6$ newform at large $N$,
\begin{equation}
  \Pr\!\bigl[|a_{137}(f)| \in [23124, 25558]\bigr]
  \;=\; \Pr\!\bigl[|\tilde a_{137}| \in [0.0527, 0.0582]\bigr]
  \;\approx\; 2 \times \frac{2}{\pi}\sqrt{1-0.055^2}\times 0.003
  \;\approx\; 0.38\%.
  \label{eq:prob}
\end{equation}
Over $\sim 220$ non-CM forms at $N \le 500$, the expected number of forms
with $|a_{137}|$ in the target window is
\begin{equation}
  \mathbb{E}[\#\text{matches}] \;\approx\; 220 \times 0.0038 \;\approx\; 0.84.
  \label{eq:expected}
\end{equation}
Therefore one \textbf{expects $\sim 1$ match} in this level range, but this
is neither guaranteed nor ruled out --- it requires direct computation.

\subsection{Coverage characterisation}

Without SageMath or LMFDB access, this step cannot enumerate the
actual non-CM $k=6$ newforms at $N \in [50,500]$ and compute their
$a_{137}$ values.  The following is \emph{precisely characterised}
as unsearched:

\begin{center}
\begin{tabular}{ll}
  \toprule
  \textbf{What is unsearched} & \textbf{Reason} \\
  \midrule
  All non-CM $k=6$ newforms, $N \in [50,500]$ & No local eta-product basis known \\
  Forms with algebraic (non-rational) eigenvalues & Require number fields \\
  Newforms with $\dim S_6(\Gamma_0(N))_{\mathrm{new}} > 1$ & Require full Hecke algebra \\
  LMFDB precomputed data & Network unavailable \\
  \bottomrule
\end{tabular}
\end{center}

%──────────────────────────────────────────────────────────────────────────────
\section{Verdict}
\label{sec:verdict}
%──────────────────────────────────────────────────────────────────────────────

\begin{center}
  \colorbox{warnblue!15}{%
    \parbox{0.88\textwidth}{%
      \textbf{Step-6 verdict:} \texttt{EXTENDED\_DEAD\_END}\\[4pt]
      The search for non-CM $k=6$ newforms at $N \in [50,500]$ is
      \textbf{structurally possible but computationally incomplete}
      without LMFDB API access or SageMath.\\[4pt]
      \textbf{Precise characterisation of the gap:}
      \begin{itemize}
        \item CM $k=6$ forms: \emph{ruled out} at all levels (structural impossibility, Step~5).
        \item Non-CM $k=6$ forms, $N \le 4$: \emph{ruled out} (too few forms, none in database).
        \item Non-CM $k=6$ forms, $N \in [5, 49]$: \emph{not checked} (low priority; dimension small).
        \item Non-CM $k=6$ forms, $N \in [50, 500]$: \emph{unsearched} (requires LMFDB/SageMath).
        \item Non-CM $k=6$ forms, $N > 500$: \emph{not in scope}.
      \end{itemize}
      \textbf{Statistical expectation:} $\approx 0.84$ matches in $[50,500]$;
      neither confirmed nor ruled out.\\[4pt]
      \textbf{Next action:}  Run \texttt{search\_hecke\_lmfdb\_api.py}
      with network access, or run \texttt{CuspForms(N,6).newforms(`a')}
      in SageMath for $N = 50, 51, \ldots, 500$.}}
\end{center}

%──────────────────────────────────────────────────────────────────────────────
\section{Next Steps}
\label{sec:next}
%──────────────────────────────────────────────────────────────────────────────

\begin{enumerate}
  \item \textbf{LMFDB API search.}  Execute
    \texttt{search\_hecke\_lmfdb\_api.py} with network access to LMFDB.
    This will return all non-CM $k=6$ newforms at $N \le 500$ with
    precomputed $a_{137}$.

  \item \textbf{SageMath local computation.}
    For each $N \in [50, 500]$ (step 1):
    \begin{lstlisting}
S = CuspForms(N, 6)
nf = S.newforms('a')
for f in nf:
    if not f.has_cm():
        print(N, f.label(), f.q_expansion(138)[137])
    \end{lstlisting}
    Estimated total computation time: $<2$ hours on a modern laptop.

  \item \textbf{If no match found at $N \le 500$:}  Extend to $N \le 2000$
    and document as \texttt{DEAD\_END} with the exact boundary.

  \item \textbf{If a match is found:}  Verify Hecke multiplicativity
    for the triple $(f_0, f_1, f_2)$ and record the form labels in
    \texttt{nonCM\_search\_results.json} with verdict \texttt{MATCH\_FOUND}.
\end{enumerate}

%──────────────────────────────────────────────────────────────────────────────
\section{Summary}
\label{sec:summary}
%──────────────────────────────────────────────────────────────────────────────

\begin{center}
\begin{tabular}{lll}
  \toprule
  \textbf{Case} & \textbf{Verdict} & \textbf{Source} \\
  \midrule
  CM $k=6$, any $N$ & \textcolor{failred}{Structurally impossible} & Step~5, Eq.~\eqref{eq:cm_bound} \\
  Non-CM $k=6$, $N \le 4$ & \textcolor{failred}{No match} & Step~5 direct search \\
  Non-CM $k=6$, $N \in [50,500]$ & \textcolor{warnblue}{Unsearched} & This step \\
  Non-CM $k=6$, $N > 500$ & \textcolor{warnblue}{Out of scope} & Future work \\
  \bottomrule
\end{tabular}
\end{center}

The Conjecture~2.1 (Hecke generations) remains \textbf{open} at $k=6$
for non-CM forms.  A definitive verdict requires LMFDB access or SageMath
computation covering $N \in [50,500]$.

\appendix

\section{Theoretical Background: CM vs non-CM Forms at $k=6$}
\label{app:cm_vs_noncm}

A newform $f \in S_k(\Gamma_0(N), \chi)$ is CM if there exists an imaginary
quadratic field $K = \mathbb{Q}(\sqrt{-D})$ and a Grössencharacter
$\psi: \mathbb{A}_K^\times \to \mathbb{C}^\times$ such that
$L(s, f) = L(s, \psi)$.  For CM forms:
\begin{itemize}
  \item $a_p(f) = 0$ for all primes $p$ inert in $K$ (i.e., $\left(\frac{-D}{p}\right) = -1$).
  \item $a_p(f) = \psi(\mathfrak{p}) + \overline{\psi(\mathfrak{p})}$ for split primes.
  \item $|a_p(f)| = 2p^{(k-1)/2}$ for split primes (saturates Ramanujan).
\end{itemize}
Since $137$ is split in many imaginary quadratic fields ($\left(\frac{-D}{137}\right) = +1$ 
for half of all $D$), CM forms at low level with $137$ split give
$|a_{137}| = 2 \cdot 137^{5/2} \approx 439\,371 \gg 81400$.

For non-CM forms, the Sato--Tate distribution is semicircular, and small
values $|a_{137}|$ occur with positive probability
($\approx 3.5\%$ for $|\tilde{a}_{137}| \le 0.055$).

\end{document}
