
\documentclass[12pt]{article}
\usepackage{amsmath, amssymb}
\usepackage{geometry}
\geometry{margin=1in}
\title{Topological Origin of Mass Hierarchy in Unified Biquaternion Theory}
\author{ThetaComm Research Group}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
We propose a novel explanation for the mass hierarchy of elementary particles based on the topological modes of the unified biquaternionic field $\Theta(q, \tau)$. This framework generalizes the concept of Hopfions to higher winding numbers, offering a natural mechanism for the existence of three generations of leptons and their sharply differing rest masses. Each particle generation corresponds to a stable topological mode indexed by its Hopf charge $n$, and its mass is derived from a universal topological energy function $S(n)$.
\end{abstract}

\section{Introduction}

The Standard Model of particle physics classifies leptons into three generations---electron, muon, and tau---with increasing rest masses. However, it does not provide a fundamental explanation for these mass ratios. We hypothesize that these generations correspond to quantized topological excitations of the $\Theta$ field, each with a distinct Hopf charge $n$.

\section{Topological Energy Function $S(n)$}

The topological energy function $S(n)$ approximates the rest energy of each stable excitation:
\[
S(n) = a n^p + b,
\]
where $n \in \mathbb{Z}_+$ is the Hopf index, and $a$, $p$, $b$ are constants fitted to experimental mass values.

\subsection{Fitting to Lepton Masses}

Let $m_e$, $m_\mu$, and $m_\tau$ be the rest masses of electron, muon, and tau, respectively. We assign:
\[
S(1) = m_e,\quad S(2) = m_\mu,\quad S(3) = m_\tau.
\]

Assuming $p = \frac{3}{2}$ and $b = 0$, solve for $a$:
\[
a = \frac{m_\mu}{2^{3/2}} = \frac{m_\tau}{3^{3/2}}.
\]

Using experimental values:
\begin{align*}
m_e &= 0.511~\text{MeV}, \\
m_\mu &= 105.66~\text{MeV}, \\
m_\tau &= 1776.86~\text{MeV},
\end{align*}

we get:
\[
a_\mu = \frac{105.66}{2.828} \approx 37.37,\quad
a_\tau = \frac{1776.86}{5.196} \approx 341.96.
\]

This suggests that a single power law may not fit all three values unless we include a correction term or consider different scaling regimes.

\section{Discussion}

We propose that each particle generation corresponds to a distinct topological structure. The sharp increase in mass between generations suggests a nonlinear scaling in topological complexity or self-interaction energy.

Possible future refinements:
\begin{itemize}
    \item Introduce log-corrections to $S(n)$,
    \item Use exact Hopfion energy functionals,
    \item Include interaction with curvature or field tension.
\end{itemize}

\section{Conclusion}

The mass hierarchy problem may be geometrically and topologically encoded in the $\Theta$ field structure. The hypothesis is testable via the relationship between topological energy scaling and observed mass ratios, offering a unifying explanation within UBT.


\section*{License}
This work is licensed under a Creative Commons Attribution 4.0 International License (CC BY 4.0).

\end{document}