% © 2025 Ing. David Jaroš — CC BY-NC-ND 4.0
%
% This work is licensed under a Creative Commons Attribution-NonCommercial-NoDerivatives 
% 4.0 International License (CC BY-NC-ND 4.0).
%
% License History: Earlier drafts (up to v0.3) were released under CC BY 4.0. 
% From v0.4 onward, all material is released under CC BY-NC-ND 4.0 to protect 
% the integrity of the theoretical work during ongoing academic development.
%
% See LICENSE.md for full license text.

\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{booktabs}
\geometry{a4paper, margin=1in}

% Define keywords command
\providecommand{\keywords}[1]{\textbf{Keywords:} #1}

\title{Unified Biquaternion Theory (UBT): Complex Time, Consciousness, and Field Unification}
\author{Ing.~David~Jaroš}
\date{November 2025}

\begin{document}
\maketitle

% License Notice - Visible in PDF
\noindent
\textbf{License:} © 2025 Ing. David Jaroš. This work is licensed under a Creative Commons Attribution-NonCommercial-NoDerivatives 4.0 International License (CC BY-NC-ND 4.0). Earlier drafts (up to v0.3) were released under CC BY 4.0. From v0.4 onward, all material is released under CC BY-NC-ND 4.0 to protect the integrity of the theoretical work during ongoing academic development. See \url{https://creativecommons.org/licenses/by-nc-nd/4.0/} for details.

\vspace{1em}

\begin{abstract}
The Unified Biquaternion Theory (UBT) presents a unified framework combining General Relativity, Quantum Field Theory, and Standard Model symmetries within a biquaternionic field defined over complex time $\tau = t + i\psi$. This theory generalizes Einstein's General Relativity and, in the real-valued limit, exactly reproduces Einstein's field equations for all curvature regimes. UBT extends the framework through additional biquaternionic degrees of freedom that may correspond to dark sector physics and quantum gravitational corrections.
\end{abstract}

\noindent\textbf{Biquaternionic Geometry Lock-in Statement:}

Throughout this work, all geometric and dynamical structures are defined at the biquaternionic level ($\mathbb{B} = \mathbb{H} \otimes \mathbb{C}$):

\begin{itemize}
\item \textbf{Metric:} Fundamental object is $\mathcal{G}_{\mu\nu}(x) \in \mathbb{B}$, derived from tetrads $E_\mu$ via $\mathcal{G}_{\mu\nu} = \text{Sc}(E_\mu E_\nu^\dagger)$. The classical metric $g_{\mu\nu} := \text{Re}(\mathcal{G}_{\mu\nu})$ is a derived projection only.

\item \textbf{Connection:} Fundamental connection is $\Omega_\mu(x) \in \mathbb{B}$. Christoffel symbols $\Gamma^\lambda_{\mu\nu} = \text{Re}(\Omega^\lambda_{\mu\nu})$ are derived, not fundamental.

\item \textbf{Curvature:} Fundamental curvature is $\mathcal{R}_{\mu\nu} = \partial_\mu \Omega_\nu - \partial_\nu \Omega_\mu + [\Omega_\mu, \Omega_\nu]$. Ricci tensor: $R_{\mu\nu} := \text{Re}(\mathcal{R}_{\mu\nu})$.

\item \textbf{Stress-energy:} Fundamental tensor is $\mathcal{T}_{\mu\nu} = \langle D_\mu\Theta, D_\nu\Theta \rangle_\mathbb{B} - \frac{1}{2}\mathcal{G}_{\mu\nu}\langle D\Theta, D\Theta \rangle$, where $T_{\mu\nu} := \text{Re}(\mathcal{T}_{\mu\nu})$ is a geometric phase response, not an external matter source.

\item \textbf{Field equations:} The fundamental equation is $\mathcal{E}_{\mu\nu} = \kappa \mathcal{T}_{\mu\nu}$ (biquaternionic Einstein tensor). Classical Einstein equations $G_{\mu\nu} = 8\pi G T_{\mu\nu}$ arise only after taking $\text{Re}(\cdot)$.
\end{itemize}

Any real-valued spacetime metric, curvature, or stress--energy tensor represents a Hermitian projection corresponding to an observer-restricted sector. No physical conclusion should be interpreted at the level of the real projection alone. General Relativity is the real projection of fundamental biquaternionic geometry.

\keywords{biquaternion, complex time, consciousness, unified field theory, Hermitian gravity, SU(3) symmetry, theta function}

\tableofcontents

\section{Introduction}

The Unified Biquaternion Theory (UBT) aims to provide a comprehensive framework for understanding fundamental physics by embedding General Relativity, Quantum Field Theory, and the Standard Model within a single mathematical structure based on biquaternionic fields over complex time.

The core equation of UBT is:
\begin{equation}
\nabla^\dagger\nabla\Theta(q,\tau) = \kappa \mathcal{T}(q,\tau)
\end{equation}
where $\Theta(q,\tau)$ is the unified biquaternionic field, $q \in \mathbb{C}\otimes\mathbb{H}$ represents biquaternion coordinates, $\tau = t + i\psi$ is complex time, $\nabla^\dagger\nabla$ is the gauge-covariant d'Alembertian, $\kappa$ is the coupling constant related to $8\pi G$, and $\mathcal{T}$ is the biquaternionic stress--energy encoding phase inconsistency of $\Theta$. The Hermitian projection $T_{\mu\nu}:=\Re(\mathcal{T}_{\mu\nu})$ is interpreted only as the observable real-sector response, not as a fundamental ``matter source.''

\section{Mathematical Foundations}

\subsection{Biquaternionic Structure}

The biquaternion algebra $\mathbb{B} = \mathbb{C}\otimes\mathbb{H}$ combines complex numbers with quaternions, providing a rich algebraic structure capable of encoding both spacetime geometry and internal gauge symmetries.

\subsection{Complex Time}

Complex time $\tau = t + i\psi$ extends the real time coordinate $t$ by adding an imaginary component $\psi$ that represents phase-like degrees of freedom. This extension enables the unified treatment of quantum phases and gravitational curvature within a single geometric framework.

\subsection{Biquaternionic Metric and Projection}

The fundamental geometry is encoded in the biquaternionic metric $\mathcal{G}_{\mu\nu}(\Theta)$ built from the field $\Theta$. All curvature tensors $\mathcal{R}_{\mu\nu}$ and connections are derived from $\mathcal{G}_{\mu\nu}$. The classical metric is a projection
\[
g_{\mu\nu} \;:=\; \Re\!\left(\mathcal{G}_{\mu\nu}\right),
\qquad
R_{\mu\nu} \;:=\; \Re\!\left(\mathcal{R}_{\mu\nu}\right),
\qquad
T_{\mu\nu} \;:=\; \Re\!\left(\mathcal{T}_{\mu\nu}\right),
\]
used only to describe the Hermitian (real) sector. The connection $\Omega_\mu$ and curvature are defined at the biquaternionic level; any Levi--Civita symbol or Christoffel term is understood solely as the Hermitian projection $\Re(\Omega_\mu)$.

\section{Physical Content}

\subsection{General Relativity Recovery}

In the Hermitian (real) projection $\mathcal{G}_{\mu\nu}\mapsto g_{\mu\nu}$ and $\mathcal{R}_{\mu\nu}\mapsto R_{\mu\nu}$, UBT exactly reproduces Einstein's field equations:
\begin{equation}
R_{\mu\nu} - \frac{1}{2}g_{\mu\nu}R = 8\pi G T_{\mu\nu}
\end{equation}
This ensures full compatibility with all experimental confirmations of General Relativity, including perihelion precession, gravitational waves, and cosmological solutions.

\subsection{Why GR Misses Exotic Regimes}

Classical GR observes only $g_{\mu\nu}=\Re(\mathcal{G}_{\mu\nu})$, so any solution with $\Im(\mathcal{G}_{\mu\nu})\neq 0$ projects to the same real geometry and remains invisible. Antigravitational responses, metric cloaking, or temporal drift arise from these imaginary components and are suppressed in the projection. They are not forbidden by physics; they are hidden by the Hermitian reduction. UBT therefore extends GR without contradiction: the real sector matches Einsteinian dynamics, while the full biquaternionic geometry supports exotic but observationally hidden regimes.

\subsection{Standard Model Symmetries}

The gauge group $SU(3) \times SU(2) \times U(1)$ of the Standard Model emerges naturally from the automorphism group of the biquaternionic structure, providing a geometric origin for the strong, weak, and electromagnetic interactions.

\subsection{Dark Sector Physics}

The additional degrees of freedom in the biquaternionic field, beyond those required for reproducing General Relativity, may provide candidates for dark matter and dark energy phenomena through p-adic extensions of the theory.

\section{Future-Proofing Rule for Extensions}

Any future extension of UBT must (i) define new dynamics at the biquaternionic level, (ii) state explicitly how the GR sector is obtained as $\Re(\,\cdot\,)$, and (iii) avoid introducing classical GR objects (metrics, Levi--Civita symbols, or stress--energy tensors) as axioms without such a projection. This applies to all new appendices, phenomenological discussions, and experimental proposals.

\section{Acknowledgments}

The author acknowledges A.H. Chamseddine (2025) and E. Verlinde (2025).  
The present biquaternionic formulation extends these approaches through a unified complex-time and toroidal topology model.

\appendix

% Include appendices
\input{Appendix_F_Hermitian_Limit}
\input{Appendix_G_Emergent_SU3}
\input{Appendix_H_Theta_Phase_Emergence}
\input{appendix_QB_quantum_bridge}

\bibliographystyle{plain}
\bibliography{references}

\section*{License}
© 2025 Ing. David Jaroš — CC BY-NC-ND 4.0

This work is licensed under a Creative Commons Attribution-NonCommercial-NoDerivatives 4.0 International License (CC BY-NC-ND 4.0).

\textbf{License History:} Earlier drafts (up to v0.3) were released under CC BY 4.0. From v0.4 onward, all material is released under CC BY-NC-ND 4.0 to protect the integrity of the theoretical work during ongoing academic development.

\end{document}
