\section*{Appendix 14: Experimental Ethics, Calibration, and Safety Guidelines}

\input{../THEORY_STATUS_DISCLAIMER}
\ConsciousnessDisclaimer

As the Unified Biquaternion Theory (UBT) introduces hypotheses that potentially link physical fields to consciousness through the internal phase component \(\psi\), and as future experimental designs such as the Psi Resonator aim to detect or modulate this component, it becomes crucial to address the ethical, safety, and calibration framework necessary for any related empirical research.

\subsection*{14.1 Ethical Principles for Consciousness-Related Experiments}

\begin{itemize}
  \item \textbf{Informed Consent:} Any experiment involving human subjects must ensure clear, documented, and reversible consent, including full disclosure of aims, risks, and possible mental or emotional effects.
  \item \textbf{Non-invasive Priority:} All protocols must prioritize non-invasive, observational or biofeedback-only methods unless justified by rigorous scientific rationale and approved by an independent ethics board.
  \item \textbf{Privacy and Subjective State Respect:} The subjective content, states, or experiences of participants must not be recorded, disclosed, or interpreted without their consent and complete anonymity.
  \item \textbf{Responsibility:} Experiments must be monitored by trained professionals with experience in both biofeedback devices and psychological effects of sensory modulation.
\end{itemize}

\subsection*{14.2 Calibration of Phase-sensitive Devices}

\begin{itemize}
  \item \textbf{Null Tests:} Devices such as the Psi Resonator should be calibrated using non-living control systems to detect spurious resonances.
  \item \textbf{Biological Baseline:} Healthy, relaxed individuals should be used to generate baseline biofield data under known and repeatable conditions.
  \item \textbf{Environmental Shielding:} Electromagnetic shielding and thermal isolation should be applied to reduce background interference.
\end{itemize}

\subsection*{14.3 Safety Protocols}

\begin{itemize}
  \item \textbf{Power Limits:} No electromagnetic or sonic output should exceed safety standards for consumer-grade biofeedback or MRI devices.
  \item \textbf{Continuous Monitoring:} Real-time monitoring of EEG, heart rate, and skin response is advised during any live test.
  \item \textbf{Emergency Shutdown:} All experimental setups must include a manual and automated shutdown mechanism.
\end{itemize}

\subsection*{14.4 Long-term Considerations}

This theory implies the possibility of interacting with structures related to subjective awareness. Even though current experiments are not expected to create cognitive effects, the precautionary principle applies. Future work must proceed with interdisciplinary collaboration—bridging physics, neuroscience, and ethics.

