\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,geometry}
\geometry{margin=1in}

\title{Damping-Compensated Tensor Lifting and Approximate Diagonalization via 2D FFT}
\author{Unified Biquaternion Theory Notes}
\date{April 10, 2026}

\newcommand{\C}{\mathbb{C}}
\newcommand{\R}{\mathbb{R}}

\begin{document}
\maketitle

\begin{abstract}
We present a computational framework that lifts a one-dimensional complex signal into a two-index interaction tensor, applies radial damping compensation, and maps the result through a two-dimensional discrete Fourier transform. The goal is structured spectral concentration rather than exact diagonalization. We provide explicit definitions of signal, lifting, weighting, and transform, and we introduce a diagonal-dominance metric for quantitative testing. We separate definitions, assumptions, and hypothesis-level claims. The framework is compatible with UBT-style quadratic-index structures and is intended as a rigorous, testable construction.
\end{abstract}

\section*{License}
This work is licensed under the Creative Commons Attribution 4.0 International License (CC BY 4.0).
You are free to share and adapt the material, provided appropriate credit is given.

\section{Introduction}
A recurring problem in damped complex systems is that natural representations are rich in interaction terms but difficult to analyze in an orthogonal spectral basis. Standard one-dimensional Fourier analysis gives exact diagonalization for circulant operators, yet damped two-index interaction objects generally retain nontrivial off-diagonal coupling.

The present note introduces a bridge:
\begin{enumerate}
    \item tensor lifting from a mode vector to a two-index interaction tensor,
    \item radial compensation of damping by an exponential weight,
    \item two-dimensional FFT as an efficient spectral map.
\end{enumerate}
The resulting representation is computationally efficient and structurally aligned with heat-kernel-like damping and theta-like quadratic phase organization. We emphasize from the outset: approximate diagonalization is a hypothesis to test, not a universal theorem.

\paragraph{Motivation for 2D lifting.}
The 2D lifting exposes pairwise mode interactions that are not accessible in a purely 1D spectral representation. This interaction-space view is the main reason the framework can isolate coherence structure beyond standard mode-energy analysis.

\paragraph{Heat-kernel interpretation.}
The construction can be interpreted as a discrete heat-kernel transform in interaction space: $\phi$ controls diffusion scale before the spectral map, while $\psi$ controls post-transform spectral regularization.
The transform maps quadratic interactions into approximately separable spectral components.

\section{Definitions, Assumptions, and Core Objects}
We fix an odd mode count $N=2K+1$ with indices $n,m\in\{-K,\dots,K\}$.

\subsection*{Definition 1 (Signal)}
Let
\begin{equation}
a=(a_{-K},\dots,a_K)\in\C^N
\label{eq:signal_def}
\end{equation}
be a complex mode vector.
Equation \eqref{eq:signal_def} defines the primary mode-domain object used in all later constructions.

\subsection*{Definition 2 (Tensor Lifting)}
Define the lifted interaction tensor
\begin{equation}
\rho_{nm}:=a_n\overline{a_m},\qquad n,m\in\{-K,\dots,K\}.
\label{eq:rho_def}
\end{equation}
Equivalently, $\rho=aa^{\ast}$.

\subsection*{Definition 3 (Radial Damping Compensation)}
For a real parameter $\phi\in\R$, define
\begin{equation}
W_{nm}(\phi):=\exp\!\big(\pi\phi(n^2+m^2)\big),
\label{eq:weight_def}
\end{equation}
and weighted tensor
\begin{equation}
\rho^{(w)}:=W\odot\rho,
\label{eq:rho_weighted}
\end{equation}
where $\odot$ is elementwise multiplication.

\subsection*{Definition 4 (2D Transform)}
Let $\mathcal{F}_2$ denote the 2D DFT on $N\times N$ arrays. Define
\begin{equation}
\widetilde{\rho}:=\mathcal{F}_{2,\mathrm{basis\mbox{-}weighted}}(\rho).
\label{eq:fft2_def}
\end{equation}
In index form,
\begin{equation}
\widetilde{\rho}_{pq}=\sum_{n=-K}^{K}\sum_{m=-K}^{K}\rho_{nm}
\exp\!\left(-2\pi i\frac{pn+qm}{N}\right)
\exp\!\big(-\pi\phi(n^2+m^2)\big),\qquad p,q\in\{-K,\dots,K\}.
\label{eq:fft2_index}
\end{equation}
Equation \eqref{eq:fft2_index} is the explicit coordinate-level form of the transform in \eqref{eq:fft2_def}.

The exponential factor $\exp(-\pi\phi(n^2+m^2))$ is applied to the transform kernel (basis functions), not directly to the data tensor. For computational efficiency, this is implemented as a pointwise multiplication prior to FFT, which is mathematically equivalent.
The equivalence between basis weighting and pre-FFT multiplication follows from linearity of the transform and should be interpreted as an implementation shortcut rather than a literal modification of data.

\subsection*{Assumption A (Envelope-Compensation Match)}
Observed attenuation is approximately compatible with a quadratic radial envelope in $(n,m)$, so the weight in \eqref{eq:weight_def} partially compensates damping before spectral projection.

\section{Tensor Lifting}
Using \eqref{eq:rho_def}, each entry $\rho_{nm}$ encodes pair interaction between modes $n$ and $m$.

\subsection*{Proposition 1}
The lifted tensor $\rho$ is Hermitian positive semidefinite, and in the single-vector noiseless construction has rank one.

\subsection*{Proof sketch}
Hermitian symmetry follows from $\rho_{mn}=a_m\overline{a_n}=\overline{\rho_{nm}}$. For any $x\in\C^N$, $x^{\ast}\rho x=x^{\ast}aa^{\ast}x=|a^{\ast}x|^2\ge 0$. Rank-one follows from the outer-product form $aa^{\ast}$.

\paragraph{Interpretation.}
Diagonal entries $\rho_{nn}=|a_n|^2$ encode mode energy, while off-diagonal entries encode pairwise phase coherence/interference in interaction space.

\section{Damping and Radial Weighting}
The weighting defined in \eqref{eq:weight_def} and applied in \eqref{eq:rho_weighted} is radially symmetric in $r^2=n^2+m^2$.

\paragraph{Proven property.}
$W$ is separable and radially symmetric in index-space radius $r^2=n^2+m^2$.

\paragraph{Role.}
If observed coefficients are damped by factors approximately proportional to $\exp(-\pi\phi(n^2+m^2))$, then \eqref{eq:weight_def} compensates this envelope before spectral analysis.
The weighting $W$ corresponds to an inverse heat kernel, compensating pre-existing damping in the signal. While the transform kernel contains $\exp(-\pi\phi(n^2+m^2))$, the implementation uses $\exp(+\pi\phi(n^2+m^2))$ to rebalance attenuated high-index components before projection.

\section{2D Fourier Transform}
Equation \eqref{eq:fft2_def} defines the transform map from compensated interaction space to a spectral lattice.

\paragraph{Proven property.}
$\mathcal{F}_2$ is invertible with norm-conserving properties up to normalization convention; Parseval identities therefore hold modulo that convention.

\paragraph{Expected effect.}
When compensation aligns with damping structure, spectral mass in $\widetilde{\rho}$ may concentrate toward diagonal coordinates in a chosen ordering, reducing effective cross-coupling.

\section{Forward and Inverse Transform}
The framework defines an invertible transform pair:
\begin{equation}
\widetilde{\rho} = \mathcal{F}_{2,\mathrm{basis\mbox{-}weighted}}(\rho),
\label{eq:forward_transform}
\end{equation}
where $W$ is elementwise invertible (all entries nonzero). The inverse transform recovers $\rho$ from $\widetilde{\rho}$:
\begin{equation}
\rho^{(w)} = \mathcal{F}_2^{-1}(\widetilde{\rho}),
\label{eq:inverse_fft}
\end{equation}
\begin{equation}
\rho = W^{-1} \odot \rho^{(w)},
\label{eq:inverse_weight}
\end{equation}
where $W^{-1}_{nm} := \exp(-\pi\phi(n^2+m^2))$ is the inverse weighting operator. The composition $\mathcal{T} = \mathcal{F}_2 \circ W$ is invertible:
\begin{equation}
\mathcal{T}^{-1} = W^{-1} \circ \mathcal{F}_2^{-1}.
\label{eq:inverse_operator}
\end{equation}
Equations \eqref{eq:forward_transform}--\eqref{eq:inverse_operator} define the forward path (lifting $\to$ weight $\to$ FFT) and inverse path (inverse FFT $\to$ inverse weight).

After the transform, a spectral damping matrix is applied:
\begin{equation}
D_{k\ell}:=\exp\!\big(-\pi\psi(k^2+\ell^2)\big),
\label{eq:spectral_damping}
\end{equation}
with final spectrum
\begin{equation}
\widetilde{\rho}^{(d)} := D \odot \widetilde{\rho}.
\label{eq:damped_spectrum}
\end{equation}
This step suppresses high-frequency spectral modes.
The transform is strictly invertible only when $\psi=0$. For $\psi>0$, spectral damping introduces regularization and the map becomes information-reducing.

\section{Role of $\phi$ and $\psi$}
The two damping parameters play distinct roles:
\begin{itemize}
    \item $\phi$ controls pre-transform diffusion in interaction space (input geometry smoothing),
    \item $\psi$ controls post-transform damping in spectral space (spectral regularization).
\end{itemize}
This separation enables independent control of interaction geometry and spectral stability.

\section{Properties}
\paragraph{Rotational symmetry in interaction space.}
The radial factor $\exp(-\pi\phi(n^2+m^2))$ depends only on $n^2+m^2$, so the weighting geometry is approximately invariant under discrete rotations preserving $n^2+m^2$. This gives isotropic damping structure in interaction coordinates.

\section{Limiting Cases}
The framework has transparent limiting regimes:
\begin{itemize}
    \item $\phi\to 0$: the basis-weighted transform reduces to standard 2D FFT,
    \item $\phi\to \infty$: only low-index interactions contribute appreciably,
    \item $\psi\to 0$: no post-transform spectral damping,
    \item $\psi\to \infty$: only lowest-frequency spectral modes remain.
\end{itemize}

\section{Integration over $\phi$}
A multiscale extension is defined by integrating over diffusion scale:
\begin{equation}
\widetilde{\rho}^{(\mathrm{int})}_{k\ell} = \int \mu(\phi) \sum_{n,m} \rho_{nm}
\exp\!\left(-2\pi i\frac{kn+\ell m}{N}\right)
\exp\!\big(-\pi\phi(n^2+m^2)\big)\, d\phi.
\label{eq:phi_integrated_transform}
\end{equation}
A practical discrete approximation is
\begin{equation}
\widetilde{\rho}^{(\mathrm{int})} \approx \sum_j w_j\,\widetilde{\rho}^{(\phi_j)}.
\label{eq:phi_integrated_discrete}
\end{equation}
This induces the multiscale radial kernel
\begin{equation}
K(n,m):=\int \mu(\phi)\exp\!\big(-\pi\phi(n^2+m^2)\big)\,d\phi,
\label{eq:multiscale_kernel}
\end{equation}
which generalizes a single-scale Gaussian to a scale-mixture kernel.
This effectively replaces a single Gaussian kernel with a mixture of Gaussians, leading to heavier tails or sharper cutoffs depending on $\mu(\phi)$.
Therefore, $\phi$-integration is not equivalent to a constant scaling, but produces a structurally different radial filter.

\subsection*{Special cases of $\phi$-integration}
\begin{itemize}
    \item $\mu(\phi)=\delta(\phi-\phi_0)$ recovers the fixed-$\phi$ transform,
    \item broad $\mu(\phi)$ yields multiscale smoothing,
    \item normalized $\mu(\phi)$ yields a scale-averaged transform.
\end{itemize}

\subsection*{Remark: Exact vs Practical Transform}
The forward transform defined in \eqref{eq:forward_transform},
\[
\widetilde{\rho} = \mathcal{F}_{2,\mathrm{basis\mbox{-}weighted}}(\rho),
\]
does not, by itself, constitute a fully orthogonal theta transform in the exact mathematical sense. A mathematically rigorous formulation would require compensating for the weighting operator $W$ via inverse weighting (deconvolution), performed directly in the spectral domain before FFT or after. In the present work, we intentionally defer the inverse weighting to the reconstruction step \eqref{eq:inverse_weight}, treating it as part of the inverse transform pipeline.

This design choice has practical advantages:
\begin{itemize}
    \item spectral deconvolution (performing $W^{-1}$ in spectral space before FFT) is numerically ill-conditioned for high-frequency modes,
    \item deferring inverse weighting to the spatial-domain inverse avoids this instability,
    \item the invertible structure is preserved: $\mathcal{T}^{-1} = W^{-1} \circ \mathcal{F}_2^{-1}$.
\end{itemize}

As a result, the presented method should be understood as:
\begin{quote}
    \emph{A theta-inspired weighted spectral transform that is invertible and approximately orthogonal under a weighted metric, rather than a strict discrete theta transform in the functional-analytic sense.}
\end{quote}
This distinction is important for reviewers and future users: the method trades exact orthogonality for numerical robustness and efficiency.

\section{Weighted Metric and Energy Interpretation}
The transform is not unitary in the standard Euclidean sense on $\rho$, since $W$ introduces exponential scaling. However, a weighted inner product can be defined:
\begin{equation}
\langle \rho_1, \rho_2 \rangle_W := \sum_{n,m} W_{nm}^{-1} \rho_{1,nm} \overline{\rho_{2,nm}},
\label{eq:weighted_inner}
\end{equation}
giving the norm
\begin{equation}
\|\rho\|_W^2 := \sum_{n,m} W_{nm}^{-1} |\rho_{nm}|^2.
\label{eq:weighted_norm}
\end{equation}
Under this metric, the forward and inverse transforms approximately preserve energy:
\begin{itemize}
    \item The forward transform concentrates energy spectral distribution.
    \item The inverse transform recovers magnitude structure (up to regularization and numerical precision).
\end{itemize}

\paragraph{Interpretation.}
The weighted metric reflects the geometry induced by damping compensation. Entries with large $W$ (high-index pairs) contribute less to the weighted norm; entries with small $W$ (low-index pairs) dominate. This weighting ensures that the inversion step does not amplify high-index noise.

\section{Diagonal Metric and Approximate Diagonalization Hypothesis}
Define diagonal and off-diagonal magnitudes directly on $\widetilde{\rho}$:
\begin{equation}
E_{\mathrm{diag}} := \sum_{j=1}^{N} |\widetilde{\rho}_{jj}|,
\qquad
E_{\mathrm{off}} := \sum_{i\neq j}|\widetilde{\rho}_{ij}|.
\label{eq:energies}
\end{equation}
Equation \eqref{eq:energies} supplies the two scalar aggregates used by the ratio in \eqref{eq:R_def}.
Define the diagonal-dominance ratio
\begin{equation}
R := \frac{\sum_{i\neq j}|\widetilde{\rho}_{ij}|}{\sum_{j=1}^{N}|\widetilde{\rho}_{jj}|+\varepsilon}, \qquad \varepsilon>0\text{ small}.
\label{eq:R_def}
\end{equation}
Equation \eqref{eq:R_def} is the operational metric used throughout this note.

A normalized variant can also be used:
\[
R' = \frac{E_{\mathrm{off}}}{E_{\mathrm{off}} + E_{\mathrm{diag}}},
\]
which ensures boundedness and scale stability.

Interpretation of \eqref{eq:R_def}: $R\ll 1$ indicates near-diagonal structure, while $R\approx 1$ indicates no strong diagonal organization.

\subsection*{Hypothesis H (Approximate Diagonalization)}
For classes of damped structured signals and suitable $\phi$, the transformed tensor $\widetilde{\rho}$ obtained via \eqref{eq:fft2_def} exhibits reduced off-diagonal mass, yielding smaller $R$ than uncompensated baselines.

\paragraph{Status.}
H is a testable hypothesis. It is not claimed as a universal theorem over all signals and all $\phi$.

\section{Algorithmic Pipeline}
The computational pipeline is:
\begin{enumerate}
    \item Build $\rho$ from $a$ using \eqref{eq:rho_def}.
    \item Apply basis weighting (implemented via pre-FFT multiplication).
    \item Compute $\widetilde{\rho}$ using \eqref{eq:fft2_def}.
    \item Optionally compute multiscale aggregation using \eqref{eq:phi_integrated_discrete}.
    \item Apply spectral damping using \eqref{eq:damped_spectrum}.
    \item Compute $R$ using \eqref{eq:R_def} on the damped spectrum.
\end{enumerate}

\noindent Pseudocode:
\begin{verbatim}
Input: complex vector a[n], parameter phi
rho[n,m] = a[n] * conjugate(a[m])
W[n,m] = exp(pi * phi * (n*n + m*m))
rho_w = W .* rho   # computational implementation of basis weighting
rho_tilde = FFT2(rho_w)

# optional phi integration
rho_tilde_int = 0
for (phi_j, w_j) in phi_grid:
    W_j[n,m] = exp(pi * phi_j * (n*n + m*m))
    rho_tilde_j = FFT2(W_j .* rho)
    rho_tilde_int = rho_tilde_int + w_j * rho_tilde_j

if use_phi_integration:
    rho_tilde = rho_tilde_int

D[k,l] = exp(-pi * psi * (k*k + l*l))
rho_tilde_d = D .* rho_tilde

diag = sum_n |rho_tilde_d[n,n]|
off  = sum_{n!=m} |rho_tilde_d[n,m]|
R = off / (diag + eps)
\end{verbatim}

\section{Numerical Stability and Regularized Inverse}
The inverse weighting step \eqref{eq:inverse_weight} uses $W^{-1}_{nm} = \exp(-\pi\phi(n^2+m^2))$. For large indices and large $\phi$, this exponential decay becomes very small, risking numerical underflow or amplification of rounding errors when rescaling.

\paragraph{Regularization strategy.}
To stabilize the inverse, we employ a regularized inverse weight:
\begin{equation}
W^{(-1)}_{nm}(\lambda) := \frac{\exp(-\pi\phi(n^2+m^2))}{1 + \lambda \exp(-\pi\phi(n^2+m^2))},
\label{eq:regularized_inverse}
\end{equation}
where $\lambda > 0$ is a regularization parameter. This formula achieves several goals:
\begin{itemize}
    \item it keeps $W^{(-1)} \le 1$ for all $(n,m)$,
    \item it suppresses amplification at high indices,
    \item as $\lambda \to 0$, it recovers the exact inverse $W^{-1}$.
\end{itemize}

\paragraph{Alternative approaches.}
Other stabilization methods include:
\begin{itemize}
    \item truncation of high-index modes before inversion,
    \item Wiener-style filtering of $\widetilde{\rho}$ before inverse FFT,
    \item bounded inverse weighting to cap amplification at a fixed threshold.
\end{itemize}

\section{Operator Interpretation}
The full transformation can be written compactly as:
\begin{equation}
\widetilde{\rho} = \mathcal{T}(\rho), \quad \text{where} \quad \mathcal{T} := \mathcal{F}_{2,\mathrm{basis\mbox{-}weighted}}.
\label{eq:operator}
\end{equation}
This separates the action into two independent steps:
\begin{itemize}
    \item basis weighting: geometric transformation (damping compensation, radial scaling),
    \item $\mathcal{F}_2$: spectral orthogonalization (2D FFT basis change),
    \item $D$: post-transform spectral regularization.
\end{itemize}
The inverse operator is:
\begin{equation}
\mathcal{T}^{-1} := W^{(-1)} \circ \mathcal{F}_2^{-1},
\label{eq:inverse_decomposition}
\end{equation}
recovering the original tensor (up to regularization error).

\section{Interpretation and Discussion}
In this framework:
\begin{itemize}
    \item diagonal entries in $\widetilde{\rho}$ are interpreted as dominant spectral energy channels,
    \item off-diagonal entries represent residual interference and phase coupling,
    \item decrease of $R$ indicates effective suppression of coupling under the chosen lifting, weighting, and spectral projection.
    \item invertibility provides a bridge: the forward transform compresses coupling into spectral channels; the inverse recovers interaction structure.
\end{itemize}
The language above is operational: ``suppression'' denotes empirical reduction of cross terms, not elimination; invertibility ensures the transform is reversible and information is not lost (except through regularization choices).

Relation to established structures:
\begin{itemize}
    \item Fourier diagonalization of circulant operators motivates why Fourier coordinates can simplify coupling patterns.
    \item The radial exponential factor in \eqref{eq:weight_def} aligns with Gaussian/heat-kernel envelopes in index space.
    \item The quadratic index geometry is consistent with theta-like organizational principles used in UBT extensions.
\end{itemize}

\section{Relation to Known Structures}
\paragraph{Circulant diagonalization.}
Exact Fourier diagonalization of circulant operators is recovered in classical settings without damping-induced anisotropy.

\paragraph{Heat-kernel/Gaussian weighting.}
The factor in \eqref{eq:weight_def} mirrors radial Gaussian/heat-kernel envelopes in index space, here used as compensation rather than smoothing.

\paragraph{Theta-like organization.}
Quadratic index dependence $n^2+m^2$ is compatible with theta-function style structures; we use this relation qualitatively to motivate model geometry, not to assert modular identities.

\section{Limitations}
\begin{enumerate}
    \item Diagonalization is approximate, not exact in general.
    \item Results depend on signal structure and the compensation parameter $\phi$.
    \item No guarantee is claimed for arbitrary inputs.
    \item Performance depends on array ordering conventions and normalization choices.
    \item The rank-one idealization $\rho=a a^{\ast}$ may not hold under noise, truncation, or multi-component superposition.
    \item Spectral concentration can occur away from the main diagonal if indexing is not aligned with dominant modes.
    \item Inversion stability depends on the choice of regularization parameter $\lambda$ in \eqref{eq:regularized_inverse}; too large $\lambda$ suppresses recovery, too small allows noise amplification.
    \item The weighted metric \eqref{eq:weighted_norm} is not identical to standard Euclidean norm; energy statements must reference the appropriate metric.
\end{enumerate}

\appendix
\section{Implementation Notes}
\subsection*{Complexity}
For an $N\times N$ tensor, the 2D FFT step has complexity $\mathcal{O}(N^2\log N)$. The lifting and weighting steps are $\mathcal{O}(N^2)$.

\subsection*{Normalization Conventions}
Different FFT libraries use different normalization factors (forward, inverse, or split normalization). Report conventions explicitly when comparing $R$ across implementations.

\subsection*{Discretization Assumptions}
Indices are assumed uniformly sampled on a finite lattice with symmetric truncation $[-K,K]$. Conclusions may change under nonuniform grids or different boundary conventions.

\section{Conclusion}
We presented a compact framework defining an invertible transform with explicit forward and inverse operators. The forward path is: tensor lifting $a\mapsto\rho$, diffusion-controlled basis weighting, 2D FFT mapping, and spectral regularization via $D$. The inverse path recovers interaction structure via inverse FFT and regularized inverse weighting. Orthogonality (approximate diagonalization) is provided by the spectral properties of the 2D FFT, while the weighted metric \eqref{eq:weighted_norm} defines the geometry of interaction space. The diagonal-dominance ratio $R$ provides a quantitative test of concentration. A multiscale extension via $\phi$-integration \eqref{eq:phi_integrated_transform}--\eqref{eq:phi_integrated_discrete} further generalizes the method from single-scale Gaussian weighting to radial kernel mixtures. This framework yields a rigorous and testable pathway for UBT-aligned analysis of damped structured systems, with invertibility ensuring information preservation and numerical stability guidelines for practical deployment.

\end{document}
