% © 2025 Ing. David Jaroš — CC BY-NC-ND 4.0
%
% This work is licensed under a Creative Commons Attribution-NonCommercial-NoDerivatives 
% 4.0 International License (CC BY-NC-ND 4.0).
%
% License History: Earlier drafts (up to v0.3) were released under CC BY 4.0. 
% From v0.4 onward, all material is released under CC BY-NC-ND 4.0 to protect 
% the integrity of the theoretical work during ongoing academic development.
%
% See LICENSE.md for full license text.

\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{xcolor}
\geometry{a4paper, margin=1in}

\title{Attempting a Non-Circular Derivation of Planck's Constant\\within Unified Biquaternion Theory}
\author{Ing.~David~Jaroš}
\date{February 2026}

\begin{document}
\maketitle

% License Notice - Visible in PDF
\noindent
\textbf{License:} © 2025 Ing. David Jaroš. This work is licensed under a Creative Commons Attribution-NonCommercial-NoDerivatives 4.0 International License (CC BY-NC-ND 4.0). See \url{https://creativecommons.org/licenses/by-nc-nd/4.0/} for details.

\vspace{1em}

\begin{abstract}
We attempt a logically non-circular derivation of the numerical value of Planck's constant $\hbar$ within the framework of Unified Biquaternion Theory (UBT). This document is structured to make circularity transparent: we explicitly separate geometric quantization from numerical scale-fixing, avoid defining energy via $E = \hbar\omega$, and do not introduce $\hbar$ as a normalization factor. We demonstrate that while UBT can derive the \emph{existence} of an action quantum from topological constraints on the biquaternionic field $\Theta(q,\tau)$, fixing its \emph{numerical value} requires either (a) an additional dimensionless input from geometry/topology, or (b) empirical identification. The document concludes that UBT currently provides a structural explanation for quantization of action, but not a complete ab initio numerical prediction of $\hbar$ without external calibration.
\end{abstract}

\tableofcontents
\newpage

\section{Motivation and Scope}

\subsection{The Problem of Circularity}

In conventional quantum mechanics, Planck's constant $\hbar$ appears as a fundamental dimensional constant introduced by postulate. Standard derivations of $\hbar$ are circular in the following sense:

\begin{enumerate}
\item \textbf{Energy-frequency relation circularity}: One defines energy via $E = \hbar\omega$, then ``derives'' $\hbar$ by measuring $E$ and $\omega$ independently---but the independent measurement of $E$ itself presupposes quantum mechanics and thus $\hbar$.

\item \textbf{Normalization circularity}: One introduces $\hbar$ as a unit conversion factor between action (angular momentum) and dimensionless phase, making it definitional rather than derived.

\item \textbf{Empirical circularity}: One fits $\hbar$ to match experimental data (e.g., blackbody radiation, photoelectric effect), which is valid operationally but does not constitute a theoretical derivation.
\end{enumerate}

\subsection{Formal Definition of Circular Derivation}

A derivation of $\hbar$ is \textbf{circular} if it commits any of the following:

\begin{itemize}
\item \textbf{Type I (Energy-frequency insertion)}: Assumes $E = \hbar\omega$ as a definition before deriving $\hbar$.
\item \textbf{Type II (Action normalization insertion)}: Introduces $\hbar$ as a normalization factor for action quantization (e.g., $S = n\hbar$) and then ``derives'' $\hbar$ from the normalization.
\item \textbf{Type III (Empirical constant insertion)}: Uses empirically measured values of $\hbar$, $\alpha$, $e$, $m_e$, etc., and algebraically solves for $\hbar$, treating the algebra as derivation.
\item \textbf{Type IV (Commutator postulate)}: Assumes canonical commutation relations $[\hat{x}, \hat{p}] = i\hbar$ as a starting postulate, then extracts $\hbar$ from the postulate.
\end{itemize}

\textbf{This document explicitly avoids all four types of circularity.}

\subsection{Criteria for Success or Failure}

A \textbf{non-circular derivation} would require:
\begin{itemize}
\item Deriving the existence of quantized action from geometric/topological principles.
\item Fixing the \emph{numerical scale} of the action quantum using only dimensionless geometric/topological data (e.g., $\pi$, winding numbers, Chern classes, group ranks).
\item Avoiding insertion of $\hbar$ or any empirical constant ($e$, $\alpha$, $m_e$, etc.) except as final identification labels.
\end{itemize}

We declare success if:
\begin{itemize}
\item We derive that action is quantized: $S = n \cdot S_0$ for some fundamental $S_0$.
\item We determine $S_0$ numerically from purely dimensionless geometric/topological invariants.
\item We identify $S_0 = \hbar$ by dimensional analysis and experimental correspondence.
\end{itemize}

We declare \textbf{partial success} if:
\begin{itemize}
\item We derive quantization of action but cannot fix $S_0$ without external input.
\item We clarify what additional physical principle (if any) would be needed to complete the derivation.
\end{itemize}

We declare \textbf{honest failure} if:
\begin{itemize}
\item We identify a fundamental obstruction preventing numerical determination of $\hbar$ from geometry alone.
\item We explicitly state the need for empirical identification.
\end{itemize}

\textbf{This document accepts partial success or honest failure as valid outcomes.} The goal is logical transparency, not propaganda.

\subsection{Philosophical Position}

UBT is permitted to:
\begin{itemize}
\item Derive the existence of quantized action from topological constraints.
\item Identify the experimentally measured value of $\hbar$ with the derived action quantum.
\end{itemize}

UBT is \textbf{not} permitted to:
\begin{itemize}
\item Claim numerical prediction of $\hbar$ unless derived from dimensionless geometry/topology alone.
\item Insert $\hbar$ as a definition or normalization factor at any intermediate step.
\item Define energy via $E = \hbar\omega$ before deriving $\hbar$.
\end{itemize}

\subsection{Structural vs. Numerical Results: Key Distinction}

\fbox{\begin{minipage}{0.95\textwidth}
\textbf{Distinction Critical to Avoiding Circular Claims:}

\textbf{Structural Result (What UBT Can Derive):}
\begin{itemize}
\item \emph{Existence} of an action quantum $S_0$ from topological closure/winding in complex time.
\item Quantization condition: physical action satisfies $S/S_0 \in \mathbb{Z}$ (or $2\pi\mathbb{Z}$ depending on normalization).
\item This is a dimensionless, topological statement independent of SI units.
\end{itemize}

\textbf{Numerical Value (What Requires External Input):}
\begin{itemize}
\item The \emph{numerical value} of $S_0$ in SI units: $[S_0] = \text{J}\cdot\text{s}$.
\item Converting dimensionless geometric ratios to dimensional physical constants requires at least one external scale (e.g., Planck length, speed of light, gravitational constant).
\end{itemize}

\textbf{Identification Step (Honest and Non-Circular):}
\begin{itemize}
\item We \emph{identify} $\hbar$ as the empirically measured SI value of the topologically derived action quantum $S_0$.
\item This is not circular because we derived the \emph{structure} (existence and quantization) independently, and only use empirical data to fix the \emph{scale}.
\end{itemize}

\textbf{This distinction is repeated at the conclusion of this document.}
\end{minipage}}

\vspace{1em}

\section{Minimal Mathematical Framework}

\subsection{Biquaternion Algebra}

A \textbf{biquaternion} is an element of $\mathbb{B} := \mathbb{C} \otimes \mathbb{H}$, where $\mathbb{H}$ is the quaternion algebra and $\mathbb{C}$ is the complex numbers. Concretely:
\[
b = a_0 + a_1 i + a_2 j + a_3 k + a_4 (i \cdot i) + a_5 (i \cdot j) + a_6 (i \cdot k) + a_7 (i \cdot i \cdot j \cdot k),
\]
where $a_n \in \mathbb{R}$ and $(i, j, k)$ are quaternion units satisfying $i^2 = j^2 = k^2 = ijk = -1$.

The biquaternion algebra has:
\begin{itemize}
\item Real dimension: 8
\item Complex dimension: 4
\item Natural $\mathbb{Z}_2 \times \mathbb{Z}_2$ automorphism group
\item Conjugation operations: quaternion conjugation, complex conjugation, and their composition
\end{itemize}

\subsection{The Θ-Field and Complex Time}

UBT postulates a fundamental field $\Theta(q, \tau)$ where:
\begin{itemize}
\item $q \in \mathbb{B}^4$ is a biquaternionic spacetime coordinate
\item $\tau = t + i\psi$ is complex time, with $t \in \mathbb{R}$ (real time) and $\psi \in \mathbb{R}$ (imaginary/phase time)
\item $\Theta: \mathbb{B}^4 \times \mathbb{C} \to \mathbb{B} \otimes \text{Spin}(3,1) \otimes \text{SU}(3) \times \text{SU}(2) \times \text{U}(1)$
\end{itemize}

The field $\Theta$ carries:
\begin{enumerate}
\item Biquaternionic structure (geometric degrees of freedom)
\item Spinor indices (fermionic content)
\item Gauge indices (Standard Model symmetries)
\end{enumerate}

\textbf{Critical point}: At this stage, we have introduced \emph{no} physical constants. All structures are purely algebraic/geometric.

\subsection{Biquaternionic Metric and Curvature}

The fundamental geometric object is the biquaternionic metric:
\[
\mathcal{G}_{\mu\nu}(q, \tau) = \text{Sc}(E_\mu E_\nu^\dagger) \in \mathbb{B},
\]
where $E_\mu$ are biquaternionic tetrads and $\text{Sc}$ denotes the scalar part projection.

The real projection
\[
g_{\mu\nu} := \text{Re}(\mathcal{G}_{\mu\nu})
\]
recovers the classical metric tensor of General Relativity in the appropriate limit.

The biquaternionic connection is:
\[
\Omega_\mu(q, \tau) \in \mathbb{B},
\]
from which the classical Christoffel symbols are derived via real projection.

\textbf{Still no dimensional constants have appeared.}

\section{Cyclic States and Action as a Topological Quantity}

\subsection{Complex-Time Torus}

Consider the complex time manifold as a product torus:
\[
\mathcal{T} = S^1_t \times S^1_\psi,
\]
where:
\begin{itemize}
\item $S^1_t$ has period $T_t$ (real-time period)
\item $S^1_\psi$ has period $T_\psi$ (phase-time period)
\end{itemize}

We parameterize with angular coordinates:
\[
t = R_t \theta_t, \quad \psi = R_\psi \theta_\psi, \quad \theta_t, \theta_\psi \in [0, 2\pi).
\]

The metric on the torus (in the $(t, \psi)$ directions only) is:
\[
ds^2 = R_t^2 d\theta_t^2 + R_\psi^2 d\theta_\psi^2.
\]

The ratio
\[
\alpha_{\text{geo}} := \frac{R_t}{R_\psi}
\]
is a \textbf{dimensionless geometric parameter} characterizing the torus shape.

\subsection{Topological Quantization of Winding}

For a field $\Theta$ defined on the torus $\mathcal{T}$, topological consistency requires:
\[
\Theta(\theta_t + 2\pi, \theta_\psi) = e^{2\pi i n_t} \Theta(\theta_t, \theta_\psi),
\]
\[
\Theta(\theta_t, \theta_\psi + 2\pi) = e^{2\pi i n_\psi} \Theta(\theta_t, \theta_\psi),
\]
where $n_t, n_\psi \in \mathbb{Z}$ are winding numbers.

These are \textbf{topological quantum numbers}, derived from the fundamental group $\pi_1(\mathcal{T}) = \mathbb{Z} \times \mathbb{Z}$.

\subsection{Action as Phase Circulation}

Define the action geometrically as the integral of the phase gradient around a closed cycle:
\[
S_{\text{cycle}} := \oint_{\gamma} \nabla_\tau \arg(\Theta) \, d\tau,
\]
where $\gamma$ is a closed path on the torus.

For the fundamental cycles $\gamma_t$ (once around $S^1_t$) and $\gamma_\psi$ (once around $S^1_\psi$):
\[
S_t = \int_0^{2\pi} \frac{\partial \arg(\Theta)}{\partial \theta_t} d\theta_t = 2\pi n_t,
\]
\[
S_\psi = \int_0^{2\pi} \frac{\partial \arg(\Theta)}{\partial \theta_\psi} d\theta_\psi = 2\pi n_\psi.
\]

\textbf{Critical note on units}: These integrals are dimensionless. The factor $2\pi$ arises from the geometry of a circle, and $n_t, n_\psi \in \mathbb{Z}$ are pure integers. We have \emph{not} introduced $\hbar$ here.

\textbf{Result}: Action along fundamental cycles is quantized in units of $2\pi$ times an integer. This is a \textbf{topological quantization}, not a dynamical one.

\subsection{What This Does and Does Not Imply}

\begin{itemize}
\item \textbf{Derived}: Action is quantized topologically: $S_{\text{dimensionless}} \in 2\pi \mathbb{Z}$.
\item \textbf{Not derived}: The dimensional scale of physical action in units of $[\text{energy} \times \text{time}]$.
\item \textbf{Missing}: A bridge between dimensionless topological winding and dimensional action in SI units.
\item \textbf{Identification (external)}: If we call the fundamental action unit $S_0$, then $\hbar$ is empirically identified as the measured value of $S_0$ in SI units.
\end{itemize}

\textbf{No circularity}: We did not assume $S = n\hbar$ and then solve for $\hbar$. We derived $S/S_0 \in \mathbb{Z}$ from topology, where $S_0$ is an unknown scale to be fixed externally.

\section{Separation of Definitions to Avoid Circularity}

\subsection{Frequency vs. Energy}

In conventional quantum mechanics, one \emph{defines}:
\[
E := \hbar \omega,
\]
which is circular if we seek to derive $\hbar$.

In UBT, we define:
\begin{enumerate}
\item \textbf{Geometric frequency}: $\omega_t := 2\pi/T_t$ (cycles per unit real time)---this is purely kinematic, involving no energy.

\item \textbf{Phase frequency}: $\omega_\psi := 2\pi/T_\psi$ (cycles per unit imaginary time)---also purely geometric.

\item \textbf{Energy density}: Defined via the stress-energy tensor derived from the UBT action:
\[
\mathcal{T}_{\mu\nu} = \langle D_\mu \Theta, D_\nu \Theta \rangle_\mathbb{B} - \frac{1}{2} \mathcal{G}_{\mu\nu} \langle D\Theta, D\Theta \rangle,
\]
where $D_\mu$ is the biquaternionic covariant derivative.
\end{enumerate}

\textbf{Critical separation}: Energy is defined via field gradients and metric, \emph{not} via $E = \hbar\omega$. The relation $E \sim \hbar\omega$ (if it holds) must be \emph{derived}, not assumed.

\subsection{Where Standard Physics Implicitly Inserts $\hbar$}

In standard QM/QFT textbooks, $\hbar$ is inserted at several stages:

\begin{enumerate}
\item \textbf{Canonical commutation relations}: $[\hat{x}, \hat{p}] = i\hbar$---this is a postulate, not derived.

\item \textbf{De Broglie relations}: $p = \hbar k$, $E = \hbar \omega$---these are definitions/postulates.

\item \textbf{Path integral normalization}: $e^{iS/\hbar}$---the appearance of $\hbar$ in the exponent is a normalization choice.

\item \textbf{Semiclassical expansion}: $\hbar \to 0$ limit---treats $\hbar$ as a small parameter, not a derived quantity.
\end{enumerate}

UBT attempts to avoid these insertions by:
\begin{itemize}
\item Defining canonical structure via biquaternionic geometry, not commutation relations.
\item Deriving energy from field dynamics, not from $E = \hbar\omega$.
\item Using topological winding numbers instead of $\hbar$-normalized action.
\end{itemize}

\section{Attempted Numerical Fixing of the Action Quantum}

\subsection{Available Dimensionless Inputs}

To fix the numerical value of the action quantum, we can use only:

\begin{enumerate}
\item \textbf{Mathematical constants}: $\pi$, $e$ (Euler's number), $\gamma$ (Euler-Mascheroni constant), etc.

\item \textbf{Topological invariants}: Winding numbers $n \in \mathbb{Z}$, Chern numbers, homotopy groups.

\item \textbf{Algebraic invariants}: Dimensions of Lie algebras (e.g., $\dim(\text{SU}(3)) = 8$), structure constants, Casimir invariants.

\item \textbf{Geometric ratios}: Aspect ratio of torus $\alpha_{\text{geo}} = R_t/R_\psi$.
\end{enumerate}

\textbf{Not allowed}: $\hbar$, $c$, $G$, $e$ (electron charge), $\alpha$ (fine-structure constant), $m_e$, etc.---these are dimensional or empirical.

\subsection{Attempt 1: Direct Geometric Scaling}

Consider the energy of a fundamental mode on the torus. The kinetic energy density (from field gradients) is:
\[
\mathcal{E}_{\text{kin}} \propto \left| \frac{\partial \Theta}{\partial \theta_t} \right|^2 \frac{1}{R_t^2} + \left| \frac{\partial \Theta}{\partial \theta_\psi} \right|^2 \frac{1}{R_\psi^2}.
\]

For a mode with winding numbers $(n_t, n_\psi)$:
\[
\Theta \sim e^{i n_t \theta_t + i n_\psi \theta_\psi} \quad \Rightarrow \quad \mathcal{E}_{\text{kin}} \propto \frac{n_t^2}{R_t^2} + \frac{n_\psi^2}{R_\psi^2}.
\]

The fundamental mode $(n_t, n_\psi) = (1, 1)$ has dimensionless winding, but $\mathcal{E}_{\text{kin}}$ has dimensions $[\text{length}]^{-2}$ in geometric units.

\textbf{Problem}: We have no dimensionful scale to convert this to physical energy. We need a fundamental length $\ell_0$ or energy $E_0$.

\subsection{Attempt 2: Planck Units (Circular)}

One might try to use Planck units:
\[
\ell_{\text{Planck}} = \sqrt{\frac{\hbar G}{c^3}}, \quad E_{\text{Planck}} = \sqrt{\frac{\hbar c^5}{G}}.
\]

But these \emph{contain $\hbar$}, making this approach circular. We cannot use Planck units to derive $\hbar$.

\subsection{Attempt 3: Geometric Constraint from Curvature}

In UBT, the torus radii $R_t$ and $R_\psi$ are not arbitrary but determined by minimizing the biquaternionic action:
\[
S[\Theta, \mathcal{G}] = \int d^4q \, d\tau \sqrt{|\text{Re}(\mathcal{G})|} \left( \mathcal{L}_{\text{kin}} + \mathcal{L}_{\text{int}} + \mathcal{L}_{\text{geom}} \right).
\]

The geometric curvature contribution $\mathcal{L}_{\text{geom}}$ includes terms like:
\[
\mathcal{L}_{\text{geom}} \sim \mathcal{R}_{\mu\nu} \mathcal{R}^{\mu\nu} + \text{topological invariants},
\]
where $\mathcal{R}_{\mu\nu}$ is the biquaternionic Ricci tensor.

For a two-torus, the Euler characteristic is $\chi(\mathcal{T}) = 0$, and the Gauss-Bonnet theorem gives:
\[
\int_\mathcal{T} K \, dA = 2\pi \chi = 0,
\]
where $K$ is the Gaussian curvature.

The Gaussian curvature of the embedded torus is:
\[
K = \frac{1}{R_t R_\psi} \cos(\theta_t).
\]

\textbf{Still dimensionless}: The integral constraint relates $R_t$ and $R_\psi$ but does not fix their absolute scale.

\subsection{Attempt 4: Minimizing the Effective Potential}

Following Appendix C of UBT, the fine-structure constant is related to the torus aspect ratio:
\[
\alpha_{\text{geo}} = \frac{R_t}{R_\psi}.
\]

The effective potential for a mode with winding number $n$ is:
\[
V_{\text{eff}}(n) = A n^2 - B n \log n,
\]
where $A, B > 0$ are dimensionless coefficients from the geometry.

Minimizing over prime integers gives $n_\star = 137$, hence:
\[
\alpha_{\text{geo}} = \frac{1}{137} \approx 0.00730.
\]

This matches the fine-structure constant $\alpha \approx 1/137$, \emph{but this is dimensionless}.

\textbf{Question}: Can we use this to fix $\hbar$?

If we assume (as an \textbf{additional postulate}) that:
\[
\alpha = \frac{e^2}{4\pi \epsilon_0 \hbar c},
\]
and we know $\alpha = 1/137$ geometrically, then:
\[
\hbar = \frac{e^2}{4\pi \epsilon_0 \alpha c}.
\]

\textbf{But this requires knowing $e$, $\epsilon_0$, and $c$ independently}, which are empirical constants. This is \textbf{not} a derivation of $\hbar$ from pure geometry.

\subsection{Fundamental Obstruction}

\textbf{Dimensional analysis argument}:

Any purely geometric/topological calculation produces dimensionless numbers (like $\pi$, winding numbers, $1/137$). The quantity $\hbar$ has dimensions:
\[
[\hbar] = \text{energy} \times \text{time} = \text{mass} \times \text{length}^2 \times \text{time}^{-1}.
\]

To obtain a dimensional quantity from dimensionless geometry, we need at least one fundamental dimensional constant (e.g., Planck length, speed of light, gravitational constant).

\textbf{If we use $c$ and $G$}:
\[
\ell_{\text{Planck}} = \sqrt{\frac{\hbar G}{c^3}} \quad \Rightarrow \quad \hbar = \frac{\ell_{\text{Planck}}^2 c^3}{G}.
\]

But we still need to \emph{know} $\ell_{\text{Planck}}$ numerically, which requires either:
\begin{enumerate}
\item Empirical measurement (circular).
\item A theory that predicts $\ell_{\text{Planck}}$ from dimensionless geometry (currently lacking).
\end{enumerate}

\textbf{Conclusion from Attempt 5}: There appears to be a fundamental obstruction: \emph{dimensionless geometry alone cannot determine dimensional scales}. At least one dimensional constant must be introduced externally.

\section{Anti-Numerology and Parameter-Tuning Guardrails}

\subsection{Purpose of This Section}

This section exists to preempt accusations of:
\begin{itemize}
\item ``You just tuned numbers to hit $\alpha \approx 1/137$.''
\item ``The factor $108\pi^2/8$ is cherry-picked to match experiment.''
\item ``You varied dimension counts until you got the right answer.''
\end{itemize}

\textbf{Policy}: Any heuristic approximation or geometric ratio claimed to predict a physical constant must satisfy the following criteria.

\subsection{Criteria for Non-Numerological Claims}

\textbf{Rule 1 (Geometric object specification)}: If a ratio like $108\pi^2/8$ appears, it must be derived from a specific, independently justified geometric object (e.g., volume of a specific manifold, Chern class of a bundle, rank of a Lie algebra).

\textbf{Rule 2 (Derivation transparency)}: The appearance of numerical factors (like 108) must be traced to:
\begin{itemize}
\item Dimensions of representation spaces (e.g., $\dim(\text{adjoint SU}(3)) = 8$)
\item Winding numbers or topological invariants (integers from $\pi_1$, $\pi_2$, etc.)
\item Combinatorial factors from symmetry groups
\item Mathematical constants ($\pi$, $e$, Euler-Mascheroni $\gamma$)
\end{itemize}

\textbf{Rule 3 (No post-hoc tuning)}: Dimension counts, algebra choices, and geometric structures must be justified by independent theoretical necessity, not by ``trying different values until matching experiment.''

\textbf{Example of violation}: Saying ``we choose dimension 108 because it gives $\alpha \approx 1/137$'' is numerology. Saying ``the biquaternionic algebra has dimension 8, and the gauge group $\text{SU}(3) \times \text{SU}(2) \times \text{U}(1)$ has total dimension $8 + 3 + 1 = 12$, and their product yields specific topological invariants...'' is potentially valid if the entire chain is derived.

\subsection{Status of Current UBT Claims}

\textbf{Fine-structure constant $\alpha \approx 1/137$}:
\begin{itemize}
\item \textbf{Claimed derivation}: Torus aspect ratio $R_t/R_\psi$ minimizing an effective potential with prime-number selection.
\item \textbf{Geometric object}: The complex-time torus $S^1_t \times S^1_\psi$ with modular parameter.
\item \textbf{Where 137 comes from}: Prime minimization of $V_{\text{eff}}(n) = An^2 - Bn\log n$. The functional form of $V_{\text{eff}}$ must be derived from biquaternionic action, not asserted.
\item \textbf{Status}: If $V_{\text{eff}}$ is independently derived from curvature/winding constraints, this is non-numerological. If $V_{\text{eff}}$ is guessed to fit 137, this is numerology.
\end{itemize}

\textbf{Factor $108\pi^2/8$ (if it appears)}:
\begin{itemize}
\item \textbf{108}: Must be derived from specific representation dimensions or topological data. If it's ``adjusted to fit,'' it's numerology.
\item \textbf{$\pi^2$}: Appears naturally in 2D topology (e.g., sphere volume factors, Chern classes).
\item \textbf{8}: Could be dimension of $\mathbb{B}$, dimension of $\text{SU}(3)$ adjoint, or other algebraic invariant.
\item \textbf{Requirement}: Each factor must have independent geometric justification.
\end{itemize}

\subsection{Dependency Graph and Logical Flow}

To clarify what depends on what, we provide a logical dependency graph:

\textbf{Level 0 (Pure geometry/topology, no empirical input):}
\begin{itemize}
\item Biquaternionic algebra structure ($\mathbb{C} \otimes \mathbb{H}$, dimension 8)
\item Gauge group structure ($\text{SU}(3) \times \text{SU}(2) \times \text{U}(1)$)
\item Complex-time torus topology ($S^1_t \times S^1_\psi$)
\item Winding number quantization ($n_t, n_\psi \in \mathbb{Z}$)
\end{itemize}

\textbf{Level 1 (Derived from Level 0):}
\begin{itemize}
\item Action quantization: $S/S_0 \in 2\pi\mathbb{Z}$ (dimensionless)
\item Fine-structure constant $\alpha$ as torus aspect ratio (if $V_{\text{eff}}$ is derived)
\end{itemize}

\textbf{Level 2 (Requires one external dimensional scale):}
\begin{itemize}
\item Numerical value of $\hbar$ in SI units (requires $c$, $G$, or direct empirical measurement)
\item Conversion between geometric and physical units
\end{itemize}

\textbf{Level 3 (Requires multiple empirical inputs):}
\begin{itemize}
\item Using $\alpha = e^2/(4\pi\epsilon_0\hbar c)$ to extract $\hbar$ from $\alpha$ requires knowing $e$, $\epsilon_0$, $c$ empirically.
\end{itemize}

\textbf{Current status of $\alpha$ derivation}:
\begin{itemize}
\item If $\alpha$ is fully derived at Level 1, this is major progress.
\item Deriving $\hbar$ requires moving to Level 2 or 3, which involves external input.
\end{itemize}

\textbf{What would count as genuine ab initio numerical derivation}:
\begin{itemize}
\item Deriving \emph{all} dimensional constants ($c$, $G$, $\hbar$) from a single fundamental geometric scale (e.g., Planck length) where that scale itself is determined by pure geometry (e.g., ``the Planck length is the unique length scale where the biquaternionic curvature invariant equals $2\pi$'').
\item Currently, UBT does not have such a principle. This is an open problem, not a hidden failure.
\end{itemize}

\section{Relation to Fine-Structure Constant (Careful Analysis)}

\subsection{Can $\alpha$ Be Derived Geometrically?}

In UBT, $\alpha$ is claimed to be derived geometrically as the aspect ratio of the complex-time torus:
\[
\alpha = \frac{R_t}{R_\psi} = \frac{1}{n_\star}, \quad n_\star = 137 \text{ (from minimizing } V_{\text{eff}}).
\]

This derivation is \textbf{dimensionless} and does not involve empirical input (assuming the functional form of $V_{\text{eff}}$ is derived from biquaternionic geometry).

\subsection{Using $\alpha$ to Constrain $\hbar$}

In standard physics:
\[
\alpha = \frac{e^2}{4\pi \epsilon_0 \hbar c}.
\]

If we take $\alpha = 1/137$ from geometry and \emph{also know} $e$, $\epsilon_0$, $c$ from independent measurements, we can solve for $\hbar$:
\[
\hbar = \frac{e^2}{4\pi \epsilon_0 \alpha c}.
\]

\textbf{Is this a derivation?}

\begin{itemize}
\item \textcolor{blue}{Pro}: We derived $\alpha$ geometrically without empirical input.
\item \textcolor{red}{Con}: We used empirical values of $e$, $\epsilon_0$, $c$ to extract $\hbar$.
\end{itemize}

\textbf{Verdict}: This is a \textbf{partial derivation}. We reduce the empirical content from $\{\alpha, \hbar, e, \epsilon_0, c\}$ to $\{e, \epsilon_0, c\}$, which is progress but not complete.

\subsection{If $e$ Is Also Derived}

Suppose (hypothetically) that UBT could also derive the electron charge $e$ from geometry (e.g., via topological quantization of electromagnetic flux). Then:
\[
\Phi_{\text{magnetic}} = \oint \mathbf{A} \cdot d\mathbf{l} = n \Phi_0, \quad \Phi_0 = \frac{h}{e} = \frac{2\pi\hbar}{e}.
\]

If winding numbers fix $n$ and geometric constraints fix $\Phi_0$, we could potentially solve for $e$ and $\hbar$ simultaneously.

\textbf{Status}: UBT does not currently provide such a derivation. This remains an open question.

\section{Conclusions: What Is and Is Not Derived}

\subsection{What UBT Derives Non-Circularly}

\begin{enumerate}
\item \textbf{Existence of quantized action}: From topological constraints on the complex-time torus, action along closed cycles is quantized in units of $2\pi \mathbb{Z}$ (dimensionless). \emph{No} empirical constants or $\hbar$ were inserted.

\item \textbf{Existence of a fundamental action scale}: The geometry selects a characteristic action $S_0$ associated with the fundamental winding modes. This is a structural prediction, not a numerical one.

\item \textbf{Dimensionless fine-structure constant}: UBT derives $\alpha \approx 1/137$ from the aspect ratio of the torus and prime selection (conditional on the effective potential being independently derived from biquaternionic geometry, not guessed).
\end{enumerate}

\subsection{What UBT Does Not Currently Derive}

\begin{enumerate}
\item \textbf{Numerical value of $\hbar$ in SI units}: Without external dimensional input (e.g., $c$, $G$, or direct empirical measurement of $\hbar$), pure geometry cannot produce a dimensional constant.

\item \textbf{Independent derivation of $c$ or $G$}: These appear as conversion factors between geometric and physical units but are not derived from UBT geometry.

\item \textbf{Complete elimination of empirical input}: At least one dimensional constant (or equivalently, a fundamental length/energy/time scale) must be fixed empirically or via an additional geometric principle not yet identified.
\end{enumerate}

\subsection{Summary Table}

\begin{center}
\begin{tabular}{|l|c|c|}
\hline
\textbf{Quantity} & \textbf{Derived from UBT Geometry?} & \textbf{Method} \\
\hline
Quantization of action & \textcolor{blue}{Yes} & Topology of $S^1_t \times S^1_\psi$ \\
Fine-structure constant $\alpha$ & \textcolor{blue}{Yes (conditional)} & Torus aspect ratio, prime selection \\
Numerical value of $\hbar$ & \textcolor{red}{No} & Requires external dimensional scale \\
Speed of light $c$ & \textcolor{red}{No} & Unit conversion, not derived \\
Gravitational constant $G$ & \textcolor{red}{No} & Unit conversion, not derived \\
Electron charge $e$ & \textcolor{orange}{Open} & Potential future derivation \\
\hline
\end{tabular}
\end{center}

\subsection{What Would Count as a Genuine Non-Circular Numerical Derivation}

A complete, non-circular derivation of $\hbar$ would require:

\begin{enumerate}
\item \textbf{Geometric scale-fixing principle}: A purely geometric constraint that determines a fundamental length/time/energy scale. Example: ``The Planck length is the unique scale where the biquaternionic curvature scalar equals $2\pi$ in natural units.''

\item \textbf{Internal consistency}: All dimensional constants ($c$, $G$, $\hbar$) must be expressible in terms of this single geometric scale via dimensionless geometric ratios.

\item \textbf{No external dimensional input}: The geometric scale itself must not be defined via empirical constants.
\end{enumerate}

\textbf{Current UBT status}: No such scale-fixing principle has been identified. This is an open theoretical problem, not a hidden failure to be denied.

\subsection{Honest Identification vs. Derivation}

\fbox{\begin{minipage}{0.95\textwidth}
\textbf{UBT's Current Position (Rigorous Statement):}

\begin{itemize}
\item \textbf{Derived}: The \emph{structure} of quantum mechanics (action quantization, discrete spectrum, fine-structure constant as a dimensionless geometric ratio).
\item \textbf{Identified (not derived)}: The numerical value of $\hbar$ in SI units, obtained by matching the topologically derived action quantum $S_0$ to experimental measurements.
\item \textbf{Distinction}: This is analogous to deriving Maxwell's equations from symmetry principles (structure) but requiring empirical input to fix $\epsilon_0$ or $c$ (scale). It is honest and non-circular.
\end{itemize}
\end{minipage}}

\vspace{1em}

\subsection{Philosophical Interpretation}

UBT provides a \textbf{structural explanation} for why action is quantized and why the fine-structure constant has the value it does. This is significant progress beyond standard QM/QFT, which treats both as empirical inputs.

However, UBT does not currently provide a \textbf{complete numerical prediction} of $\hbar$ without empirical calibration. The fundamental obstruction is dimensional: geometry produces dimensionless numbers, while $\hbar$ is dimensional.

\textbf{Possible resolutions}:

\begin{enumerate}
\item \textbf{Accept empirical identification}: Take $\hbar$ as the experimentally measured scale of the derived action quantum. This is honest and scientifically defensible.

\item \textbf{Future geometric unification}: If a more complete theory can derive \emph{all} dimensional constants ($c$, $G$, $\hbar$) from a single fundamental scale (e.g., Planck length) via geometric/topological principles, then numerical prediction becomes possible.

\item \textbf{Anthropic/observational selection}: Perhaps only certain values of dimensional constants are compatible with the existence of observers, and UBT+observational constraints uniquely fix $\hbar$. This is speculative.
\end{enumerate}

\subsection{Final Verdict}

\textbf{UBT achieves partial success}: It derives the \emph{structure} of quantum mechanics (quantization of action, discrete energy levels, fine-structure constant) from geometric first principles, but requires empirical input to fix the absolute dimensional scale.

This is not a failure of UBT but rather a reflection of the nature of dimensional analysis: \emph{dimensionless geometry alone cannot determine dimensional scales without at least one external dimensional input}.

The derivation presented here is \textbf{logically non-circular} in the sense that:
\begin{itemize}
\item We never defined $E = \hbar\omega$ before deriving $\hbar$ or discussing action quantization.
\item We never introduced $\hbar$ as a normalization factor in the quantization condition.
\item We clearly separated geometric quantization (derived) from numerical scale-fixing (empirical identification).
\item We explicitly identified all four types of circularity and avoided them.
\end{itemize}

\textbf{Recommendation}: UBT should claim a geometric derivation of the \emph{structure} of quantum mechanics, with $\hbar$ identified (not derived) as the experimentally measured scale of the topologically quantized action. This is scientifically rigorous and avoids overclaiming.

\section{Open Questions and Future Directions}

\begin{enumerate}
\item \textbf{Derivation of electron charge}: Can $e$ be derived from topological quantization of electromagnetic flux in UBT? If so, this would reduce empirical input further.

\item \textbf{Unification of dimensional constants}: Can $c$, $G$, $\hbar$ be expressed in terms of a single fundamental geometric scale (e.g., Planck length) derivable from UBT?

\item \textbf{Emergent spacetime}: If spacetime itself is emergent from biquaternionic geometry, can dimensional constants arise from the emergent structure?

\item \textbf{Observational constraints}: What role do observational selection effects (anthropic reasoning) play in fixing dimensional scales?

\item \textbf{Experimental tests}: Can UBT predict \emph{deviations} from standard values of $\hbar$ in extreme regimes (high curvature, strong fields)?
\end{enumerate}

\section*{Acknowledgments}

This document benefits from the rigorous mathematical framework developed in Unified Biquaternion Theory and aims for logical transparency above rhetorical persuasion. Any errors or oversights are the responsibility of the author.

\end{document}
