
\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{bm}
\usepackage{physics}

\title{Noether $\to$ $\alpha$ v0.9: Spectral trace formulation (\texorpdfstring{$\alpha$}{alpha} as a fiber-average)}
\author{Draft for UBT Project}
\date{\today}

\begin{document}
\maketitle

\section*{Idea}
Derive $\alpha$ from UBT \emph{without specifying an external field list}. After linearizing around the UBT vacuum,
all fluctuation degrees of freedom form a fiber $\mathcal F$ on which the $U(1)$ generator $Q$ acts.
Sums $\sum_j d_j q_j^2$ are replaced by the \emph{group-invariant trace}
\begin{equation}
\sum_j d_j q_j^2 \;\longrightarrow\; \Tr_{\mathcal F}(Q^2).
\end{equation}
Equivalently, one can use the fiber-average $\langle q^2\rangle \equiv \Tr(Q^2)/\dim \mathcal F$. No external ``field content'' is needed.

\section{Setup: fluctuation operator and spin blocks}
Let $\Delta_\Theta$ be the quadratic fluctuation operator of the UBT field(s) $\Theta$ on $M^4\times S^1_\psi$ with vacuum background
$(A_\psi^\star, A(\psi),B(\psi))$. For each spin block $B$ (Dirac, scalar, vector, \ldots) the heat kernel has
\begin{equation}
\Tr \, e^{-s\Delta_B} = \int d^5x \sqrt{|g_5|}\,\frac{1}{(4\pi s)^{5/2}} \sum_{n\ge0} s^n\, a_n^{(B)}(x).
\end{equation}
Minimal $U(1)$ coupling with generator $Q$ gives bundle curvature $\Omega_{MN}=i\,Q\,F_{MN}$ and
\begin{equation}
a_2^{(B)} \supset -\,\kappa_B \,\Tr_{\mathcal F_B}\!\big(Q^2\big)\, F_{MN}F^{MN},
\qquad \kappa_B>0 \text{ (spin weight)}.
\end{equation}
Examples in minimalist conventions: $\kappa_{\rm scalar}=\tfrac{1}{12}$, $\kappa_{\rm Dirac}=\tfrac{1}{6}$.

\paragraph{Noether normalization.}
Fix the overall scale by requiring the fundamental charged excitation to carry $Q=\pm1$. This removes any free rescaling of $\Theta$ and $Q$.

\section{Route (A): Noether + holonomy (UBT-only)}
The UBT vacuum determines the holonomy and the photon norm,
\begin{equation}
\theta_H^\star = g_5 \oint A_\psi\, d\psi = g_5\,\mathcal I_\psi^\star,
\qquad
Z^\star=\int_0^{L_\psi^\star} e^{B(\psi)-2A(\psi)} |\xi_0(\psi)|^2\, d\psi.
\end{equation}
Hence
\begin{equation}
\boxed{~ g_5=\frac{\theta_H^\star}{\mathcal I_\psi^\star} \;,\qquad
\alpha(\mu_0)=\frac{1}{4\pi}\frac{\theta_H^{\star 2}}{(\mathcal I_\psi^\star)^2 Z^\star} ~}.
\end{equation}
All symbols are fixed by the UBT vacuum $(\Theta, A_\psi, A,B)$; no external inputs.

\section{Route (B): Emergent photon as a spectral trace (UBT-only)}
Picking the $a_2$-term and integrating proper-time with UV cutoff $\varepsilon=\Lambda^{-2}$ in $D=5$ gives a linear power-law piece.
Summed over spin blocks $B$,
\begin{align}
\left(\frac{1}{g_4^2}\right)_{\rm power}
&= \Lambda\, Z^\star \sum_B C_5^{(B)} \,\Tr_{\mathcal F_B}\!\big(Q^2\big),
\\[4pt]
C_5^{(B)} &\equiv \frac{2\,\kappa_B}{(4\pi)^{5/2}} \;,
\quad
\text{e.g.}\;\; C_5^{\rm scalar}=\frac{1}{6(4\pi)^{5/2}},\;\;
C_5^{\rm Dirac}=\frac{1}{3(4\pi)^{5/2}}.
\end{align}
Equivalently, with fiber-averages $\langle q^2\rangle_B \equiv \Tr_{\mathcal F_B}(Q^2)/\dim\mathcal F_B$,
\begin{equation}
\left(\frac{1}{g_4^2}\right)_{\rm power}
= \Lambda\, Z^\star \sum_B C_5^{(B)} \,\dim\mathcal F_B \,\langle q^2\rangle_B.
\end{equation}

\paragraph{KK logarithms as a spectral integral.}
Let $\rho_B(m; L_\psi^\star,\theta_H^\star)$ denote the KK spectral density of block $B$ on the UBT vacuum.
Then the finite log part can be written without enumerating fields,
\begin{equation}
\left(\frac{1}{g_4^2}\right)_{\rm logs}
= \sum_B \frac{b_B}{4\pi}\int_0^\Lambda \! dm \;\rho_B(m;L_\psi^\star,\theta_H^\star)\, \ln\!\frac{\Lambda}{m},
\end{equation}
with $b_B$ the standard 4D weights per degree of freedom of block $B$. The total induced coupling is
\begin{equation}
\boxed{\;\frac{1}{\alpha(\mu_0)}=\Lambda Z^\star \sum_B C_5^{(B)} \Tr_{\mathcal F_B}(Q^2)
\;+\; \sum_B \frac{b_B}{4\pi}\!\int_0^\Lambda \rho_B\,\ln\!\frac{\Lambda}{m}\, dm\; .}
\end{equation}
Here, \emph{every} ingredient is determined by the UBT vacuum and the $\Theta$-bundle algebra (no external lists).

\section{Consistency and ``no numerology'' checks}
\begin{itemize}
\item \textbf{Single vacuum:} The same $(L_\psi^\star,\theta_H^\star,A,B)$ must reproduce $\alpha$ \emph{and} the lightest charged eigenvalue $m_e$.
\item \textbf{Normalization:} $Q=\pm1$ for the fundamental excitation fixes the overall scale of $Q$ once and for all.
\item \textbf{Trace vs.\ sum:} $\sum_j d_j q_j^2 = \Tr_{\mathcal F}(Q^2)$, independent of basis or labeling of modes.
\item \textbf{No fits:} $\Lambda$ is the physical UV threshold of the UBT vacuum (first heavy mode/curvature), not a free parameter.
\end{itemize}

\section*{Checklist (UBT-only $\to$ number for $\alpha$)}
\begin{enumerate}
\item Solve the UBT vacuum EOM $\Rightarrow$ obtain $(L_\psi^\star,\theta_H^\star,A(\psi),B(\psi),A_\psi(\psi))$.
\item Compute $Z^\star=\int e^{B-2A}|\xi_0|^2 d\psi$ (photon zero-mode norm).
\item (Route A) Evaluate $\mathcal I_\psi^\star=\int_0^{L_\psi^\star}\! A_\psi d\psi$ and insert
$\alpha=\theta_H^{\star 2}/\big(4\pi(\mathcal I_\psi^\star)^2 Z^\star\big)$.
\item (Route B) Build the fluctuation operator $\Delta_\Theta$; decompose into spin blocks $B$; compute the fiber traces $\Tr_{\mathcal F_B}(Q^2)$ and the spectral density $\rho_B$; assemble the master formula above.
\item Apply perturbative running from $\mu_0$ to the comparison scale (e.g.\ $M_Z$).
\end{enumerate}


\section*{License}
This work is licensed under a Creative Commons Attribution 4.0 International License (CC BY 4.0).

\end{document}