
\documentclass[11pt]{article}
\usepackage[margin=2.2cm]{geometry}
\usepackage{amsmath,amssymb,amsfonts,physics}
\usepackage{tcolorbox}
\usepackage{bm}
\usepackage{siunitx}

\newcommand{\Izero}{I_0}

\title{UBT $\to$ $\alpha$ : Ultra-simple box (trace \& warp)}
\author{One-page note}
\date{\today}

\begin{document}
\maketitle
\vspace{-1.5em}

\begin{tcolorbox}[colback=white,colframe=black,title={\large Master relation (UBT-only, no field list)}]
\[
\boxed{
\quad \alpha^{-1} \;\approx\; \Izero\!\big(2\varepsilon\big)\;\underbrace{\frac{1}{(4\pi)^{3/2}}}_{\displaystyle 0.022448\ldots}\;\kappa_{\rm eff}\;\underbrace{\Tr_{\mathcal F}\!\big(Q^2\big)}_{\displaystyle D_\Theta}\quad
}
\]
\smallskip
\textbf{Where:}
\begin{itemize}
\item $\varepsilon$ is the (dimensionless) warp amplitude of $A(\psi)$ via $Z^\* = \!\int e^{-2A} d\psi = L\,\Izero(2\varepsilon)$, hence $\Lambda Z^\* = (2\pi/L)\cdot Z^\* = 2\pi\,\Izero(2\varepsilon)$.
\item $\kappa_{\rm eff}$ is the spin weight from the $a_2$ heat-kernel block of the \emph{UBT} fluctuation operator: \(\kappa_{\rm Dirac}=\tfrac{1}{3}\), \(\kappa_{\rm scalar}=\tfrac{1}{6}\). (Mixtures $\Rightarrow$ weighted average.)
\item $D_\Theta \equiv \Tr_{\mathcal F}(Q^2)$ is the \emph{invariant fiber trace} of the $U(1)$ generator $Q$ on the linearized fiber $\mathcal F$ of $\Theta$ (``average over all modes'' with no field list).
\end{itemize}
\textbf{Note.} KK logarithms give a finite, typically smaller correction and can be added as a next-order term.
\end{tcolorbox}

\vspace{0.5em}
\noindent\textbf{Why this is input-free.}
We used only: (i) the circle identity $\Lambda=2\pi/L$, $Z^\*=L\Izero(2\varepsilon)$, (ii) the heat-kernel $a_2$ coefficient in $D=5$, and (iii) the invariant trace $\Tr(Q^2)$ instead of any sum over fields. No ``choice of content'' appears.

\section*{How to compute $D_\Theta=\Tr_{\mathcal F}(Q^2)$ (from the algebra of $\Theta$)}
\begin{enumerate}
\item \textbf{Fix the charge unit:} normalize $Q$ so that the fundamental excitation has $Q=\pm 1$ (Noether). This removes any overall field rescaling.
\item \textbf{Linearize:} write the quadratic fluctuation operator of $\Theta$ at the UBT vacuum. Its fiber $\mathcal F$ is a finite-dimensional complex vector space carrying the action of $Q$.
\item \textbf{Trace instead of sum:} diagonalize $Q$ on $\mathcal F$ (purely algebraic). If $Q\ket{i}=q_i\ket{i}$ with degeneracy $d_i$, then
\[
\Tr_{\mathcal F}(Q^2)=\sum_i d_i\,q_i^2 \qquad \text{(basis independent)}.
\]
\item \textbf{Typical quick evaluations:}
\begin{itemize}
\item If the fiber splits into $N$ modes with $q=\pm 1$ (any pattern of signs), then $\Tr(Q^2)=N$.
\item If the algebra enforces a uniform spectrum $\{\pm q\}$ with multiplicity $N$, then $\Tr(Q^2)=N q^2$.
\item If the fiber is grouped into blocks (Dirac/scalar), compute the trace \emph{per block} and combine with $\kappa_{\rm eff}$ by weights.
\end{itemize}
\end{enumerate}

\section*{Ready-to-evaluate forms}
\begin{align*}
\alpha^{-1} &\approx \Izero(2\varepsilon)\;\frac{1}{(4\pi)^{3/2}}\;\Big(\tfrac{1}{3}\,D_\Theta^{\rm Dirac}+\tfrac{1}{6}\,D_\Theta^{\rm scalar}\Big),\\
\text{or (single block)}\quad \alpha^{-1} &\approx \Izero(2\varepsilon)\;\frac{\kappa_{\rm eff}}{(4\pi)^{3/2}}\;D_\Theta.
\end{align*}

\section*{Back-of-the-envelope targets}
To reach $\alpha^{-1}\!\approx\!137$ at $\varepsilon=0$ one needs $D_\Theta \sim \mathcal O(10^4)$ for a single dominant block. Moderate warping ($\varepsilon\!\sim\!1$) reduces this by a factor $\Izero(2)\!\approx\!2.279$.

\section*{Minimal correction (optional)}
Add KK logs from the same spectrum as a finite next-order piece:
\[
\Delta\!\big(\alpha^{-1}\big)_{\rm logs} = \sum_B \frac{b_B}{4\pi}\int_0^{\Lambda}\!\rho_B(m)\,\ln\!\frac{\Lambda}{m}\,dm \quad (\text{small against the power-law term}).
\]

\end{document}
