\documentclass[11pt,a4paper]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\newtheorem{theorem}{Theorem}
\newtheorem{assumption}{Assumption}

\title{Unified Biquaternion Theory: Fit-Free Alpha Baseline\\
\large arXiv Preprint}
\author{UBT Team\\
Principal Investigator: Ing. David Jaroš}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
We present the fit-free baseline result for the fine-structure constant $\alpha$ derivation within the Unified Biquaternion Theory (UBT). Under three standard assumptions—geometric locking of mode parameters (A1), standard renormalization scheme (A2), and Thomson-limit extraction (A3)—we rigorously establish that the two-loop counterterm renormalization factor equals unity: $\mathcal{R}_{\mathrm{UBT}} = 1$. This yields $B = \frac{2\pi N_{\mathrm{eff}}}{3R_\psi}$ and $\alpha^{-1} = F(B)$ with no fitted parameters. We provide complete derivations, reproducibility checks, and a replication protocol for independent verification.
\end{abstract}

\tableofcontents

\section{Introduction}

The fine-structure constant $\alpha \approx 1/137.036$ is one of the fundamental dimensionless constants of nature. In standard quantum field theory, it enters as a free parameter. The Unified Biquaternion Theory (UBT) provides a geometric framework where $\alpha$ emerges from the topology of complex time.

This manuscript establishes the \textbf{fit-free baseline}: at two-loop order in perturbation theory, under well-defined assumptions A1--A3, the renormalization factor $\mathcal{R}_{\mathrm{UBT}}$ equals unity. This result eliminates free parameters from the alpha derivation within this framework.

\subsection{Main Result}

\begin{theorem}[CT Baseline, Informal]
Under assumptions A1 (geometric locking), A2 (standard CT scheme with Ward identities), and A3 (Thomson-limit extraction), the finite renormalized ratio at two loops equals:
\[
\boxed{\mathcal{R}_{\mathrm{UBT}} = 1}
\]
\end{theorem}

This implies:
\[
B = \frac{2\pi N_{\mathrm{eff}}}{3R_\psi} \times \mathcal{R}_{\mathrm{UBT}} = \frac{2\pi N_{\mathrm{eff}}}{3R_\psi}
\]
with no adjustable parameters.

\subsection{Scope and Limitations}

\textbf{What this result establishes:}
\begin{itemize}
\item A rigorous baseline for $\mathcal{R}_{\mathrm{UBT}}$ at two-loop order
\item A fit-free framework under explicit assumptions
\item Compatibility with standard QED in the real-time limit
\end{itemize}

\textbf{What this result does NOT claim:}
\begin{itemize}
\item Complete derivation of $\alpha$ from first principles (requires validating A1--A3 and computing $F(B)$)
\item Three-loop or higher-order results
\item Explicit complex-time effects beyond the stated assumptions
\end{itemize}

\section{Assumptions}

We state the three core assumptions explicitly:

\begin{assumption}[A1: Geometric Locking]
\label{ass:A1}
The effective number of modes $N_{\mathrm{eff}}$ and the imaginary-time radius $R_\psi$ are uniquely determined by:
\begin{itemize}
\item Mode counting on the Hermitian slice (Appendix P6)
\item Periodicity condition $\psi \sim \psi + 2\pi$
\item Lorentz invariance on the Hermitian slice
\item Thomson-limit normalization
\end{itemize}
No alternative values satisfy all four conditions simultaneously.
\end{assumption}

\begin{assumption}[A2: CT Renormalization Scheme]
\label{ass:A2}
The complex-time (CT) renormalization scheme satisfies:
\begin{itemize}
\item Dimensional regularization in $d = 4 - 2\epsilon$ dimensions
\item Ward identities: $Z_1 = Z_2$ (vertex $=$ field renormalization)
\item Reduction to QED in the real-time limit ($\psi \to 0$)
\item Standard counterterm structure up to scheme ambiguities
\end{itemize}
\end{assumption}

\begin{assumption}[A3: Observable Extraction]
\label{ass:A3}
The coupling constant is extracted via:
\begin{itemize}
\item Thomson scattering limit at $q^2 = 0$
\item Gauge-invariant observable definition
\item Transverse photon polarization (longitudinal parts vanish)
\end{itemize}
\end{assumption}

\section{CT Baseline Theorem}

The full proof is provided in the accompanying appendices. We state the theorem precisely:

\begin{theorem}[CT Baseline]
\label{thm:ct-baseline}
Under Assumptions~\ref{ass:A1}, \ref{ass:A2}, and \ref{ass:A3}, the finite renormalized two-loop ratio
\[
\mathcal{R}_{\mathrm{UBT}} := \frac{\Pi^{(2)}_{\text{CT, finite}}(q^2=0;\mu)}{\Pi^{(2)}_{\text{QED, finite}}(q^2=0;\mu)} \times \mathcal{N}_{\text{CT}\to\text{QED}}
\]
equals unity at two-loop order:
\[
\mathcal{R}_{\mathrm{UBT}} = 1 + \mathcal{O}(\alpha^3).
\]
\end{theorem}

\subsection{Proof Sketch}

The proof proceeds in three steps:

\textbf{Step 1: CT $\to$ QED reduction.} In the limit $\psi \to 0$, the complex-time scheme reduces to standard QED. All complex-time-dependent propagators and vertices reduce to their real-time counterparts.

\textbf{Step 2: Ward identities.} Gauge invariance enforces $Z_1 = Z_2$ up to $\mathcal{O}(\alpha^3)$. This eliminates scheme-dependent finite parts in the counterterms.

\textbf{Step 3: Thomson normalization.} Fixing the coupling at the physical observable $q^2 = 0$ via Thomson scattering provides a gauge-invariant extraction. Combined with Steps 1--2, this uniquely determines the finite ratio to be unity.

For complete details, see the included appendices from the main UBT consolidation document.

\section{Geometric Inputs}

\input{../consolidation_project/alpha_two_loop/tex/geometric_inputs_proof.tex}

\section{CT Scheme Definition}

\input{../consolidation_project/alpha_two_loop/tex/ct_scheme_definition.tex}

\section{$\mathcal{R}_{\mathrm{UBT}}$ Extraction}

\input{../consolidation_project/alpha_two_loop/tex/R_UBT_extraction.tex}

\section{Reproducibility Checklist}

\input{reproducibility_checklist.tex}

\section{Conclusions}

We have established a fit-free baseline for the fine-structure constant derivation in UBT:
\begin{itemize}
\item \textbf{Result}: $\mathcal{R}_{\mathrm{UBT}} = 1$ at two loops under A1--A3
\item \textbf{Implication}: $B = \frac{2\pi N_{\mathrm{eff}}}{3R_\psi}$ with no free parameters
\item \textbf{Requirement}: Any $\mathcal{R}_{\mathrm{UBT}} \neq 1$ must be derived by explicit CT calculation
\end{itemize}

This baseline provides a rigorous foundation for further development of UBT's predictive framework.

\subsection{Future Work}

\begin{enumerate}
\item Complete three-loop $\mathcal{R}_{\mathrm{UBT}}$ calculation
\item Finalize the function $F(B)$ for numeric prediction
\item Extend to weak mixing angle $\theta_W$ and strong coupling $\alpha_s$
\item Develop experimental tests of complex-time signatures
\end{enumerate}

\section*{Acknowledgments}

We thank the open-source community for code review and replication efforts. This work benefits from public scrutiny and collaborative verification.

\section*{Data Availability}

All source code, LaTeX documents, test suites, and build artifacts are publicly available at:

\url{https://github.com/DavJ/unified-biquaternion-theory}

Replication protocol: \texttt{docs/REPLICATION\_PROTOCOL.md}

\bibliographystyle{plain}
\bibliography{../consolidation_project/references}

\end{document}
