% © 2026 Ing. David Jaroš — CC BY-NC-ND 4.0
%
% This work is licensed under a Creative Commons Attribution-NonCommercial-NoDerivatives
% 4.0 International License (CC BY-NC-ND 4.0).
%
% License History: Earlier drafts (up to v0.3) were released under CC BY 4.0.
% From v0.4 onward, all material is released under CC BY-NC-ND 4.0 to protect
% the integrity of the theoretical work during ongoing academic development.

% =====================================================================
% TASK 4 — Gravitons in Schwarzschild Background
% Status: [OPEN] — research direction initiated in v57
% Source: DERIVATION_INDEX.md §"Gravitational Physics"
% Layer: [L2] — depends on temporal sector of Schwarzschild Θ₀
% =====================================================================

\documentclass[12pt]{article}
\usepackage[a4paper, margin=2.5cm]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% Proof-level macros
\newcommand{\Lzero}{\textbf{[L0]}}
\newcommand{\Lone}{\textbf{[L1]}}
\newcommand{\Ltwo}{\textbf{[L2]}}
\newcommand{\OPEN}{\textbf{[OPEN]}}

\title{\textbf{Gravitons in Schwarzschild Background\\
       from UBT Linearisation}}
\author{Ing. David Jaroš}
\date{March 2026 (v57 — research direction)}

\begin{document}
\maketitle

\begin{abstract}
We initiate the study of graviton perturbations around the Schwarzschild
background $\Theta_0^{(S)}$ in Unified Biquaternion Theory (UBT).
The flat-background case ($k^2 = 0$) is proved at level [L0].
The curved case requires linearising the UBT equation
$\nabla^\dagger\nabla\Theta = 0$ around the explicit Schwarzschild
solution $\Theta_0^{(S)} = e^{i\Phi(r)}\cdot[f(r)\mathbf{1}+g(r)\boldsymbol{e}_r]$.
This is an open problem at [L2].
\end{abstract}

% ======================================================================
\section{Background and Motivation}
% ======================================================================

\textbf{Known result (flat background):}
Linearising $\nabla^\dagger\nabla\Theta = 0$ around $\Theta_0 = \mathbf{1}$
gives dispersion relation $k^2 = 0$ \Lzero\ for spin-2 perturbations
(see \texttt{research\_tracks/research/graviton\_from\_perturbation.tex},
\texttt{canonical/geometry/gr\_completion\_attempt.tex}).

\textbf{New goal:}
Now that the full Schwarzschild $\Theta_0^{(S)}$ is available
(\texttt{canonical/geometry/biquaternionic\_vacuum\_solutions.tex §3},
\Lone), we can attempt gravitons in curved background.

% ======================================================================
\section{Setup: Linearisation Around Schwarzschild}
\label{sec:setup}
% ======================================================================

\subsection{Perturbation Ansatz}

Write the full field as
\begin{equation}
  \Theta = \Theta_0^{(S)} + \varepsilon\,\Theta_1,
  \label{eq:pert_ansatz}
\end{equation}
where $\Theta_0^{(S)}(r,\tau)$ is the exact Schwarzschild solution (Proved [L1]):
\begin{equation}
  \Theta_0^{(S)} = e^{i\Phi(r)}\cdot\bigl[f(r)\,\mathbf{1}
    + g(r)\,\boldsymbol{e}_r\bigr],
  \quad \Phi(r) = \frac{1-M/2r}{1+M/2r},\quad
  g(r) = r\Psi(r)^2,\quad \Psi = 1+\frac{M}{2r}.
  \label{eq:theta0_schw}
\end{equation}
and $\Theta_1$ is a small perturbation (the graviton field in curved background).

\subsection{Linearised UBT Equation}

The full UBT equation is
\begin{equation}
  \nabla^\dagger\nabla\Theta = 0,
  \label{eq:ubt_eq}
\end{equation}
where $\nabla^\dagger\nabla$ is the biquaternionic Laplacian in the
$\mathcal{G}_{\mu\nu}[\Theta_0^{(S)}]$ background metric.

Expanding to first order in $\varepsilon$:
\begin{equation}
  \nabla^\dagger_{(0)}\nabla_{(0)}\Theta_1
  + \bigl(\delta\nabla^\dagger\bigr)\nabla_{(0)}\Theta_0^{(S)}
  + \nabla^\dagger_{(0)}\bigl(\delta\nabla\bigr)\Theta_0^{(S)}
  = 0,
  \label{eq:linearised_eq}
\end{equation}
where $\nabla_{(0)}$ is the covariant derivative in the Schwarzschild
background and $\delta\nabla$ denotes first-order corrections from
the metric perturbation $\delta g_{\mu\nu} = g_{\mu\nu}[\varepsilon\Theta_1]$.

% ======================================================================
\section{Expected Result and Open Status}
\label{sec:expected}
% ======================================================================

\subsection{What the Linearised Equation Should Give}

In standard GR, linearisation of the vacuum Einstein equations around the
Schwarzschild metric gives the \emph{Regge-Wheeler equation} for odd-parity
perturbations and the \emph{Zerilli equation} for even-parity perturbations.

\textbf{Conjecture:} The linearised UBT equation \eqref{eq:linearised_eq}
should reproduce these equations in the appropriate sector, confirming
that UBT gravitational waves match standard GR in curved backgrounds.

\subsection{Technical Obstacles}

\begin{enumerate}
  \item The Schwarzschild background $\Theta_0^{(S)}$ has position-dependent
        phase $\Phi(r)$ — the linearisation is non-trivial and requires
        careful treatment of the phase-cancellation identity.
  \item The biquaternionic Laplacian $\nabla^\dagger\nabla$ in curved
        background involves Christoffel symbols from $\Psi(r)^4\delta_{ij}$
        spatial metric — straightforward to compute but tedious.
  \item The temporal sector ($g_{tt} = -\Phi(r)^2$, non-trivial $r$-dependence)
        introduces mixing between radial and temporal perturbations.
\end{enumerate}

\subsection{Status}

\begin{center}
\begin{tabular}{|l|l|}
\hline
\textbf{Result} & \textbf{Status} \\
\hline
Graviton $k^2=0$ (flat background) & \Lzero\ Proved \\
Schwarzschild $\Theta_0^{(S)}$ (background) & \Lone\ Proved \\
Linearised UBT around $\Theta_0^{(S)}$ & \OPEN\ [L2] \\
Regge-Wheeler / Zerilli from UBT & \OPEN\ [L2] \\
\hline
\end{tabular}
\end{center}

% ======================================================================
\section{Next Steps}
\label{sec:nextsteps}
% ======================================================================

\begin{enumerate}
  \item Compute $\nabla^\dagger_{(0)}\nabla_{(0)}$ in Schwarzschild isotropic
        coordinates $g_{ij} = \Psi^4\delta_{ij}$, $g_{tt} = -\Phi^2$.
  \item Substitute $\Theta_1 = \Theta_1(\mathbf{x},\tau)$ in \eqref{eq:linearised_eq}
        and separate angular modes (spin-2 spherical harmonics).
  \item Check whether the $\ell=2$ mode satisfies the Regge-Wheeler equation
        (odd parity) and the Zerilli equation (even parity).
  \item If yes: upgrade status to \Lone\ and promote to canonical.
  \item If no: document the deviation from standard GR and analyse implications.
\end{enumerate}

\textbf{DERIVATION\_INDEX.md} row to update:
``Graviton dispersion in Schwarzschild background'' — currently \OPEN.

\end{document}
