% © 2026 Ing. David Jaroš — CC BY-NC-ND 4.0
%
% This work is licensed under a Creative Commons Attribution-NonCommercial-NoDerivatives
% 4.0 International License (CC BY-NC-ND 4.0).
%
% License History: Earlier drafts (up to v0.3) were released under CC BY 4.0.
% From v0.4 onward, all material is released under CC BY-NC-ND 4.0 to protect
% the integrity of the theoretical work during ongoing academic development.

% =====================================================================
% Hosotani Mechanism: Higgs from Wilson Line on ψ-Circle
% Status: [OPEN/IN PROGRESS] — v63 research direction
% Source: DERIVATION_INDEX.md Gap H1
% Layer: [L2] — depends on KK spectrum and R_ψ calibration
% =====================================================================

\documentclass[12pt]{article}
\usepackage[a4paper, margin=2.5cm]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% Proof-level macros
\newcommand{\Lzero}{\textbf{[L0]}}
\newcommand{\Lone}{\textbf{[L1]}}
\newcommand{\Ltwo}{\textbf{[L2]}}
\newcommand{\OPEN}{\textbf{[OPEN]}}

\title{\textbf{Hosotani Mechanism on the $\psi$-Circle:\\
       Higgs Potential from Wilson Line Condensation in UBT}}
\author{Ing. David Jaroš}
\date{March 2026 (v63 — research direction)}

\begin{document}
\maketitle

\begin{abstract}
The free kinetic action $S[\Theta] = \int \text{Sc}[(\nabla^\dagger\Theta)^\dagger(\nabla^\dagger\Theta)]$
gives $a_4 = 0$ at tree and one-loop level (Gap H1 Dead End; see
\texttt{higgs\_yukawa\_scan.md}~§2.3).
We investigate whether the Hosotani mechanism — spontaneous symmetry breaking
driven by non-trivial Wilson lines along the compact $\psi$-circle — can generate
a non-zero quartic coefficient $\lambda$ and a Higgs vacuum expectation value $v$
within the UBT framework.
The $\psi$-circle radius $R_\psi$ calibrated from the fine-structure constant
($n_R = 138$, $R_\psi \approx \hbar/(m_e c \cdot 137)$) sets the mass scale.
\end{abstract}

\tableofcontents

% ======================================================================
\section{Background and Motivation}
\label{sec:background}
% ======================================================================

\subsection{Gap H1 — Current Status}

The Gap H1 problem asks: \emph{does the UBT action $S[\Theta]$ generate
the quartic term $\lambda|\Theta|^4$ of the Higgs potential?}

The v63 analysis (Dead End, \texttt{higgs\_yukawa\_scan.md}~§2.3) showed:
\begin{itemize}
  \item \textbf{Tree level}: A constant background $\Theta_0 = v \cdot \mathbf{1}$
        satisfies $\nabla^\dagger\Theta_0 = 0$, so $S[\Theta_0]$ is independent
        of $v$.  Therefore $a_4^{\text{tree}} = 0$.
  \item \textbf{One-loop}: The Kaluza-Klein spectrum on the $\psi$-circle
        is $\omega_n^2 = n^2/R_\psi^2$, independent of $|\Theta_0|$.
        The Coleman-Weinberg sum gives a $v$-independent constant, so
        $a_4^{\text{1-loop}} = 0$.
\end{itemize}

\textbf{Conclusion}: Non-trivial $\lambda$ requires either (a) explicit
interaction terms in $S[\Theta]$, or (b) a dynamical mechanism that
couples the background to the KK spectrum.  The Hosotani mechanism
provides option (b) via Wilson lines along $\psi$.

\subsection{Hosotani Mechanism (1983)}

In Kaluza-Klein theories with a compact extra dimension of circumference
$L = 2\pi R$, gauge fields can develop a non-zero vacuum expectation value
for the extra-dimensional component, $\langle A_\psi \rangle \neq 0$.
This \emph{Wilson line} (or Hosotani) background is gauge-invariant:
\begin{equation}
  W = \mathcal{P}\exp\!\left(ig \oint_0^{2\pi R_\psi}\!\! A_\psi\,d\psi\right)
  \in G_{\text{SM}},
\end{equation}
and shifts the KK masses through the covariant momentum
$p_n \to (n + \theta_a)/R_\psi$ for KK level $n$ and Wilson line angle
$\theta_a$.  The resulting effective potential $V(\theta_a)$ is
radiatively generated and may have a minimum at $\theta_a \neq 0$,
signalling spontaneous symmetry breaking without a tree-level mass term.

In UBT, the compact imaginary-time direction $\psi \in [0, 2\pi R_\psi)$
plays the role of the Hosotani extra dimension.

% ======================================================================
\section{UBT Setup: $A_\psi$ Identification}
\label{sec:setup}
% ======================================================================

\subsection{Covariant Derivative in UBT}

From the canonical SM gauge structure
(\texttt{canonical/interactions/sm\_gauge.tex}), the UBT covariant
derivative is
\begin{equation}
  D_\mu = \partial_\mu + ig_s T^a G^a_\mu + ig\,\tau^i W^i_\mu + ig' Y B_\mu,
\end{equation}
where $\mu$ runs over the four real coordinates $(t, x, y, z)$ and the
imaginary-time direction $\psi$.  The $\psi$-component gives the
extra-dimensional gauge field
\begin{equation}
  A_\psi = W^i_\psi\,\tau^i + B_\psi\,Y
  \quad \in\; \mathfrak{su}(2)_L \oplus \mathfrak{u}(1)_Y.
\end{equation}

\subsection{Wilson Line Holonomy}

The gauge-invariant Wilson line along the closed $\psi$-circle is
\begin{equation}
  \label{eq:wilson_line}
  W = \exp\!\left(2\pi i R_\psi\,g\,W^i_\psi\,\tau^i\right)
      \cdot \exp\!\left(2\pi i R_\psi\,g'\,B_\psi\,Y\right).
\end{equation}
In the $SU(2)_L$ sector we write $W^i_\psi\,\tau^i = \theta_W\,\hat{n}\cdot\boldsymbol{\tau}/(gR_\psi)$,
so that
\begin{equation}
  W_{\text{SU(2)}} = \exp(2\pi i\,\theta_W\,\hat{n}\cdot\boldsymbol{\tau}).
\end{equation}
The single real parameter $\theta_W \in [0, \pi)$ controls spontaneous
symmetry breaking: $\theta_W = 0$ is the unbroken phase; $\theta_W \neq 0$
breaks $SU(2)_L$ to a residual $U(1)$.

\subsection{Z\texorpdfstring{$_2$}{2} Mirror-Sector Constraint}

UBT has a mirror symmetry between KK modes $n^* = 137$ and $n^{**} = 139$
about $n_R = 138$.  Under $\psi \to -\psi$ (the $Z_2$ reflection):
\begin{equation}
  A_\psi \to -A_\psi \implies \theta_W \to -\theta_W.
\end{equation}
$Z_2$ symmetry forces the effective potential $V(\theta_W)$ to be even:
\begin{equation}
  V(\theta_W) = V(-\theta_W).
\end{equation}
A non-trivial minimum therefore occurs at $\theta_W = \pi/2$ (maximal
mixing) if $V''(\theta_W)\big|_{\theta_W=0} < 0$, i.e.\ the origin is
a local maximum.

% ======================================================================
\section{Effective Potential for the Wilson Line}
\label{sec:veff}
% ======================================================================

\subsection{General Form}

At one-loop in a five-dimensional theory compactified on $S^1$ of radius
$R$, the Hosotani effective potential from a single multiplet (mass $M_n$
at KK level $n$) is
\begin{equation}
  \label{eq:hosotani_veff_general}
  V(\theta) = -\frac{N}{(2\pi R)^5}
              \sum_{n=-\infty}^{+\infty} \int \frac{d^4k}{(2\pi)^4}
              \ln\!\left[k^2 + \frac{(n+\theta)^2}{R^2}\right],
\end{equation}
where $N > 0$ for bosons (negative contribution from sign flip)
and $N < 0$ for fermions.  After dimensional regularisation and
Poisson resummation the Coleman-Weinberg integral reduces to
\begin{equation}
  \label{eq:hosotani_veff}
  V(\theta) = \frac{3}{16\pi^6 R^5}
              \sum_{k=1}^{\infty}
              \frac{N_{\text{eff}}\,\cos(2\pi k\,\theta)}{k^5},
\end{equation}
where $N_{\text{eff}} = N_b - 4N_f$ counts boson and fermion degrees
of freedom with appropriate Clifford factors ($N_b$ bosons, $N_f$
Weyl fermions, factor 4 from $\gamma$-trace in $d=4$).

\subsection{Application to UBT $\psi$-Circle}

Setting $R = R_\psi$ and restricting to the SU(2)$_L$ Wilson line
$\theta = \theta_W$:
\begin{equation}
  \label{eq:hosotani_ubt}
  V(\theta_W) = \frac{3}{16\pi^6 R_\psi^5}
                \sum_{k=1}^{\infty}
                \frac{N_{\text{eff}}\,\cos(2\pi k\,\theta_W)}{k^5}.
\end{equation}

\noindent\textbf{Relevant KK spectrum}: The mirror sector analysis
(\texttt{higgs\_yukawa\_scan.md}~§3.1) gives $n_R = 138$, with the
modes near $n_R$ contributing the most to the $\theta_W$-dependent
part of the potential.  Modes with $|n| \gg n_R$ contribute a
$\theta_W$-independent constant to $V$.

\subsection{$N_{\text{eff}}$ Count at $n_R = 138$}
\label{sec:neff_count}

% v66: N_eff = 8 at n_R=138 (bosonic); SSB confirmed [L1]

\subsubsection{KK Parity Rule (from \texttt{fermionic\_statistics\_bridge.tex})}

The canonical bridge document
(\texttt{canonical/bridges/fermionic\_statistics\_bridge.tex})
proves the KK parity rule via the Grassmannian path-integral measure:

\begin{proposition}[KK Parity Rule, {\normalfont\textbf{[L1]}}]
Under exchange of two identical KK modes at winding number $n$, the
path-integral acquires the geometric phase $(-1)^n$.  Consequently:
\begin{align}
  n \text{ even} &\implies (-1)^n = +1 \implies \text{bosonic statistics},\\
  n \text{ odd}  &\implies (-1)^n = -1 \implies \text{fermionic statistics}.
\end{align}
\end{proposition}

This rule assigns a definite statistics to every KK level from the
topology of the $\psi$-circle alone, without additional input.

\subsubsection{Degrees of Freedom of $\Theta \in \mathbb{C}\otimes\mathbb{H}$}

The UBT field satisfies $\Theta \in \mathbb{C}\otimes\mathbb{H} \cong
\mathrm{Mat}(2,\mathbb{C})$.  The real dimension is
\begin{equation}
  \dim_{\mathbb{R}}(\mathbb{C}\otimes\mathbb{H})
  = \dim_{\mathbb{R}}(\mathrm{Mat}(2,\mathbb{C})) = 8.
  \label{eq:dimR_eight}
\end{equation}
Each independent real component of $\Theta$ constitutes one degree
of freedom; therefore every KK level carries exactly $\mathbf{8}$
real degrees of freedom.

\subsubsection{Count at $n_R = 138$}

At KK level $n_R = 138$ (even):

\begin{itemize}
  \item By the KK parity rule, $n_R = 138$ is \textbf{bosonic}: all
        8 real DOF of $\mathbb{C}\otimes\mathbb{H}$ are bosonic.
  \item The adjacent levels $n^* = 137$ and $n^{**} = 139$ are
        \textbf{odd}, hence fermionic; they do not contribute bosonic
        modes at $n_R = 138$.
\end{itemize}

Therefore the mode count \emph{at the mirror level $n_R = 138$} is:
\begin{equation}
  \label{eq:neff_count}
  \boxed{N_b = 8,\quad N_f = 0,\quad
  N_{\text{eff}} = N_b - 4N_f = 8 - 0 = 8 > 0.}
\end{equation}

\subsubsection{Consequence: SSB at $\theta_W = \pi/2$}

With $N_{\text{eff}} = 8 > 0$ inserted into the mass-term coefficient
\eqref{eq:a2_hosotani}:
\begin{equation}
  a_2^{\text{Hos}} = -\frac{3 \times 8 \times \zeta(3)}{8\pi^4 R_\psi^5}
                   = -\frac{3\,\zeta(3)}{\pi^4 R_\psi^5} < 0.
\end{equation}
This is \textbf{negative}, confirming the SSB condition
\eqref{eq:ssb_condition}: the origin $\theta_W = 0$ is a local
maximum and the $Z_2$ constraint forces the minimum to
$\theta_W = \pi/2$.

\begin{remark}
The $N_{\text{eff}}$ count uses only the $\mathbb{R}$-dimension of
$\mathbb{C}\otimes\mathbb{H}$ (8 DOF) and the topological KK parity
rule (even $n \to$ bosonic).  Both inputs are [L1] results;
the conclusion $N_{\text{eff}} = 8 > 0$ is therefore a \Lone\ result.
\end{remark}

\subsection{Expansion Near $\theta_W = 0$}

Using $\cos(2\pi k\,\theta_W) = 1 - 2\pi^2 k^2\,\theta_W^2
+ \tfrac{2}{3}\pi^4 k^4\,\theta_W^4 + \cdots$ and the standard sums
$\sum_{k=1}^\infty k^{-3} = \zeta(3)$,
$\sum_{k=1}^\infty k^{-1} = \zeta(1)_{\text{reg}}$ (requires
zeta-function or Pauli-Villars regularisation; see §4.2):
\begin{equation}
  V(\theta_W) = V_0
    - \frac{3N_{\text{eff}}}{8\pi^4 R_\psi^5}\,\zeta(3)\,\theta_W^2
    + \frac{N_{\text{eff}}}{8\pi^2 R_\psi^5}\,\zeta(1)_{\text{reg}}\,\theta_W^4
    + O(\theta_W^6).
\end{equation}

\noindent\textbf{Mass term coefficient}:
\begin{equation}
  \label{eq:a2_hosotani}
  a_2^{\text{Hos}} = -\frac{3N_{\text{eff}}\,\zeta(3)}{8\pi^4 R_\psi^5}.
\end{equation}
For SSB we need $a_2 < 0$, i.e.\ $N_{\text{eff}} > 0$
(bosons dominate over fermions).

\noindent\textbf{Quartic coefficient}:
\begin{equation}
  \label{eq:a4_hosotani}
  a_4^{\text{Hos}} = \frac{N_{\text{eff}}\,\zeta(1)_{\text{reg}}}{8\pi^2 R_\psi^5}.
\end{equation}
The regularised value $\zeta(1)_{\text{reg}}$ requires a UV cutoff or
zeta-function regularisation; at the scale $1/R_\psi$ this gives a
finite positive number for $N_{\text{eff}} > 0$.

% ======================================================================
\section{Minimum and SSB Condition}
\label{sec:minimum}
% ======================================================================

\subsection{Condition for Non-Trivial Minimum}

The potential $V(\theta_W)$ has a non-trivial minimum at
$\theta_W \neq 0$ if and only if
\begin{equation}
  \label{eq:ssb_condition}
  N_{\text{eff}} > 0
  \quad\text{(bosons dominate)}
\end{equation}
and the $Z_2$ symmetry forces the minimum to $\theta_W = \pi/2$
(or a nearby value determined by higher-order terms).

\subsection{Higgs VEV from Wilson Line}

Once $\theta_W \neq 0$, the zero-mode of $A_\psi$ acquires a VEV
\begin{equation}
  \langle A_\psi \rangle = \frac{\theta_W}{g R_\psi}.
\end{equation}
This is identified with the Higgs VEV $v$ through the relation
\begin{equation}
  \label{eq:higgs_vev}
  v = \frac{\theta_W}{g R_\psi}\,.
\end{equation}
For $\theta_W = \pi/2$ (maximal mixing from $Z_2$):
\begin{equation}
  v = \frac{\pi}{2\,g\,R_\psi}.
\end{equation}

\subsection{Quartic Coupling $\lambda$}

The quartic Higgs coupling in the standard parametrisation
$V_{\text{Higgs}}(H) = \lambda(H^\dagger H - v^2/2)^2$ is
related to the Wilson-line potential by
\begin{equation}
  \label{eq:lambda_hosotani}
  \lambda = \frac{1}{2}\frac{d^2 V}{d\theta_W^2}\bigg|_{\theta_W = \theta_W^{\min}}
            \cdot \left(\frac{g R_\psi}{\theta_W^{\min}}\right)^2.
\end{equation}
Evaluating at $\theta_W^{\min} = \pi/2$ with
\eqref{eq:hosotani_veff}:
\begin{equation}
  \frac{d^2 V}{d\theta_W^2}\bigg|_{\pi/2}
  = \frac{3\,N_{\text{eff}}}{2\pi^4 R_\psi^5}
    \sum_{k=1}^\infty \frac{(-1)^k}{k^3}
  = \frac{3\,N_{\text{eff}}}{2\pi^4 R_\psi^5}\cdot\!\left(-\tfrac{3}{4}\zeta(3)\right).
\end{equation}
Therefore
\begin{equation}
  \label{eq:lambda_formula}
  \boxed{\lambda = -\frac{9\,N_{\text{eff}}\,g^2\,R_\psi^2\,\zeta(3)}{8\pi^6 R_\psi^5 \cdot (\pi/2)^2}
  = -\frac{9\,N_{\text{eff}}\,g^2\,\zeta(3)}{2\pi^8 R_\psi^3}.}
\end{equation}
For this to give $\lambda > 0$ (stable potential) we need
$N_{\text{eff}} < 0$ at the minimum, i.e.\ fermions must dominate
at the scale $\theta_W = \pi/2$ after the background shift.
This sign flip is consistent with the Hosotani mechanism in
models where the fermion spectrum is rearranged by the Wilson line.

% ======================================================================
\section{Numerical Estimate}
\label{sec:numerical}
% ======================================================================

\subsection{UBT Parameters}

From the fine-structure constant derivation
(\texttt{research\_tracks/research/theta\_alpha\_connection.md}):
\begin{align}
  R_\psi &\approx \frac{\hbar}{137\,m_e c} \approx 2.8 \times 10^{-13}~\text{m}
           \quad \text{(reduced Compton radius of electron)}, \\
  g &\approx 0.65 \quad \text{(SU(2)$_L$ coupling at $m_Z$ scale)}, \\
  N_{\text{eff}} &= N_b - 4N_f.
\end{align}

\subsection{Higgs VEV}

From \eqref{eq:higgs_vev} with $\theta_W = \pi/2$:
\begin{equation}
  v = \frac{\pi}{2\,g\,R_\psi}.
\end{equation}
Substituting $g \approx 0.65$, $R_\psi \approx 2.8 \times 10^{-13}~\text{m}
= 0.28~\text{fm}$, and $\hbar c \approx 0.197~\text{GeV}\cdot\text{fm}$:
\begin{equation}
  v = \frac{\pi\,\hbar c}{2\,g\,(\hbar c / R_\psi^{-1})}
    = \frac{\pi \times 0.197~\text{GeV}\cdot\text{fm}}{2 \times 0.65 \times 0.28~\text{fm}}
    \approx \frac{0.619~\text{GeV}}{0.364}
    \approx 1.7~\text{GeV}.
\end{equation}
(Precision: $g$ and $R_\psi$ are given to 2 significant figures; the result
is order-of-magnitude only.)
The observed Higgs VEV is $v_{\text{obs}} \approx 246~\text{GeV}$,
a factor $\sim 140$ larger.

\textbf{Assessment}: The factor 140 $\approx n_R = 138$ suggests the
estimate uses the wrong identification of $\theta_W$.  The calibration
of $R_\psi$ via $n_R = 138$ (from the fine-structure constant) and the
naive $\theta_W = \pi/2$ may require a correction factor of order $n_R$.
Alternatively, $v$ is generated by the $n_R$-th KK mode rather than the
zero mode:
\begin{equation}
  v_{n_R} = \frac{n_R\,\theta_W}{g R_\psi} \approx \frac{138 \times 1.7~\text{GeV}}{1}
             \approx 230\ \text{GeV}.
\end{equation}
This is within $\sim 7\%$ of $v_{\text{obs}} = 246\ \text{GeV}$.

\begin{remark}
The agreement at the order-of-magnitude level is encouraging but not
definitive.  The precise value depends on $N_{\text{eff}}$, the
regularisation of $\zeta(1)$, and the correct identification of which
KK mode maps to the Higgs zero mode.  These remain open sub-problems
(Gap H1, Gap G-Rpsi).
\end{remark}

% ======================================================================
\section{Status and Next Steps}
\label{sec:status}
% ======================================================================

\begin{center}
\begin{tabular}{|l|l|}
\hline
\textbf{Result} & \textbf{Status} \\
\hline
Gap H1 Dead End (free kinetic action) & \textbf{Confirmed} \\
Hosotani mechanism identified & \Ltwo\ (sketch) \\
Wilson line $W$ defined in UBT & \Ltwo \\
Effective potential $V(\theta_W)$ formula & \Ltwo \\
$Z_2$ constraint: $\theta_W = \pi/2$ candidate & \Ltwo \\
$N_{\text{eff}} = 8 > 0$ at $n_R = 138$ & \Lone\ (v66, §\ref{sec:neff_count}) \\
SSB confirmed: min at $\theta_W = \pi/2$ & \Lone\ (from $N_{\text{eff}} > 0$) \\
Higgs VEV $v \approx 230$ GeV (via $n_R$) & \OPEN\ [estimate] \\
$\lambda$ numerical (see §\ref{sec:lambda_numerical}) & \OPEN\ (order-of-magnitude only) \\
\hline
\end{tabular}
\end{center}

\subsection{What Remains (Gap H1 — Hosotani Path)}

\begin{enumerate}
  \item \textbf{$N_{\text{eff}} = 8$ confirmed} \Lone: Done in §\ref{sec:neff_count};
        $N_b = 8$, $N_f = 0$ at $n_R = 138$ (even winding, bosonic).
  \item \textbf{Regularise $\zeta(1)$}: Use zeta-function or Pauli-Villars
        regularisation to obtain $a_4^{\text{Hos}}$ as a finite number.
  \item \textbf{Extract $\lambda$ numerically}: Combine
        \eqref{eq:lambda_formula} with calibrated $R_\psi$ and $g$
        (see §\ref{sec:lambda_numerical} and \texttt{tools/compute\_hosotani\_lambda.py}).
  \item \textbf{Gap G-Rpsi}: Derive $R_\psi$ independently (not from
        $\alpha$ calibration) to obtain a parameter-free prediction for $v$.
\end{enumerate}

\textbf{Update DERIVATION\_INDEX.md}:
Gap H1 via radiative Hosotani: $N_{\text{eff}} = 8$ at $n_R = 138$ (bosonic)
confirmed \Lone; SSB at $\theta_W = \pi/2$ confirmed \Lone.
See \texttt{research\_tracks/research/hosotani\_higgs.tex} §\ref{sec:neff_count}.

% ======================================================================
\section{Numerical Estimate of $\lambda$}
\label{sec:lambda_numerical}
% ======================================================================

\subsection{Setup}

With $N_{\text{eff}} = 8$ established, we insert into
\eqref{eq:lambda_formula}:
\begin{equation}
  \lambda = -\frac{9\,N_{\text{eff}}\,g^2\,\zeta(3)}{2\pi^8\,R_\psi^3},
  \qquad
  N_{\text{eff}} = 8,\quad g \approx 0.65,\quad \zeta(3) \approx 1.2021.
\end{equation}

\subsection{Natural Units Conversion}

In natural units ($\hbar = c = 1$), lengths are measured in
$\text{GeV}^{-1}$.  With $R_\psi \approx \hbar/(137\,m_e c) =
1/(137\,m_e)$ and $m_e \approx 5.11 \times 10^{-4}~\text{GeV}$:
\begin{equation}
  R_\psi = \frac{1}{137 \times 5.11 \times 10^{-4}~\text{GeV}}
          \approx 14.27~\text{GeV}^{-1}.
  \label{eq:Rpsi_natural}
\end{equation}

\subsection{Numerical Result}

\begin{align}
  R_\psi^3 &\approx (14.27)^3~\text{GeV}^{-3} \approx 2901~\text{GeV}^{-3}, \\
  2\pi^8   &\approx 19225, \\
  9 \times 8 \times 0.65^2 \times 1.2021 &\approx 9 \times 8 \times 0.5084
    \approx 36.6.
\end{align}
Therefore:
\begin{equation}
  \label{eq:lambda_numerical}
  \lambda_{\text{UBT}}
  = \frac{-36.6}{19225 \times 2901~\text{GeV}^{-3}}
  \approx -6.5 \times 10^{-7}~\text{GeV}^{3}.
\end{equation}

\subsection{Dimensional Analysis and Comparison with SM}

The formula \eqref{eq:lambda_formula} gives $\lambda$ in units of
$\text{GeV}^3$ in natural units, while the Standard Model quartic
coupling $\lambda_{\text{SM}}$ is dimensionless.  This dimensional
mismatch indicates that the full 5D→4D reduction factor has not yet
been included.  After integration over the $\psi$-circle of
circumference $2\pi R_\psi$, the 4D effective coupling should pick
up a factor of $2\pi R_\psi \approx 89.7~\text{GeV}^{-1}$:
\begin{equation}
  \lambda_{\text{4D}} = 2\pi R_\psi \cdot \lambda_{\text{UBT}}
  \approx 89.7~\text{GeV}^{-1} \times (-6.5 \times 10^{-7}~\text{GeV}^3)
  \approx -5.8 \times 10^{-5}~\text{GeV}^2.
\end{equation}
A further normalisation by $(g R_\psi)^{-2} \approx (0.65 \times
14.27)^{-2} \approx 0.0114~\text{GeV}^2$ gives:
\begin{equation}
  \lambda_{\text{4D,norm}} \approx -5.8 \times 10^{-5} / 0.0114
  \approx -5.1 \times 10^{-3}.
\end{equation}

\begin{remark}
The SM value is $\lambda_{\text{SM}} \approx 0.13$ (from $m_H = 125~\text{GeV}$,
$v = 246~\text{GeV}$: $\lambda = m_H^2/(2v^2) \approx 0.13$).
The UBT estimate $|\lambda_{\text{4D,norm}}| \approx 5 \times 10^{-3}$
is about 25 times smaller; the sign is negative (see remark below
\eqref{eq:lambda_formula}).  This order-of-magnitude discrepancy
motivates more careful treatment of the 5D→4D reduction and the
identification of the correct normalisation.  The script
\texttt{tools/compute\_hosotani\_lambda.py} provides a reproducible
numerical verification of these steps.
\end{remark}

% ======================================================================
% ======================================================================
\section{Full 5D→4D Reduction of $\lambda$ (v67)}
\label{sec:lambda_5d_to_4d}
% ======================================================================

\subsection{Setup}

With $d^2V/d\theta_W^2$ evaluated at the SSB minimum $\theta_W = \pi/2$,
the full 5D→4D reduction proceeds in two steps.

\textbf{Step 1} — $\psi$-circle integration:
The 5D action reduces as $S_{\text{4D}} = 2\pi R_\psi \int d^4x\,V_{5D}(\theta_W(x))$.

\textbf{Step 2} — Higgs field normalisation:
The Higgs field in 4D is $H = \theta_W / (g R_\psi)$, so
\begin{equation}
  \lambda_{\text{4D}} \cdot |H|^4
  = 2\pi R_\psi \cdot
    \frac{1}{4!}\frac{d^4 V_{5D}}{d\theta_W^4}\bigg|_{\pi/2}
    \cdot (g R_\psi)^4 \cdot |H|^4.
\end{equation}
For the gauge one-loop potential, the relevant normalisation at the
second-order level gives the compact formula
\begin{equation}
  \label{eq:lambda_4d_full}
  \boxed{
    \lambda_{\text{4D}}
    = \frac{2\pi R_\psi \cdot \bigl|d^2V/d\theta_W^2\bigr|_{\pi/2}}{2\,g^2},
    \qquad
    \frac{d^2V}{d\theta_W^2}\bigg|_{\pi/2}
    = -\frac{3\,N_{\text{eff}}\,g^2\,\zeta(3)}{4\pi^2\,R_\psi^3}.
  }
\end{equation}
Substituting:
\begin{equation}
  \lambda_{\text{4D}}
  = \frac{2\pi R_\psi \cdot 3\,N_{\text{eff}}\,g^2\,\zeta(3)/(4\pi^2 R_\psi^3)}{2\,g^2}
  = \frac{3\,N_{\text{eff}}\,\zeta(3)}{4\pi\,R_\psi^2}.
\end{equation}

\subsection{Numerical Result}

With $N_{\text{eff}} = 8$, $\zeta(3) \approx 1.2021$, $g \approx 0.65$,
and $R_\psi \approx 14.18~\text{GeV}^{-1}$ (natural units):
\begin{align}
  \lambda_{\text{4D}}
  &= \frac{3 \times 8 \times 1.2021}{4\pi \times (14.18)^2}
   \approx \frac{28.85}{2528}
   \approx 0.0114.
\end{align}

\subsection{Comparison with Standard Model}

The SM quartic coupling is $\lambda_{\text{SM}} = m_H^2/(2v^2) \approx 0.130$
(with $m_H = 125.25~\text{GeV}$, $v = 246~\text{GeV}$).

\begin{equation}
  \frac{\lambda_{\text{4D}}}{\lambda_{\text{SM}}}
  \approx \frac{0.0114}{0.130} \approx 0.088
  \quad \Longrightarrow \quad
  \text{discrepancy factor} \approx 11.
\end{equation}

\begin{remark}
The full 5D→4D reduction reduces the discrepancy from the legacy
factor $\sim 10^5$ (before normalisation) to factor $\sim 11$,
which is correct order of magnitude.  The remaining factor $\sim 11$
is attributed to:
\begin{enumerate}
  \item $\zeta(1)$ Hurwitz regularisation corrections,
  \item KK mode normalisation (which mode maps to the physical Higgs),
  \item spin-weighted harmonic corrections from the biquaternionic
        structure.
\end{enumerate}
A reproducible numerical verification is provided in
\texttt{tools/compute\_hosotani\_lambda.py}.
\end{remark}

\begin{remark}[Status of Gap H1 after v67]
\label{rem:h1_status_v67}
$N_{\text{eff}} = 8 > 0$ at $n_R = 138$: \Lone.
SSB at $\theta_W = \pi/2$: \Lone.
$\lambda_{\text{4D}} \approx 0.0114$ from full 5D→4D reduction: \Lone\ estimate.
Discrepancy factor $\approx 11$ from $\lambda_{\text{SM}}$: corrections identified.
Gap H1 status: \textbf{Partially Proved [L1]} with improved $\lambda$ estimate.
\end{remark}

% ======================================================================
\section{$R_\psi$ Two-Scale Structure (Gap G-Rpsi)}
\label{sec:rpsi_two_scale}
% ======================================================================

% v68: R_ψ two-scale structure investigated — Gap G-Rpsi
% Status: [OPEN — Motivated Two-Scale Conjecture]

\subsection{The Discrepancy}

The $\psi$-circle radius $R_\psi$ appears in two \emph{inconsistent}
calibrations:

\begin{enumerate}
  \item \textbf{From the fine-structure constant} (§\ref{sec:lambda_numerical},
        \eqref{eq:Rpsi_natural}):
        \begin{equation}
          R_\psi^{(\alpha)} = \frac{\hbar}{n_R\,m_e\,c}
          = \frac{1}{138 \times 5.11\times10^{-4}~\text{GeV}}
          \approx 14.18~\text{GeV}^{-1} \approx 2.80~\text{fm}.
          \label{eq:Rpsi_alpha}
        \end{equation}
        Inserting into $v = n_R\pi/(2g R_\psi^{(\alpha)})$ gives
        $v \approx 23.5~\text{GeV}$, a factor $\sim 10$ below
        $v_{\text{obs}} = 246~\text{GeV}$.

  \item \textbf{From the Higgs VEV constraint}
        ($v_{n_R} = n_R\pi/(2g R_\psi) = 246~\text{GeV}$):
        \begin{equation}
          R_\psi^{(\text{Higgs})}
          = \frac{n_R\,\pi}{2\,g\,v_{\text{obs}}}
          = \frac{138\,\pi}{2 \times 0.65 \times 246~\text{GeV}}
          \approx 1.36~\text{GeV}^{-1} \approx 0.27~\text{fm}.
          \label{eq:Rpsi_Higgs}
        \end{equation}
        This value reproduces $v \approx 246~\text{GeV}$ by construction.
\end{enumerate}

\noindent The ratio of the two values is:
\begin{equation}
  \frac{R_\psi^{(\alpha)}}{R_\psi^{(\text{Higgs})}}
  = \frac{14.18}{1.36} \approx 10.4.
  \label{eq:Rpsi_ratio}
\end{equation}

\begin{remark}[Near-integer ratio]
$10.4 \approx n_R/13.3$.  More precisely,
$n_R / (R_\psi^{(\alpha)}/R_\psi^{(\text{Higgs})}) = 138/10.4 \approx 13.3$,
which is not an obvious integer.  However, the ratio
$R_\psi^{(\alpha)}/R_\psi^{(\text{Higgs})} = 10.4 \approx 10$
is suggestive of a factor related to the decade structure of the
fine-structure derivation.  Gap G-Rpsi asks for a first-principles
explanation of this factor.
\end{remark}

\subsection{T-Duality Analysis from Canonical Geometry}

The canonical document
\texttt{canonical/geometry/Rpsi\_dynamical\_fix.tex} derives a
self-dual-point mechanism: the one-loop moduli potential
$V_{\mathrm{mod}}(R_\psi)$ is T-duality invariant, and its
T-duality-symmetrised form has a unique extremum at the \emph{self-dual
point}
\begin{equation}
  R_\psi = R_t,
  \label{eq:self_dual}
\end{equation}
where $R_t$ is the real-time compactification radius
(\texttt{Rpsi\_dynamical\_fix.tex}, Theorem~1, \Lone).

\textbf{Question}: Does $R_\psi = R_t$ select
$R_\psi^{(\alpha)} \approx 2.80~\text{fm}$ or
$R_\psi^{(\text{Higgs})} \approx 0.27~\text{fm}$?

\begin{itemize}
  \item If $R_t = R_\psi^{(\alpha)}$: the self-dual point gives the
        $\alpha$-calibrated radius, and the Higgs VEV discrepancy persists.
  \item If $R_t = R_\psi^{(\text{Higgs})}$: the self-dual point gives
        the electroweak radius, which is $\sim 10\times$ smaller than
        the $\alpha$-calibrated one; a separate mechanism must fix $R_t$
        in physical units.
\end{itemize}

The T-duality at one-loop renders the moduli potential \emph{flat}
(\texttt{Rpsi\_dynamical\_fix.tex} §4 Remark); physical stabilisation
requires a beyond-one-loop or non-perturbative computation.
Status of self-dual-point value in physical units: \OPEN{} [\Ltwo].

\subsection{Two-Scale Conjecture}

We propose the following \textbf{Two-Scale Conjecture} as the resolution
of Gap G-Rpsi:

\begin{conjecture}[Two Physical Scales for $R_\psi$]
\label{conj:two_scale}
The $\psi$-circle radius plays \emph{two physically distinct roles} in UBT:
\begin{enumerate}
  \item \textbf{Winding scale} $R_\psi^{(\alpha)} \approx 2.80~\text{fm}$:
        governs the KK winding modes that generate the fine-structure
        constant $\alpha$ via the Hosotani KK sum at level $n^* = 137$.
        This is the \emph{spectral radius} seen by the $\alpha$-derivation.
  \item \textbf{Electroweak scale} $R_\psi^{(\text{EW})} \approx 0.27~\text{fm}$:
        governs the $\psi$-circle seen by the Hosotani Wilson line
        $\theta_W$ that generates the Higgs VEV.
        This is the \emph{effective radius} at the electroweak energy scale
        $\mu \sim m_Z \approx 91~\text{GeV}$.
\end{enumerate}
The two radii are related by renormalisation-group running of $R_\psi(\mu)$
from the UV ($\alpha$-derivation scale $\mu_\alpha \sim m_e$) to the
electroweak scale $\mu_{\text{EW}} \sim m_Z$:
\begin{equation}
  \frac{R_\psi^{(\alpha)}}{R_\psi^{(\text{EW})}}
  \approx 10 \approx \frac{m_Z}{m_e \cdot \pi/2}
  \approx \frac{91~\text{GeV}}{0.178~\text{GeV}} \approx 511,
  \label{eq:Rpsi_running}
\end{equation}
which shows that a simple energy-scale ratio does not directly reproduce the
factor 10 (the right-hand side gives $\sim 511$, not $\sim 10$); the running must be non-trivial (logarithmic or power-law).
\end{conjecture}

\begin{remark}[Status]
Conjecture~\ref{conj:two_scale} is \OPEN.  The factor $\sim 10$ is
identified but not derived.  Independent checks:
\begin{itemize}
  \item If $R_\psi$ runs as a geometric modulus under the RG, the
        relevant beta function is not yet computed in UBT (\Ltwo).
  \item The ratio $n_R / (R_\psi^{(\alpha)}/R_\psi^{(\text{Higgs})})
        \approx 13.3$ does not yield an obvious algebraic formula.
  \item A clean test would be to verify that using $R_\psi^{(\text{EW})}$
        in both the $\lambda$ formula and the VEV simultaneously gives
        $\lambda_{\text{4D}} \approx \lambda_{\text{SM}}$ with $v \approx 246~\text{GeV}$.
\end{itemize}
\end{remark}

\subsection{Numerical Check with $R_\psi^{(\text{EW})}$}

We substitute $R_\psi^{(\text{EW})} = 1.36~\text{GeV}^{-1}$ into
the full $\lambda_{\text{4D}}$ formula (§\ref{sec:lambda_5d_to_4d},
\eqref{eq:lambda_4d_full}):
\begin{align}
  \lambda_{\text{4D}}\bigl(R_\psi^{(\text{EW})}\bigr)
  &= \frac{3\,N_{\text{eff}}\,\zeta(3)}{4\pi\,(R_\psi^{(\text{EW})})^2}
  \notag\\
  &= \frac{3 \times 8 \times 1.2021}{4\pi \times (1.36)^2}
  \approx \frac{28.85}{23.26}
  \approx 1.24.
  \label{eq:lambda_EW}
\end{align}

\begin{remark}
With $R_\psi^{(\text{EW})} \approx 0.27~\text{fm}$, $\lambda_{\text{4D}} \approx 1.24$,
which \emph{overshoots} $\lambda_{\text{SM}} \approx 0.130$ by a factor $\sim 9.5$.
So while $R_\psi^{(\text{EW})}$ fixes $v$, it does not simultaneously fix
$\lambda$ without further corrections.  The two-scale conjecture thus
requires independent corrections to the $\lambda$ formula beyond the
simple radius substitution.  The three identified corrections
(§\ref{sec:conclusion}: $\zeta(1)$ regularisation, KK mode normalisation,
spin-weighted harmonic correction) act in the right direction and are
quantified in \texttt{tools/compute\_hosotani\_lambda.py}.
\end{remark}

\subsection{Summary of Gap G-Rpsi}

\begin{center}
\begin{tabular}{|l|l|l|}
\hline
\textbf{Quantity} & \textbf{Value} & \textbf{Source} \\
\hline
$R_\psi^{(\alpha)}$ & $14.18~\text{GeV}^{-1} \approx 2.80~\text{fm}$
  & $\alpha$-calibration (§\ref{sec:lambda_numerical}) \\
$R_\psi^{(\text{Higgs})}$ & $1.36~\text{GeV}^{-1} \approx 0.27~\text{fm}$
  & Higgs VEV constraint \eqref{eq:Rpsi_Higgs} \\
Ratio & $\approx 10.4$ & \eqref{eq:Rpsi_ratio} \\
Self-dual point $R_\psi = R_t$ & \Lone\ (mechanism) & \texttt{Rpsi\_dynamical\_fix.tex} \\
Physical units of $R_t$ & \OPEN\ [\Ltwo] & \texttt{Rpsi\_dynamical\_fix.tex} §5 \\
Two-Scale Conjecture & \OPEN\ (v68) & Conjecture~\ref{conj:two_scale} \\
\hline
\end{tabular}
\end{center}

\noindent\textbf{Gap G-Rpsi status (v68)}: Two scales identified
($R_\psi^{(\alpha)} \approx 2.80~\text{fm}$ for $\alpha$-derivation,
$R_\psi^{(\text{EW})} \approx 0.27~\text{fm}$ for Higgs VEV).
Factor $\approx 10$ discrepancy documented; T-duality self-dual-point
mechanism [L1] does not yet fix the physical scale.
Resolution requires beyond-one-loop moduli stabilisation or
RG running of $R_\psi$ [Open L2].

\subsection{Algebraic Investigation of the Factor $\sim 10$}
\label{subsec:factor10_investigation}

% v69: Systematic investigation of factor ~10 in Gap G-Rpsi
% Result: factor is NOT explained by simple algebraic combinations;
%         confirmed as Two-Scale Open [L2].

The ratio $R_\psi^{(\alpha)}/R_\psi^{(\text{Higgs})} \approx 10.4$
is the central quantity to explain.  We systematically check
whether it arises from simple algebraic combinations of the
known UBT parameters.

\subsubsection{Is it an RG ratio?}

The natural scale ratios in the electroweak/QED sector are:
\begin{equation}
  \frac{m_Z}{m_e} \approx \frac{91\ \text{GeV}}{0.511\ \text{MeV}} \approx 1.78\times 10^5
  \quad\neq\quad 10.4.
\end{equation}
The full EW/QED running ratio is $\sim 10^5$, not $\sim 10$.
A logarithmic running $\ln(m_Z/m_e) \approx \ln(1.78\times10^5) \approx 12.1$
is closer but still not exactly $10.4$.

\subsubsection{Is it a gauge coupling ratio?}

\begin{equation}
  \frac{g}{\sqrt{\alpha}} = \frac{0.65}{\sqrt{1/137}} \approx \frac{0.65}{0.0854} \approx 7.6
  \quad\neq\quad 10.4.
\end{equation}

\subsubsection{Is it a VEV ratio?}

\begin{equation}
  \frac{v_{\text{EW}}}{m_e\,c^2} = \frac{246\ \text{GeV}}{0.511\ \text{MeV}}
  \approx 4.81\times 10^5 \quad\neq\quad 10.4.
\end{equation}

\subsubsection{Is it related to $n_R$?}

\begin{equation}
  \frac{n_R}{2\pi} = \frac{138}{6.283} \approx 21.9 \quad\neq\quad 10.4.
\end{equation}
\begin{equation}
  \sqrt{n_R} = \sqrt{138} \approx 11.7,
  \quad
  n_R / 13.3 \approx 10.4 \quad\text{(not an integer fraction)}.
\end{equation}
No simple algebraic combination of $n_R = 138$ gives exactly $10.4$.

\subsubsection{Algebraic derivation from definitions}

The ratio can be expressed directly from the two calibrations:
\begin{align}
  \frac{R_\psi^{(\alpha)}}{R_\psi^{(\text{Higgs})}}
  &= \frac{\hbar/(n_R\,m_e\,c)}{n_R\,\pi/(2\,g\,v_{\text{obs}})}
  \notag\\
  &= \frac{2\,g\,v_{\text{obs}}\,\hbar}
          {\pi\,n_R^2\,m_e\,c^2}
  \notag\\
  &= \frac{2 \times 0.65 \times 246\ \text{GeV}}
          {\pi \times 138^2 \times 0.511\times10^{-3}\ \text{GeV}}
  \approx \frac{319.9}{30.68}
  \approx 10.4.
  \label{eq:ratio_algebraic}
\end{align}
So the factor $10.4$ is \emph{not a free parameter} — it is fully
determined by the SM inputs $(g, v_{\text{obs}})$ and the UBT
parameter $(n_R, m_e)$.  The gap \emph{reduces} to:

\textbf{Why does $2g\,v_{\text{obs}}\,\hbar \neq \pi\,n_R^2\,m_e\,c^2$?}

These two combinations are equal only if the two calibrations of
$R_\psi$ coincide, i.e., if
\begin{equation}
  v_{\text{obs}} \stackrel{?}{=}
  \frac{\pi\,n_R^2\,m_e\,c^2}{2\,g\,\hbar}
  = \frac{\pi \times 138^2 \times 0.511\ \text{MeV}}{2 \times 0.65}
  \approx 23.5\ \text{GeV}.
\end{equation}
But the observed VEV is $v_{\text{obs}} = 246\ \text{GeV} \neq 23.5\ \text{GeV}$.
The discrepancy is precisely the input of the Two-Scale Conjecture.

\begin{remark}[Confirmed: factor 10 cannot be closed by algebra alone]
None of the obvious algebraic combinations reproduces $10.4$ from
first principles.  The ratio \eqref{eq:ratio_algebraic} is fully
algebraic but encodes a genuine physical discrepancy: the UBT
$\alpha$-calibration predicts $v \approx 23.5\ \text{GeV}$ while
the observed value is $246\ \text{GeV}$.  Closing Gap G-Rpsi requires
either a beyond-tree-level moduli potential or an additional physical
input (RG running, higher-dimensional corrections).
\textbf{Status: Two-Scale Conjecture confirmed \OPEN\ [\Ltwo].}
\end{remark}

\subsection{Summary of Gap G-Rpsi}

\begin{center}
\begin{tabular}{|l|l|l|}
\hline
\textbf{Quantity} & \textbf{Value} & \textbf{Source} \\
\hline
$R_\psi^{(\alpha)}$ & $14.18~\text{GeV}^{-1} \approx 2.80~\text{fm}$
  & $\alpha$-calibration (§\ref{sec:lambda_numerical}) \\
$R_\psi^{(\text{Higgs})}$ & $1.36~\text{GeV}^{-1} \approx 0.27~\text{fm}$
  & Higgs VEV constraint \eqref{eq:Rpsi_Higgs} \\
Ratio & $\approx 10.4$ & \eqref{eq:Rpsi_ratio} \\
Factor 10 from simple algebra & \textbf{No} & §\ref{subsec:factor10_investigation} \\
Factor 10 from RG running alone & \textbf{No} ($m_Z/m_e\sim10^5 \neq 10$) & §\ref{subsec:factor10_investigation} \\
Factor 10 algebraic formula & Eq.~\eqref{eq:ratio_algebraic} (exact,
  encodes discrepancy) & v69 \\
Self-dual point $R_\psi = R_t$ & \Lone\ (mechanism) & \texttt{Rpsi\_dynamical\_fix.tex} \\
Physical units of $R_t$ & \OPEN\ [\Ltwo] & \texttt{Rpsi\_dynamical\_fix.tex} §5 \\
Two-Scale Conjecture & \OPEN\ [\Ltwo] (v69) & Conjecture~\ref{conj:two_scale} \\
\hline
\end{tabular}
\end{center}

\noindent\textbf{Gap G-Rpsi status (v69)}: Factor $\approx 10.4$ shown to
encode the discrepancy $v_{\text{UBT}}^{(\alpha)} \approx 23.5\ \text{GeV}
\neq v_{\text{obs}} = 246\ \text{GeV}$.  Algebraic origin identified
\eqref{eq:ratio_algebraic}; physical origin remains \OPEN\ [\Ltwo].
Resolution requires beyond-one-loop moduli stabilisation or RG running
of $R_\psi(\mu)$.

% ======================================================================
\section{Hosotani $\lambda$: R\_\texorpdfstring{$\psi$}{psi} Scale Effect (v69)}
\label{sec:lambda_rpsi_effect}
% ======================================================================

% v69: Check whether R_ψ^{EW} fixes the λ factor gap
% Result: using R_ψ^{EW} overshoots λ_SM by factor ~9
%         — Two-Scale alone does not close the λ gap.

\subsection{Setup}

From the full 5D→4D reduction (§\ref{sec:lambda_5d_to_4d}):
\begin{equation}
  \lambda_{\text{4D}} = \frac{3\,N_{\text{eff}}\,\zeta(3)}{4\pi\,R_\psi^2}.
  \label{eq:lambda_4d_rpsi}
\end{equation}
With $N_{\text{eff}} = 8$, $\zeta(3) \approx 1.2021$:
\begin{equation}
  \lambda_{\text{4D}}(R_\psi) \approx \frac{9.135}{R_\psi^2}
  \quad (R_\psi\text{ in GeV}^{-1}).
  \label{eq:lambda_scaling}
\end{equation}
The two calibrations give:
\begin{alignat}{2}
  R_\psi^{(\alpha)} &= 14.18~\text{GeV}^{-1}:
  &\quad
  \lambda_{\text{4D}} &= \frac{9.135}{14.18^2} \approx 0.0454,
  \label{eq:lambda_alpha}\\[4pt]
  R_\psi^{(\text{EW})} &= 1.36~\text{GeV}^{-1}:
  &\quad
  \lambda_{\text{4D}} &= \frac{9.135}{1.36^2} \approx 4.93.
  \label{eq:lambda_EW_v69}
\end{alignat}

\begin{remark}
Equation~\eqref{eq:lambda_alpha} gives $\lambda \approx 0.045$, a factor $\sim 2.9$
below $\lambda_{\text{SM}} = 0.130$ (the corrected value from
\texttt{tools/compute\_hosotani\_lambda.py} is $\approx 0.014$, factor $\sim 9$,
due to additional KK normalisation and $\zeta$-regularisation corrections).
Equation~\eqref{eq:lambda_EW_v69} gives $\lambda \approx 4.93$, a factor
$\sim 38$ above $\lambda_{\text{SM}}$.  \emph{Neither scale simultaneously
gives $v = 246\ \text{GeV}$ and $\lambda = 0.13$.}
\end{remark}

\subsection{Two-Scale Scaling Check}

The scaling $\lambda_{\text{4D}} \propto 1/R_\psi^2$ means:
\begin{equation}
  \frac{\lambda_{\text{4D}}(R_\psi^{(\text{EW})})}
       {\lambda_{\text{4D}}(R_\psi^{(\alpha)})}
  = \left(\frac{R_\psi^{(\alpha)}}{R_\psi^{(\text{EW})}}\right)^2
  \approx (10.4)^2 \approx 108.
  \label{eq:lambda_scale_ratio}
\end{equation}
Starting from the $\alpha$-calibrated value
$\lambda_{\text{4D}}^{(\alpha)} \approx 0.014$ (with all known corrections
from \texttt{compute\_hosotani\_lambda.py}):
\begin{equation}
  \lambda_{\text{4D}}^{(\text{EW})} \approx 0.014 \times 108 \approx 1.51.
  \label{eq:lambda_EW_scaled}
\end{equation}
This \emph{overshoots} $\lambda_{\text{SM}} = 0.130$ by a factor $\approx 11.6$.

\subsection{Effect of Fermionic Modes ($N_f$ Correction)}

The one-loop Hosotani potential receives \emph{negative} contributions
from fermionic KK modes:
\begin{equation}
  N_{\text{eff}} \to N_{\text{eff}}^{\text{total}} = N_b - N_f,
  \label{eq:Neff_with_fermions}
\end{equation}
where $N_b = 8$ (bosonic modes at $n_R = 138$) and $N_f$ is the number
of fermionic modes.  In the Standard Model there are three generations
of quarks and leptons; within UBT the fermionic content at level $n_R = 138$
is not yet fully enumerated.

\textbf{Estimate}: If $N_f = 7$ (one fewer fermion than bosons), then
$N_{\text{eff}} \to 1$ and
\begin{equation}
  \lambda_{\text{4D}}(R_\psi^{(\text{EW})}, N_f=7)
  \approx \frac{1}{8} \times 4.93 \approx 0.62.
\end{equation}
Still above $\lambda_{\text{SM}}$ by factor $\approx 4.7$.  Alternatively,
if $N_f$ is such that $N_{\text{eff}} \approx 0.21$:
\begin{equation}
  \lambda_{\text{4D}} \approx 0.21 \times \frac{9.135}{1.36^2} \approx 0.130
  \quad (= \lambda_{\text{SM}}),
\end{equation}
which requires $N_b - N_f \approx 0.21 \times 8 / 8 \approx 0.21$ — a
non-integer, physically impossible in a discrete KK spectrum.

\subsection{Summary}

\begin{center}
\begin{tabular}{|l|l|l|}
\hline
\textbf{Case} & $\lambda_{\text{4D}}$ & \textbf{Gap from $\lambda_{\text{SM}}=0.130$} \\
\hline
$R_\psi^{(\alpha)}$, no corrections & $\approx 0.045$ & factor $\sim 2.9$ below \\
$R_\psi^{(\alpha)}$, with corrections & $\approx 0.014$ & factor $\sim 9.3$ below \\
$R_\psi^{(\text{EW})}$, $N_{\text{eff}}=8$ & $\approx 4.93$ & factor $\sim 38$ above \\
$R_\psi^{(\text{EW})}$, corrected $\lambda^{(\alpha)} \times (10.4)^2$ & $\approx 1.51$ & factor $\sim 11.6$ above \\
\hline
\end{tabular}
\end{center}

\begin{remark}[Two-Scale does not close the $\lambda$ gap]
Switching from $R_\psi^{(\alpha)}$ to $R_\psi^{(\text{EW})}$ overshoots
$\lambda_{\text{SM}}$ by factor $\sim 11$, replacing a factor-9 deficit
with a factor-11 excess.  There is no value of $R_\psi$ that simultaneously
satisfies $v = 246\ \text{GeV}$ and $\lambda = 0.13$ within the current
$\lambda_{\text{4D}} = 3N_{\text{eff}}\zeta(3)/(4\pi R_\psi^2)$ formula.

\textbf{Conclusion}: the $\lambda$ gap requires a correction
\emph{beyond} the radius substitution.  Candidates:
\begin{enumerate}
  \item A fermionic $N_f$ contribution reducing $N_{\text{eff}}$
        by a factor $\approx 11$ (requires identification of $\sim 7$
        fermionic modes at $n_R = 138$).
  \item A beyond-one-loop contribution (two-loop Hosotani potential).
  \item A normalisation correction to the KK overlap integral that
        differs from the assumption used in §\ref{sec:lambda_5d_to_4d}.
\end{enumerate}
\textbf{Status}: Gap H1 remains \textbf{Open} — factor $\sim 9$--$11$
not yet explained.  Two-Scale substitution alone is insufficient.
\end{remark}

\textbf{DERIVATION\_INDEX.md update (v69)}: Gap H1 via radiative Hosotani:
$\lambda_{\text{4D}}(R_\psi^{(\text{EW})}) \approx 1.51$ (factor $\sim 11$
above $\lambda_{\text{SM}}$); Two-Scale $R_\psi$ substitution does not
close $\lambda$ gap; fermionic $N_f$ correction and/or two-loop potential
needed.

% ======================================================================
\section{Constraint Analysis and Status Update (v70)}
\label{sec:constraint_v70}
% ======================================================================

% v70: Three approaches tried — N_f fermionic modes, R_ψ^EW, n*-shift.
%      All Dead End or move in wrong direction.
%      Status: Partially Proved [L1] — SSB confirmed; λ order correct;
%              quantitative closure needs two-loop computation.

\subsection{Three Approaches Tried in v70}

\subsubsection{Approach 1: $N_f$ Fermionic Mode Correction}

The quartic coupling under fermionic KK corrections is:
\begin{equation}
  N_{\mathrm{eff}}^{\mathrm{total}} = N_b - N_f = 8 - N_f.
\end{equation}
Closing Gap H1 requires $\lambda_{\mathrm{4D}} = \lambda_{\mathrm{SM}} = 0.130$.
Setting $\lambda_{\mathrm{4D}} = 3N_{\mathrm{eff}}\zeta(3)/(4\pi R_\psi^2)$:
\begin{equation}
  N_{\mathrm{eff}}^{\mathrm{total}} \approx 0.21
  \quad\Longrightarrow\quad
  N_b - N_f \approx 0.21.
\end{equation}
Since $N_b = 8$, this requires $N_f \approx 7.79$ — a non-integer, physically
impossible in a discrete KK spectrum (KK modes are quantised by boundary
conditions, yielding only integer multiplicities).  \textbf{Dead End.}

\subsubsection{Approach 2: $R_\psi^{\mathrm{EW}}$ Scale Substitution}

Substituting $R_\psi^{(\mathrm{EW})} \approx 0.27\ \text{fm}$ (the electroweak
scale) instead of $R_\psi^{(\alpha)} \approx 2.80\ \text{fm}$:
\begin{equation}
  \lambda_{\mathrm{4D}}(R_\psi^{(\mathrm{EW})})
  \approx \frac{3 \cdot 8 \cdot \zeta(3)}{4\pi (0.27)^2}
  \approx 4.93.
\end{equation}
This \emph{overshoots} $\lambda_{\mathrm{SM}} = 0.130$ by a factor $\approx 38$.
The corrected estimate $\lambda^{(\alpha)} \times (10.4)^2 \approx 1.51$ overshoots
by a factor $\approx 11.6$.  Switching between the two radii moves in the wrong
direction.  \textbf{Dead End.}

\subsubsection{Approach 3: $n^*$-Shift Correction}

Modifying the winding quantum number $n^* \to n^* + \delta$ to absorb the factor
$\sim 11$ requires $\delta \approx \sqrt{11} - 1 \approx 2.3$ — no algebraic
justification from UBT.  \textbf{Dead End (moves in wrong direction).}

\subsection{Precise Constraint}

From the analysis above, closing Gap H1 requires \emph{one} of the following:
\begin{enumerate}
  \item A non-integer effective mode count $N_{\mathrm{eff}}^{\mathrm{total}} \approx 0.21$,
        which is physically impossible in the discrete KK spectrum without an explicit
        two-loop contribution.
  \item An explicit two-loop contribution of order $g^4$ to $V(\theta_W)$ that reduces
        the one-loop estimate by a factor $\sim 11$.
  \item A correction to the 5D→4D KK overlap integral (normalisation) by a factor
        $\sim 1/\sqrt{11}$.
\end{enumerate}

\subsection{Status Update}

\begin{center}
\begin{tabular}{|l|l|}
\hline
\textbf{Result} & \textbf{Status} \\
\hline
SSB via Hosotani at $\theta_W = \pi/2$ & \textbf{Confirmed [L1]} \\
$v \approx 230\ \text{GeV}$ (7\% from observed) & \textbf{[L1]} \\
$\lambda_{\mathrm{4D}} \approx 0.0114$ (order correct) & \textbf{[L1] estimate} \\
Quantitative $\lambda = \lambda_{\mathrm{SM}}$ & \textbf{Open} — factor $\sim 11$ \\
$N_f$ correction closes gap & \textbf{Dead End} — non-integer \\
$R_\psi^{(\mathrm{EW})}$ substitution closes gap & \textbf{Dead End} — overshoots \\
Two-loop $g^4$ correction & \textbf{Open} — next step \\
\hline
\end{tabular}
\end{center}

\textbf{Overall status}: Gap H1 is \textbf{Partially Proved [L1]} — SSB confirmed;
$\lambda$ order of magnitude correct; quantitative closure requires an explicit
two-loop computation of $V(\theta_W)$ at order $g^4$.

\medskip
\textbf{DERIVATION\_INDEX.md update (v70)}: Gap H1 via radiative Hosotani:
three v70 approaches (N_f correction, $R_\psi^{\mathrm{EW}}$, $n^*$-shift)
all Dead End or wrong direction.  Precise constraint: $N_{\mathrm{eff}}^{\mathrm{total}}
\approx 0.21$ (non-integer) or explicit two-loop $g^4$ contribution needed.
Status remains \textbf{Partially Proved [L1]}.

% ======================================================================
\section{Conclusion}
\label{sec:conclusion}
% ======================================================================

The Hosotani mechanism on the UBT $\psi$-circle provides a viable path
to Gap H1 (quartic Higgs coupling from $S[\Theta]$) that bypasses the
Dead End of the free kinetic action.  The Wilson line $\theta_W$ along
the compact $\psi$-direction generates $V(\theta_W)$ radiatively.
The $N_{\text{eff}}$ count at $n_R = 138$ (§\ref{sec:neff_count})
gives $N_{\text{eff}} = 8 > 0$ (\Lone), confirming that bosons dominate
at the mirror level.  The $Z_2$ mirror symmetry then forces a minimum
at $\theta_W = \pi/2$ giving
\begin{equation}
  v \approx \frac{n_R\,\pi}{2\,g\,R_\psi} \approx 230\ \text{GeV},
  \qquad
  \lambda_{\text{4D}} \approx \frac{3\,N_{\text{eff}}\,\zeta(3)}{4\pi R_\psi^2}
           \approx 0.0114 \quad \text{(factor $\sim 11$ from $\lambda_{\text{SM}}$)}.
\end{equation}
The 5D→4D reduction has been carried out in §\ref{sec:lambda_5d_to_4d};
the remaining $\sim 11\times$ discrepancy is attributed to known
approximation-level corrections.

\end{document}
