% © 2026 Ing. David Jaroš — CC BY-NC-ND 4.0
%
% This work is licensed under a Creative Commons Attribution-NonCommercial-NoDerivatives
% 4.0 International License (CC BY-NC-ND 4.0).
%
% License History: Earlier drafts (up to v0.3) were released under CC BY 4.0.
% From v0.4 onward, all material is released under CC BY-NC-ND 4.0 to protect
% the integrity of the theoretical work during ongoing academic development.

% =====================================================================
% TASK 8 — R Factor Two-Loop Analysis
% Status: [OPEN HARD PROBLEM] — B3 approach, coefficient c not derived
% Source: DERIVATION_INDEX.md §"Fine Structure Constant" R-factor row
% =====================================================================

\documentclass[11pt,a4paper]{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=2.5cm}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{conjecture}{Conjecture}
\theoremstyle{remark}
\newtheorem{remark}{Remark}

\newcommand{\Lzero}{\textbf{[L0]}}
\newcommand{\Lone}{\textbf{[L1]}}
\newcommand{\OPEN}{\textbf{[OPEN]}}

\title{\textbf{R Factor: Two-Loop Analysis on the $\psi$-Circle}}
\author{Ing. David Jaroš}
\date{March 2026 (v57)}

\begin{document}
\maketitle

\begin{abstract}
The correction factor $R \approx 1.114$ in the UBT fine-structure derivation
($B = B_{\mathrm{base}} \times R$) is an open hard problem.  The best
algebraic candidate is $R = 1 + \alpha(N_{\mathrm{eff}} + \pi + 1/4) \approx 1.1123$
(error $0.15\%$), found numerically (\texttt{tools/explore\_r\_factor.py}).  Neither
the coefficient $c = N_{\mathrm{eff}} + \pi + 1/4 \approx 15.39$ nor the formula
$R = 1 + c\alpha$ has been derived from $\mathcal{S}[\Theta]$.  This document
analyses the two-loop beta-function approach (Approach~B3) and documents the
result.
\end{abstract}

% ======================================================================
\section{Setup: The $R$ Factor Problem}
\label{sec:setup}
% ======================================================================

The UBT prediction for $\alpha^{-1}$ requires:
\begin{equation}
  \alpha = \frac{B_{\mathrm{base}} \times R}{n^{*\,2}},
  \qquad
  B_{\mathrm{base}} = N_{\mathrm{eff}}^{3/2} \approx 41.57,
  \qquad n^* = 137.
  \label{eq:alpha_formula}
\end{equation}
With $\alpha = 1/137.036$:
\begin{equation}
  R = \frac{\alpha \cdot n^{*\,2}}{B_{\mathrm{base}}}
    = \frac{137.036^2}{137^2 \cdot 41.57 \cdot 137.036}
    \approx 1.114.
  \label{eq:R_target}
\end{equation}

\noindent\textbf{Status of $R$:}
\begin{center}
\begin{tabular}{|l|l|}
\hline
$B_{\mathrm{base}} = N_{\mathrm{eff}}^{3/2}$ & Partially Proved [L1] \\
$n^* = 137$ (prime attractor) & Proved [L1] \\
$R \approx 1.114$ & \OPEN\ Hard Problem \\
\hline
\end{tabular}
\end{center}

% ======================================================================
\section{Current State: \texttt{explore\_r\_factor.py}}
\label{sec:current}
% ======================================================================

Running \texttt{tools/explore\_r\_factor.py} gives the closest algebraic
candidates (sorted by error):

\begin{center}
\begin{tabular}{lll}
\hline
\textbf{Candidate} & \textbf{Value} & \textbf{Error} \\
\hline
$1 + \alpha(N_{\mathrm{eff}} + \pi + 1/4)$ & 1.11235 & 0.148\% \\
$1 + \alpha N_{\mathrm{eff}}^{3/2}/e$ & 1.11162 & 0.213\% \\
$1 + \alpha(N_{\mathrm{eff}} + \pi)$ & 1.11052 & 0.312\% \\
$2^{1/6}$ & 1.12246 & 0.760\% \\
\hline
\end{tabular}
\end{center}

Best candidate: $R = 1 + \alpha(N_{\mathrm{eff}} + \pi + 1/4)$.
This is a \textbf{Numerical Observation} only — not derived from $\mathcal{S}[\Theta]$.

% ======================================================================
\section{Two-Loop Beta Function on the $\psi$-Circle (Approach B3)}
\label{sec:twoloop}
% ======================================================================

\subsection{One-Loop Structure}

At one loop, the renormalisation group equation on the $\psi$-circle is:
\begin{equation}
  \mu\frac{d\alpha}{d\mu} = \beta_1\,\alpha^2 + \mathcal{O}(\alpha^3),
  \qquad \beta_1 = -\frac{2N_{\mathrm{eff}}}{3\pi}.
  \label{eq:one_loop_beta}
\end{equation}
The one-loop running gives the standard QED-like correction:
$\alpha(M_Z^2) \approx 1/128$, consistent with the observed running of $\alpha$.

\subsection{Two-Loop Contribution}

At two loops, the beta function is:
\begin{equation}
  \mu\frac{d\alpha}{d\mu}
    = \beta_1\,\alpha^2 + \beta_2\,\alpha^3 + \mathcal{O}(\alpha^4),
  \label{eq:two_loop_beta}
\end{equation}
where $\beta_2$ arises from the two-loop diagram on the $\psi$-circle.

The two-loop correction to the effective coupling is:
\begin{equation}
  \alpha_{\mathrm{eff}}
    = \alpha_0\bigl(1 + c\,\alpha_0\bigr) + \mathcal{O}(\alpha_0^3),
  \label{eq:alpha_two_loop}
\end{equation}
where $c$ is the two-loop coefficient.

\subsection{Identifying $c$}

The best numerical candidate gives $c \approx N_{\mathrm{eff}} + \pi + 1/4 = 15.39$.

Decomposing this:
\begin{itemize}
  \item $N_{\mathrm{eff}} = 12$: counting effective bosonic modes on $\psi$-circle.
  \item $\pi$: candidate origin — volume factor from angular integration over
    the $\psi$-circle half-period ($\mathrm{vol}(S^1)/(2\cdot 1) = \pi$).
  \item $1/4$: candidate origin — Weyl spinor loop factor
    ($2$ real DOF / $4D$ spacetime = $1/2$ per DOF; $1/2 \times 1/2 = 1/4$
    for two fermionic lines in the two-loop diagram).
\end{itemize}

\begin{conjecture}[Two-loop coefficient, B3]
\label{conj:two_loop}
The two-loop coefficient on the $\psi$-circle is:
\begin{equation}
  c = N_{\mathrm{eff}} + \pi + \frac{1}{4},
  \label{eq:c_conjecture}
\end{equation}
where $\pi$ comes from half-period integration and $1/4$ from the Weyl
spinor loop factor.  This gives $R = 1 + c\alpha \approx 1.1123$.
\end{conjecture}

\textbf{Status of Conjecture~\ref{conj:two_loop}:}
\textbf{Numerical Observation only} — neither the $\pi$ nor the $1/4$
contribution has been derived from an explicit two-loop Feynman diagram
computation in $\mathcal{S}[\Theta]$ on the $\psi$-circle.

% ======================================================================
\section{Two-Scale Approach (v70): Dead End}
\label{sec:two_scale}
% ======================================================================

\subsection{Setup}

The two $\psi$-circle radii established in UBT are:
\begin{equation}
  R_\psi^{(\alpha)} = \frac{\hbar}{n_R\,m_e\,c} \approx 2.80\ \text{fm},
  \qquad
  R_\psi^{(\mathrm{EW})} = \frac{n_R\,\pi}{2g\,v_{\mathrm{obs}}} \approx 0.27\ \text{fm}.
  \label{eq:two_scale_radii}
\end{equation}
Their ratio is $R_\psi^{(\alpha)}/R_\psi^{(\mathrm{EW})} \approx 10.4$.  The
question: can this ratio generate $R \approx 1.114$?

\subsection{Key Computation}

Setting $\rho = R_\psi^{(\mathrm{EW})}/R_\psi^{(\alpha)} \approx 0.096$, one
observes the numerical coincidence:
\begin{equation}
  \rho \approx 0.096 \approx \alpha \cdot N_{\mathrm{eff}}
  \quad\text{(numerical observation only).}
\end{equation}
Attempting $R = 1 + \rho \cdot \alpha \cdot \beta$ requires
$\beta \approx 163$ — no algebraic value identified.
Alternatively, $R = \rho^\varepsilon$ gives $\varepsilon \approx 0.046$
— again no algebraic value.

\subsection{Verdict}

\begin{center}
\textbf{Dead End} — the ratio $R_\psi^{(\alpha)}/R_\psi^{(\mathrm{EW})}$ does
not generate $R \approx 1.114$ through any natural algebraic combination.
\end{center}

% ======================================================================
\section{Modular Correction Approach (v70): Dead End}
\label{sec:modular}
% ======================================================================

\subsection{Setup}

Near the self-dual point $\tau = i$ of the $\psi$-circle modular parameter,
the partition function $Z(\tau)$ receives corrections from the Eisenstein series
$E_2(\tau)$.  Setting $\tau = i + \varepsilon$ with $\varepsilon \ll 1$:
\begin{equation}
  Z(\tau) \approx Z(i)\!\left[1 + \frac{3\pi\,\varepsilon}{2} + \mathcal{O}(\varepsilon^2)\right].
\end{equation}

\subsection{Key Computation}

With $\varepsilon = 1/(2\pi \cdot 137)$ (one-loop shift at the fine-structure
scale):
\begin{equation}
  \text{correction} \approx 1 + \frac{3\pi}{2}\cdot\frac{1}{2\pi\cdot 137}
  \approx 1.0055.
\end{equation}
This is a factor $\approx 20$ too small compared to $R - 1 \approx 0.114$.
No value of $\varepsilon$ derived from UBT parameters reproduces $R$.

\subsection{Verdict}

\begin{center}
\textbf{Dead End} — modular correction near $\tau = i$ is of order $\alpha \sim 1/137$,
yielding only $R \approx 1.005$, not $1.114$.
\end{center}

% ======================================================================
\section{Heat Kernel on Flat Torus (v70): Dead End}
\label{sec:heat_kernel}
% ======================================================================

\subsection{Setup}

The Seeley--DeWitt heat-kernel expansion on the $\psi$-circle (topologically
$T^1$ or more generally $T^3$ for the three imaginary directions) gives
subleading corrections to the effective action as a series in the background
curvature:
\begin{equation}
  \mathrm{Tr}\,e^{-t\Delta} \sim \frac{1}{(4\pi t)^{d/2}}
  \sum_{n=0}^\infty t^n\,c_n[g],
\end{equation}
where $c_n[g]$ are the Seeley--DeWitt coefficients.

\subsection{Key Computation}

On the flat torus $T^3$ (constant metric, zero curvature):
\begin{equation}
  R_{\mu\nu\rho\sigma} = 0
  \quad\Longrightarrow\quad
  c_n[g] = 0 \;\text{ for all } n \geq 1.
\end{equation}
Therefore all Seeley--DeWitt subleading corrections vanish identically on the
flat $\psi$-torus.  No correction of the form $R \approx 1 + c\alpha$ can arise
from the heat kernel on a flat background.

\subsection{Verdict}

\begin{center}
\textbf{Dead End} — flat $T^3$ has $c_n = 0$; no heat-kernel correction
generates $R \approx 1.114$.
\end{center}

% ======================================================================
\section{Summary: R Remains Open Hard Problem (v70)}
\label{sec:v70_summary}
% ======================================================================

After 25+ approaches across all versions, the $R$ factor remains an
\textbf{Open Hard Problem}.

\medskip
\begin{center}
\begin{tabular}{|l|l|l|}
\hline
\textbf{Approach (v70)} & \textbf{Result} & \textbf{Status} \\
\hline
Two-scale ratio $R_\psi^{(\alpha)}/R_\psi^{(\mathrm{EW})}$
  & $\beta \approx 163$ (no algebraic value) & Dead End \\
Modular correction near $\tau=i$
  & correction $\approx 1.005$ (factor 20 too small) & Dead End \\
Heat kernel on flat $T^3$
  & $c_n = 0$ (zero curvature) & Dead End \\
\hline
\end{tabular}
\end{center}

\medskip
\textbf{Best candidate (unchanged)}: $R = 1 + \alpha(N_{\mathrm{eff}} + \pi + 1/4)
\approx 1.1123$, error $0.15\%$.
\begin{itemize}
  \item $\pi$: candidate origin — $S^1$ angular integration.
  \item $1/4$: candidate origin — Weyl spinor (2 DOF / $2\pi^2$ from 4D).
  \item Neither has been derived from $\mathcal{S}[\Theta]$.
\end{itemize}

\textbf{Status}: $R$ is a \textbf{Motivated Conjecture (B3)} — numerical
observation only.  The two-loop beta-function approach has the correct
qualitative structure $R = 1 + c\alpha$; the coefficient $c$ is not
algebraically derived.

\textbf{Next step}: Explicit two-loop $\beta$-function coefficient on
$T^3 \times S^1$ (direct Feynman diagram computation in $\mathcal{S}[\Theta]$).

\medskip
\textbf{DERIVATION\_INDEX.md update (v70)}: R factor row — three additional
approaches (Two-Scale, Modular, Heat Kernel) all Dead End; best candidate
unchanged; status remains \textbf{Open Hard Problem}.

% ======================================================================
\section{Explicit Two-Loop $\beta$-Function Coefficient on $T^3 \times S^1$ (v71)}
\label{sec:twoloop_explicit}
% ======================================================================

% v71: Explicit computation of two-loop beta-function coefficient c.
% Approach: scalar field, SU(2), SU(2)×U(1) with N_f fermions.
% Result: no natural particle content reproduces c ≈ 15.39 — Dead End.

\subsection{Setup: Two-Loop Vacuum Diagram on $T^3$}

The UBT action on the $\psi$-circle $S^1$ at Kaluza-Klein level $n_R = 138$
is effectively a field theory on $T^3 \times S^1$ (three imaginary directions
times one compact real direction).  The two-loop beta function is controlled
by the two-loop vacuum bubble (sunset diagram):
\begin{equation}
The two-loop vacuum bubble (sunset diagram) contributes as:
\begin{equation}
  \Gamma^{(2)}\big|_{\text{2-loop}} \sim g^4 \!\!\int\!\! \frac{d^dp\,d^dk}{(2\pi)^{2d}}
  \frac{1}{p^2\,(k^2)\,((p+k)^2)},
  \label{eq:twoloop_integral}
\end{equation}
where $d = 3$ (the three compact directions of $\mathrm{Im}(\mathbb{H})$)
and the momenta are Kaluza-Klein sums $p_n = 2\pi n/R_\psi$.

The two-loop $\beta$-function coefficient $c_{2\text{loop}}$ is defined by:
\begin{equation}
  \beta_2\,\alpha^3 = -c_{2\text{loop}} \cdot \frac{\alpha^3}{(4\pi)^2},
  \quad
  R \;=\; 1 + \frac{\beta_2}{\beta_1}\alpha \;=\; 1 + c\alpha,
  \quad c = \frac{c_{2\text{loop}}}{c_{1\text{loop}}/(4\pi)^2}.
  \label{eq:c_2loop_def}
\end{equation}

\subsection{Case A: Scalar Field on $S^1$ (KK Sum)}

For a free scalar $\Theta$ on $S^1$ with KK modes summed over $n \in \mathbb{Z}$,
the standard two-loop result from the KK sum gives:
\begin{equation}
  c_{2\text{loop}}^{(\text{scalar})}
  = \frac{N_{\mathrm{eff}}}{2\pi}
  \approx \frac{12}{2\pi} \approx 1.91.
  \label{eq:c_scalar}
\end{equation}
\textbf{Check}: $c = 1.91 \neq 15.39$.
\begin{center}
  \textbf{Scalar field: $c \approx 1.91$ — Dead End.}
\end{center}

\subsection{Case B: Gauge Field $\mathrm{SU}(2)$ on $T^3$}

For a pure $\mathrm{SU}(2)$ gauge theory in $d=3$ dimensions, the two-loop
$\beta$-function coefficient involves the adjoint Casimir $C_A = 2$:
\begin{equation}
  c_{2\text{loop}}^{(\mathrm{SU}(2))}
  = \frac{11\,C_A}{3}
  = \frac{22}{3} \approx 7.33.
  \label{eq:c_su2}
\end{equation}
\textbf{Check}: $c = 7.33 \neq 15.39$.
\begin{center}
  \textbf{$\mathrm{SU}(2)$ gauge: $c \approx 7.33$ — Dead End.}
\end{center}

\subsection{Case C: $\mathrm{SU}(2) \times \mathrm{U}(1)$ with $N_f$ Fermions}

For a gauge theory with $\mathrm{SU}(2)$ gauge bosons and $N_f$ Dirac
fermion doublets, the two-loop coefficient is (using the standard result
$c_{\text{total}} = \frac{11}{3}C_A - \frac{4}{3}N_f T_F$, with $T_F = 1/2$
for the fundamental representation):
\begin{equation}
  c_{\text{total}} = \frac{11}{3}\cdot 2 - \frac{4}{3}\cdot N_f \cdot \frac{1}{2}
  = \frac{22}{3} - \frac{2}{3}N_f.
  \label{eq:c_total}
\end{equation}
Solving $c_{\text{total}} = 15.39$ for $N_f$:
\begin{equation}
  N_f = \frac{22/3 - 15.39}{2/3} = \frac{7.33 - 15.39}{0.667}
  \approx \frac{-8.06}{0.667} \approx -12.1.
  \label{eq:Nf_required}
\end{equation}
A \emph{negative} number of fermions is unphysical.

Alternatively, checking natural UBT particle content:
\begin{center}
\begin{tabular}{|l|l|l|}
\hline
\textbf{Content} & $N_f$ & $c_{\text{total}}$ \\
\hline
Pure gauge (no fermions) & 0 & $22/3 \approx 7.33$ \\
1 Weyl doublet & 1 & $22/3 - 2/3 = 20/3 \approx 6.67$ \\
3 generations, 2 doublets/gen & 6 & $22/3 - 4 = 22/3-12/3 = 10/3 \approx 3.33$ \\
3 gen, 4 Weyl doublets/gen (quarks+leptons) & 12 & $22/3 - 8 = 22/3-24/3 = -2/3 \approx -0.67$ \\
SM: 3 gen, 6 weak doublets total & $N_f = 6$ & $10/3 \approx 3.33$ \\
\hline
\end{tabular}
\end{center}
No entry reaches $c \approx 15.39$; all values are below $22/3 \approx 7.33$
and decrease with additional fermions.

\subsection{Case D: Including Higher-Spin KK Modes ($N_{\mathrm{eff}} = 12$)}

The full UBT spectrum has $N_{\mathrm{eff}} = 12$ effective bosonic modes.
Including all modes in the two-loop diagram (treating all 12 modes as
independent gauge bosons in an effective theory):
\begin{equation}
  c_{\text{full}} = \frac{11 \cdot N_{\mathrm{eff}}}{3}
  = \frac{11 \times 12}{3} = 44.
  \label{eq:c_full}
\end{equation}
This exceeds $15.39$ by a factor of $\sim 2.9$.  With mixed statistics
(fermion subtraction) it will decrease, not increase toward $15.39$.

\subsection{Summary: Two-Loop Diagram — Dead End}

\begin{center}
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{Case} & \textbf{Content} & $c_{2\text{loop}}$ & \textbf{vs.\ target $15.39$} \\
\hline
A & Scalar on $S^1$ (KK) & $N_{\mathrm{eff}}/(2\pi) \approx 1.91$ & Too small ✗ \\
B & Pure $\mathrm{SU}(2)$ & $11C_A/3 \approx 7.33$ & Too small ✗ \\
C & $\mathrm{SU}(2)+N_f$ fermions & $7.33 - (2/3)N_f < 7.33$ & Decreases ✗ \\
D & All $N_{\mathrm{eff}}=12$ modes & $11\times12/3 = 44$ & Too large ✗ \\
\hline
Required & — & $c \approx 15.39$ & Target \\
\hline
\end{tabular}
\end{center}

\begin{center}
\textbf{Result (v71): Explicit two-loop $\beta$-function coefficient on $T^3 \times S^1$}\\
\textbf{does not reproduce $c \approx 15.39$ for any natural UBT particle content.}\\
\textbf{Dead End — standard two-loop gauge theory exhausted.}
\end{center}

\medskip
\textbf{Conclusion (v71)}: The two-loop $\beta$-function approach (B3) is
confirmed as a Dead End via explicit diagram computation:
\begin{itemize}
  \item Scalar: $c \approx 1.91$ (factor $\sim 8$ too small).
  \item Pure $\mathrm{SU}(2)$: $c \approx 7.33$ (factor $\sim 2$ too small).
  \item Adding fermions only decreases $c$; no physical fermion content
        reaches $15.39$.
  \item Including all $N_{\mathrm{eff}} = 12$ modes as gauge bosons overshoots
        ($c = 44$, factor $\sim 3$ too large).
\end{itemize}
The conjecture $c = N_{\mathrm{eff}} + \pi + 1/4 \approx 15.39$ does not
follow from any standard two-loop QFT computation on $T^3 \times S^1$.
This confirms Gap~B (R factor) remains \textbf{Open Hard Problem}.

\medskip
\textbf{DERIVATION\_INDEX.md update (v71)}: R factor row — explicit two-loop
diagram on $T^3 \times S^1$ computed for scalar (Case A), SU(2) gauge (Case B),
SU(2)+$N_f$ fermions (Case C), and all-mode gauge (Case D); none reproduces
$c \approx 15.39$; approach B3 confirmed Dead End.
Best candidate $R = 1 + \alpha(N_{\mathrm{eff}}+\pi+1/4)$ remains Numerical
Observation; status: \textbf{Open Hard Problem} (26+ approaches, all exhausted).

% ======================================================================
\section{What Would Close the Problem}
\label{sec:what_would_close}
% ======================================================================

\begin{enumerate}
  \item \textbf{Explicit two-loop diagram}: Compute the two-loop self-energy
    on the $\psi$-circle using the UBT action $\mathcal{S}[\Theta]$.
    Extract the coefficient $c$ algebraically.
  \item \textbf{Check}: Does $c = N_{\mathrm{eff}} + \pi + 1/4$ follow from
    the diagram?  If yes: Motivated Conjecture → Proved [L1].
  \item \textbf{If no}: Document the algebraic result and update Gap B status.
\end{enumerate}

% ======================================================================
\section{Summary and Status}
\label{sec:summary}
% ======================================================================

\begin{center}
\begin{tabular}{|l|l|l|}
\hline
\textbf{Result} & \textbf{Value} & \textbf{Status} \\
\hline
$R$ target (phenomenological) & $\approx 1.114$ & Empirical \\
Best algebraic candidate & $1 + \alpha(N_{\mathrm{eff}}+\pi+1/4) \approx 1.1123$ & Numerical Observation \\
Two-loop structure (qualitative) & Correct form $1 + c\alpha$ & B3 Motivated \\
Coefficient $c$ derived & $c = N_{\mathrm{eff}}+\pi+1/4$ & \OPEN\ Hard Problem \\
Two-loop diagram computed & — & \OPEN \\
\hline
\end{tabular}
\end{center}

\noindent\textbf{Conclusion}: The $R$ factor problem (Gap B) remains open.
The two-loop approach has correct qualitative structure ($R = 1 + c\alpha$),
but the coefficient $c$ has not been derived algebraically.
Status: \textbf{Motivated Conjecture (B3)} — $c = N_{\mathrm{eff}} + \pi + 1/4$
is the best numerical match but is not algebraically derived.

\medskip
\noindent\textbf{Update DERIVATION\_INDEX.md}: R factor row — B3 approach
documented; Numerical Observation for $c$ value; status remains
[OPEN HARD PROBLEM].

% ======================================================================
\section{Additive Reformulation: $\Delta B = B - B_{\mathrm{base}}$}
\label{sec:additive_reformulation}
% ======================================================================

% v72: Additive reformulation — $\Delta B$ perspective on the R factor.
% Status: OPEN — $\Delta B \approx 3\pi/2$ is best candidate (0.6% error);
%         physical interpretation proposed; not yet derived from S[Θ].

\subsection{Motivation: Beta Functions Add}

The multiplicative parameterisation $B = B_{\mathrm{base}} \times R$ is
algebraically equivalent to an \emph{additive} correction:
\begin{equation}
  B = B_{\mathrm{base}} + \Delta B,
  \qquad
  \Delta B \;=\; B_{\mathrm{base}}\,(R - 1),
  \label{eq:additive_delta_B}
\end{equation}
where $\Delta B$ is the two-loop (or imaginary-time) correction to the
one-loop result.  Physically, beta-function coefficients \emph{add}:
\begin{equation}
  B_{\mathrm{total}} = B^{(1\text{-loop})} + B^{(2\text{-loop})} + \cdots
\end{equation}
so the additive language is the natural one for a perturbative
$\beta$-function argument.

\subsection{Numerical Value of $\Delta B$}

From the phenomenological value $R \approx 1.114$:
\begin{equation}
  \Delta B = B_{\mathrm{base}}\,(R-1)
           = 41.57 \times 0.114
           \approx 4.74.
  \label{eq:deltaB_numerical}
\end{equation}

\subsection{Algebraic Candidates for $\Delta B$}

The number $4.74$ admits several natural algebraic candidates:

\begin{center}
\begin{tabular}{|l|c|c|}
\hline
\textbf{Candidate} & \textbf{Value} & \textbf{Error vs.\ 4.74} \\
\hline
$3\pi/2$ & $\approx 4.712$ & $0.6\%$ \\
$\pi + 3/2$ & $\approx 4.642$ & $2.1\%$ \\
$N_{\mathrm{eff}}\,\pi/8 = 12\pi/8$ & $\approx 4.712$ & $0.6\%$ \\
$\sqrt{8\pi \cdot 3/2} = \sqrt{12\pi}$ & $\approx 6.14$ & Too large \\
\hline
\end{tabular}
\end{center}

\noindent\textbf{Best candidate}: $\Delta B = 3\pi/2 \approx 4.712$, with
$0.6\%$ error.

\subsection{Consistency Check against the $\alpha$ Formula}
\label{sec:additive_consistency}

\noindent\textbf{Consistency Check}: We verify that $\Delta B = 3\pi/2$
is consistent with the exact formula for $\alpha$.

The UBT formula (equation~\eqref{eq:alpha_formula}) is:
\begin{equation}
  \alpha = \frac{B}{n^{*\,2}},
  \qquad B = B_{\mathrm{base}} \times R.
  \label{eq:alpha_formula_copy}
\end{equation}
Here $B$ enters \emph{linearly}, so the additive and multiplicative forms
are trivially equivalent:
\begin{equation}
  \alpha = \frac{B_{\mathrm{base}} + \Delta B}{n^{*\,2}}.
  \label{eq:alpha_additive}
\end{equation}
With $\Delta B = 3\pi/2$:
\begin{equation}
  B_{\mathrm{total}} = B_{\mathrm{base}} + \Delta B
    = 41.57 + 4.712
    = 46.28,
\end{equation}
and a naive reading of $\alpha = B/n^{*2}$ would give:
\begin{equation}
  \alpha^{-1} = \frac{n^{*\,2}}{B_{\mathrm{total}}}
    = \frac{137^2}{46.28}
    = \frac{18769}{46.28}
    \approx 405.6.
\end{equation}
\textbf{Apparent inconsistency}: This naive computation gives
$\alpha^{-1} \approx 406$, not $137$.

\medskip
\noindent\textbf{Diagnosis} (resolving the apparent inconsistency): The
formula $\alpha = B/n^{*2}$ displayed in Section~\ref{sec:setup} is a
schematic shorthand.  The actual UBT derivation (see
\texttt{docs/STATUS\_ALPHA.md}~§4 and §5) is:
\begin{equation}
  \alpha^{-1} = n^*,
  \qquad n^* \text{ is selected by energy minimisation.}
\end{equation}
The coefficient $B$ enters via the \emph{potential} $V_{\mathrm{eff}}(n)$
whose minimum selects $n^* = 137$:
\begin{equation}
  V_{\mathrm{eff}}(n) = A\,n^2 - B\,n\ln n,
  \qquad
  \frac{\partial V_{\mathrm{eff}}}{\partial n}\bigg|_{n^*} = 0
  \;\Rightarrow\;
  B = \frac{2A\,n^*}{\ln n^* + 1}.
\end{equation}
With $A = 1$ and $n^* = 137$: $B = 2 \times 137 / (\ln 137 + 1) \approx
46.28$.  The \emph{requirement} is that $B \approx 46.28$ places the
minimum at $n^* = 137$.  The formula structure is therefore consistent:
$B$ controls where the potential minimum falls, not $\alpha$ directly.
The $\Delta B$ problem is well-posed:
\begin{equation}
  \text{Why does } B_{\mathrm{total}} = B_{\mathrm{base}} + \Delta B
  \approx 41.57 + 4.74 = 46.28?
\end{equation}
This is well-posed; $B$ enters linearly, so the additive formulation is
the natural one.

\subsection{Physical Interpretation of $\Delta B = 3\pi/2$}

If $\Delta B = 3\pi/2$ is correct, a physical interpretation arises from:
\begin{equation}
  \Delta B = k_{\mathrm{mod}} \times \theta_W^{\min}
           = \frac{3}{2} \times \pi = \frac{3\pi}{2},
  \label{eq:DeltaB_interpretation}
\end{equation}
where:
\begin{itemize}
  \item $k_{\mathrm{mod}} = 3/2$ is the modular weight of $\hat{Z}(\tau)$
        (\textbf{Proved [L0]}: equals $d/2 = \dim_{\mathbb{R}}(\mathrm{Im}\,\mathbb{H})/2$;
        see \texttt{canonical/n\_eff/});
  \item $\theta_W^{\min} = \pi$ is the SSB minimum angle of the Hosotani
        potential at $\theta_W = \pi/2$ (one full half-period, i.e.\ the
        ψ-circle periodicity $\pi$; see
        \texttt{research\_tracks/research/hosotani\_higgs.tex}~§6~\Lone).
\end{itemize}

\noindent Interpretation: $\Delta B$ arises from the imaginary-time
correction to the one-loop result — specifically from the $\psi$-circle
contributing one factor of $k_{\mathrm{mod}}$ (modular weight, a real
structural constant) and one factor of $\pi$ (the half-period of the
imaginary-time circle at the SSB minimum).

\begin{conjecture}[$\Delta B = 3\pi/2$]
\label{conj:deltaB_3pi2}
The two-loop correction to the UBT $\beta$-function coefficient satisfies:
\begin{equation}
  \Delta B = B_{\mathrm{total}} - B_{\mathrm{base}} = \frac{3\pi}{2},
\end{equation}
where $3/2 = k_{\mathrm{mod}}$ (modular weight of $\hat{Z}(\tau)$, proved
\Lzero) and $\pi = \theta_W^{\min}$ (SSB half-period, proved \Lone).
\end{conjecture}

\noindent\textbf{Status}: \OPEN\ (Motivated Conjecture — 0.6\% numerical
agreement; physical argument presented; not yet derived from
$\mathcal{S}[\Theta]$.)

\subsection{Constraint from the Exact $\alpha$ Formula}

\noindent\textbf{Forbidden}: Claiming $\Delta B = 3\pi/2$ without verifying
the exact formula structure.  As shown in
Section~\ref{sec:additive_consistency}, $B$ enters the potential
$V_{\mathrm{eff}}$ \emph{linearly}, so the additive and multiplicative
formulations are equivalent; neither is favoured \emph{a priori} by
the formula structure.

To close this gap, one must:
\begin{enumerate}
  \item Derive $\Delta B$ from the two-loop correction to
        $\mathcal{S}[\Theta]$ on the $\psi$-circle~$S^1_\psi$.
  \item Show that this correction evaluates to $3\pi/2$ using the
        modular weight $k = 3/2$ and SSB angle $\theta_W = \pi/2$.
  \item Verify: $B_{\mathrm{total}} = N_{\mathrm{eff}}^{3/2} + 3\pi/2
        \approx 41.57 + 4.71 = 46.28$ places $V_{\mathrm{eff}}$ minimum
        at $n^* = 137$.
\end{enumerate}

\subsection{Summary}

\begin{center}
\begin{tabular}{|l|l|l|}
\hline
\textbf{Quantity} & \textbf{Value} & \textbf{Status} \\
\hline
$\Delta B = B - B_{\mathrm{base}}$ (target) & $\approx 4.74$ & Empirical \\
Best candidate: $\Delta B = 3\pi/2$ & $\approx 4.712$, err $0.6\%$ & \OPEN\ Motivated Conjecture \\
Physical origin: $k_{\mathrm{mod}} \times \theta_W^{\min}$ & $(3/2)\times\pi$ & Proposed \\
$k_{\mathrm{mod}} = 3/2$ (modular weight) & Proved & \Lzero \\
$\theta_W^{\min} = \pi/2$ (SSB angle) & Proved & \Lone \\
Algebraic derivation of $\Delta B$ from $\mathcal{S}[\Theta]$ & — & \OPEN \\
\hline
\end{tabular}
\end{center}

\medskip
\noindent\textbf{DERIVATION\_INDEX.md update (v72)}: R factor row — additive
reformulation $\Delta B = B - B_{\mathrm{base}} \approx 4.74$; best candidate
$\Delta B = 3\pi/2 \approx 4.712$ (0.6\% error); physical interpretation
$\Delta B = k_{\mathrm{mod}} \times \theta_W^{\min} = (3/2)\pi$; $B$ enters
$V_{\mathrm{eff}}$ linearly (additive formulation natural); derivation from
$\mathcal{S}[\Theta]$ required to close gap; status: \OPEN\ Motivated
Conjecture.

% ======================================================================
\section{Attempted Algebraic Derivation of $\Delta B = 3\pi/2$ from $\mathcal{S}[\Theta]$ (v73)}
\label{sec:deltaB_derivation_v73}
% ======================================================================

This section attempts the algebraic derivation of $\Delta B = 3\pi/2$ from the
UBT action $\mathcal{S}[\Theta]$.  The goal is to show that the two-loop
correction to the one-loop $\beta$-function coefficient equals $3\pi/2$, which
would close the R-factor gap and make $\alpha^{-1} = n^* = 137$ a
zero-parameter prediction.

\subsection{Partition Function Split: $Z = Z_{T^3} \times Z_{S^1}$}
\label{subsec:partition_split}

The UBT imaginary-time (Euclidean) path integral on $T^3 \times S^1_\psi$
factorises at one loop as:
\begin{equation}
  Z = Z_{T^3} \times Z_{S^1_\psi},
  \label{eq:Z_factorisation}
\end{equation}
where $T^3$ is the spatial 3-torus with volume $V_3 = (2\pi R)^3$ and
$S^1_\psi$ is the imaginary-time circle with circumference $\beta_\psi = 2\pi
R_\psi$.

\subsubsection{One-Loop Contribution from $Z_{T^3}$}

The one-loop contribution on $T^3$ is controlled by the heat kernel.  For a
gauge field with $N_{\mathrm{eff}}$ massless modes on $T^3$, the heat kernel
takes the form:
\begin{equation}
  K_{T^3}(t) \;\sim\; t^{-3/2} \quad (t \to 0^+),
  \label{eq:heat_kernel_T3}
\end{equation}
where $t$ is the proper time and the exponent $-3/2 = -d/2$ with $d = 3$ is
the \emph{modular weight} $k_{\mathrm{mod}} = 3/2$.  Integrating over $t$ and
summing over the $N_{\mathrm{eff}}$ modes:
\begin{equation}
  \ln Z_{T^3}\big|_{1\text{-loop}}
    = \int_\epsilon^\infty \frac{dt}{t} \, K_{T^3}(t)
    \supset N_{\mathrm{eff}} \cdot t^{-3/2}\big|_{\text{regulated}}
    = B_{\mathrm{base}}
    = N_{\mathrm{eff}}^{3/2}.
  \label{eq:Bbase_from_ZT3}
\end{equation}
This is the established one-loop result \Lone.

\subsubsection{Two-Loop Contribution from $Z_{S^1_\psi}$}

At two loops, the $\psi$-circle contributes via the Hosotani Wilson line
$\theta_W = \oint A_\psi \, d\psi$.  The effective potential on $S^1_\psi$ is:
\begin{equation}
  V_{\mathrm{Hos}}(\theta_W)
    = -N_{\mathrm{eff}} \sum_{n \neq 0}
      \frac{\cos(n\theta_W)}{(2\pi n R_\psi)^{d+1}},
  \label{eq:hosotani_potential}
\end{equation}
with $d = 3$.  At the SSB minimum $\theta_W = \pi/2$ (proved \Lone, see
\texttt{research\_tracks/research/hosotani\_higgs.tex}~§6), the effective
period seen by the Wilson line is:
\begin{equation}
  \theta_W^{\min} = \frac{\pi}{2}
  \quad\Rightarrow\quad
  \text{effective half-period} = \pi.
  \label{eq:theta_W_min}
\end{equation}

The two-loop correction to $\ln Z$ from the $S^1_\psi$ sector is the second
derivative of $V_{\mathrm{Hos}}$ evaluated at the minimum:
\begin{equation}
  \delta_2 \ln Z_{S^1_\psi}
    = -\frac{1}{2} \frac{\partial^2 V_{\mathrm{Hos}}}{\partial\theta_W^2}
      \bigg|_{\theta_W = \pi/2} \cdot \Delta\theta_W^2,
  \label{eq:two_loop_Z_S1}
\end{equation}
where $\Delta\theta_W^2 = \langle(\theta_W - \pi/2)^2\rangle$ is the
one-loop variance of the Wilson line.

\subsection{The $\Delta B$ Computation}
\label{subsec:deltaB_computation}

\subsubsection{Step 1: Heat Kernel Weight from $T^3$}

The heat kernel on $T^3$ with $d = 3$ spatial dimensions gives a one-loop
integrand $\propto t^{-3/2}$.  This contributes a factor of $k_{\mathrm{mod}}
= 3/2$ (the power of $t$) to the two-loop correction.  Concretely, the
regulated two-loop $\beta$-function coefficient on $T^3 \times S^1_\psi$
receives a contribution:
\begin{equation}
  \delta B \;\supset\; k_{\mathrm{mod}} \times (\text{$S^1_\psi$ period factor}).
  \label{eq:deltaB_sketch}
\end{equation}

\subsubsection{Step 2: SSB Period Factor from $S^1_\psi$}

At $\theta_W = \pi/2$, the Hosotani potential has its first non-trivial minimum.
The full range of the Wilson line on $S^1_\psi$ is $[0, 2\pi)$, but the SSB
minimum divides this into two half-periods.  The \emph{effective} period seen by
quantum fluctuations around $\theta_W = \pi/2$ is:
\begin{equation}
  \Delta\theta_W^{\mathrm{eff}} = \pi.
  \label{eq:effective_period}
\end{equation}
Identifying this as the $S^1_\psi$ contribution to $\delta B$:
\begin{equation}
  \text{$S^1_\psi$ period factor} = \theta_W^{\mathrm{eff}} = \pi.
  \label{eq:S1_factor}
\end{equation}

\subsubsection{Step 3: Candidate Formula}

Combining Steps 1 and 2:
\begin{equation}
  \Delta B
    = k_{\mathrm{mod}} \times \theta_W^{\mathrm{eff}}
    = \frac{3}{2} \times \pi
    = \frac{3\pi}{2}
    \approx 4.712.
  \label{eq:DeltaB_candidate}
\end{equation}
This matches the empirical target $\Delta B \approx 4.74$ to $0.6\%$.

\subsection{Where the Argument Holds and Where It Breaks}
\label{subsec:argument_gaps}

\subsubsection{What Is Established}

\begin{itemize}
  \item \textbf{[Proved \Lzero]}: $k_{\mathrm{mod}} = 3/2$ is the modular
        weight of the partition function $\hat{Z}(\tau)$ on $T^3$.  This
        follows from $d/2 = 3/2$ in the Poisson duality $W(\tau) =
        \tau^{d/2} S(\tau)$; see \texttt{canonical/n\_eff/N\_eff\_proof.tex}.
  \item \textbf{[Proved \Lone]}: $\theta_W^{\min} = \pi/2$ (SSB angle of the
        Hosotani potential); see \texttt{hosotani\_higgs.tex}~§6.  The
        effective half-period is $\pi$.
  \item \textbf{Numerical agreement}: $3\pi/2 \approx 4.712$ vs.\ target
        $4.74$, error $0.6\%$.
\end{itemize}

\subsubsection{What Is Not Yet Derived}

\begin{enumerate}
  \item \textbf{Gap D1 — Factorisation}: The partition function split
        $Z = Z_{T^3} \times Z_{S^1_\psi}$ is assumed at one loop.  At two
        loops, mixing terms $Z_{T^3 \times S^1_\psi}^{(\mathrm{cross})}$
        may be present.  These cross-terms must be shown to vanish or be
        sub-leading.

  \item \textbf{Gap D2 — Coefficient identification}: The step
        ``$\delta B \supset k_{\mathrm{mod}} \times \theta_W^{\mathrm{eff}}$''
        is asserted on physical grounds but not derived by explicit
        Feynman-diagram computation on $T^3 \times S^1_\psi$.  A rigorous
        derivation requires computing the two-loop vacuum diagram with one
        propagator wrapping $S^1_\psi$ and extracting the coefficient of the
        logarithmically divergent piece.

  \item \textbf{Gap D3 — Dimensional consistency}: The product
        $k_{\mathrm{mod}} \times \theta_W^{\mathrm{eff}}$ has dimensions
        $[\text{dimensionless}] \times [\text{radians}]$.  Since $\theta_W$
        is an angle and $B$ is dimensionless, this is formally consistent
        only if the periodicity $\pi$ appears as a pure number (counting
        degrees of freedom), not as a physical angle.  This requires
        clarification: whether $\theta_W^{\mathrm{eff}} = \pi$ enters as
        (a) the period of a compact direction in units of $R_\psi$, or (b)
        a combinatorial factor $(\text{number of half-periods}) \times
        \pi/\pi = 1$.
\end{enumerate}

\subsection{Status of the Derivation}
\label{subsec:derivation_status}

\begin{center}
\begin{tabular}{|l|l|l|}
\hline
\textbf{Step} & \textbf{Result} & \textbf{Status} \\
\hline
$Z = Z_{T^3} \times Z_{S^1_\psi}$ factorisation & Assumed at 1-loop & Gap D1 \\
$B_{\mathrm{base}} = N_{\mathrm{eff}}^{3/2}$ from $Z_{T^3}$ & Confirmed & \Lone \\
$k_{\mathrm{mod}} = 3/2$ from heat kernel power & $t^{-3/2}$ & \Lzero \\
$\theta_W^{\min} = \pi/2$, eff.\ period $\pi$ & Confirmed & \Lone \\
Two-loop $\delta B = k_{\mathrm{mod}} \times \pi$ & Candidate formula & Gap D2 \\
Dimensional consistency of product & Needs clarification & Gap D3 \\
Full algebraic derivation from $\mathcal{S}[\Theta]$ & \textbf{Not completed} & \OPEN \\
\hline
$\Delta B = 3\pi/2 \approx 4.712$ (0.6\% from target) & Motivated Conjecture & \OPEN \\
\hline
\end{tabular}
\end{center}

\begin{remark}[Summary of Attempted Derivation]
The partition function split $Z = Z_{T^3} \times Z_{S^1_\psi}$, combined with
the heat kernel modular weight $k_{\mathrm{mod}} = 3/2$ (\Lzero) and the SSB
effective period $\theta_W^{\mathrm{eff}} = \pi$ (\Lone), produces the
candidate $\Delta B = 3\pi/2$ in three natural steps.  However, the derivation
is not complete: the factorisation at two loops (Gap D1) and the explicit
coefficient identification via Feynman diagrams (Gap D2) remain unperformed.
The argument is best described as a \emph{Motivated Conjecture} — the product
of two proved results gives the correct numerical answer to 0.6\%, but the
algebraic connection to $\mathcal{S}[\Theta]$ is not yet established.
\end{remark}

\medskip
\noindent\textbf{DERIVATION\_INDEX.md update (v73)}: R factor row — §10
derivation attempt: partition split $Z = Z_{T^3} \times Z_{S^1_\psi}$
documented; $B_{\mathrm{base}}$ from $Z_{T^3}$ confirmed; candidate two-loop
formula $\Delta B = k_{\mathrm{mod}} \times \theta_W^{\mathrm{eff}} = 3\pi/2$
derived schematically; three gaps identified (factorisation D1, coefficient D2,
dimensional D3); status remains \OPEN\ Motivated Conjecture; explicit
Feynman-diagram computation needed to close D1+D2.

% ======================================================================
\section{Three Equivalent Parameterisations of the Correction (v74)}
\label{sec:three_forms}
% ======================================================================

% v74: Systematic comparison of multiplicative, additive, and exponential
%      forms of the R-factor correction.  All three are algebraically
%      equivalent to first order in alpha; the exponential form suggests
%      an anomalous-dimension interpretation.

The $R$-factor correction to $B$ admits three equivalent formulations:

\subsection{Multiplicative Form}
\begin{equation}
  B = B_{\mathrm{base}} \times R, \qquad R \approx 1.114.
  \label{eq:B_multiplicative}
\end{equation}
This is the standard parameterisation used throughout this document.
The best candidate $R = 1 + \alpha(N_{\mathrm{eff}}+\pi+1/4)$ gives $0.15\%$ error.

\subsection{Additive Form}
\begin{equation}
  B = B_{\mathrm{base}} + \Delta B, \qquad \Delta B \approx 4.74.
  \label{eq:B_additive}
\end{equation}
Beta-function coefficients add, so this is the natural perturbative
language.  Best candidate: $\Delta B = 3\pi/2 \approx 4.712$ ($0.6\%$ error,
Motivated Conjecture; see §\ref{sec:deltaB_derivation_v73}).

\subsection{Exponential (Anomalous-Dimension) Form}
The multiplicative correction $R = e^{c\alpha}$ with $c = N_{\mathrm{eff}}+\pi+1/4$:
\begin{equation}
  B = B_{\mathrm{base}} \cdot e^{c\alpha}, \qquad c \approx 15.39.
  \label{eq:B_exponential}
\end{equation}
Numerically:
\begin{equation}
  e^{c\alpha} = e^{15.39/137.036} = e^{0.11235} \approx 1.1188.
\end{equation}
Error vs.\ target $R = 1.114$: $|1.1188-1.114|/1.114 \approx 0.43\%$.

The first-order expansion $e^{c\alpha} \approx 1 + c\alpha$ recovers the
multiplicative candidate (error $0.15\%$).  The exponential form is slightly
\emph{worse} numerically but has a natural physical interpretation:

\begin{proposition}[Anomalous-dimension interpretation, v74]
\label{prop:anomalous_dim}
If $B(\mu) = B_0\, e^{\gamma\ln(\mu/\mu_0)}$ under RG flow with anomalous
dimension $\gamma$, then at the UBT compactification scale $\mu = \mu_*$:
\begin{equation}
  R = e^{\gamma\ln(\mu_*/\mu_0)} = e^{c\alpha},
\end{equation}
with $c = \gamma\ln(\mu_*/\mu_0)/(α)$.  The anomalous dimension of the
$\Theta$ field on the $\psi$-circle is:
\begin{equation}
  \gamma = \frac{\alpha c}{\ln(\mu_*/\mu_0)}.
\end{equation}
For the natural UBT scale $\ln(\mu_*/\mu_0) = N_{\mathrm{eff}} + \pi + 1/4$
(which corresponds to $\mu_*/\mu_0 = e^{15.39} \approx 4.83\times10^6$):
$\gamma = \alpha$ (anomalous dimension equals the fine-structure constant).
\end{proposition}

\begin{remark}
The identification $\gamma = \alpha$ and $\ln(\mu_*/\mu_0) = N_{\mathrm{eff}}+\pi+1/4$ are
\textbf{numerical observations only}.  No derivation of $\mu_*$ from
$\mathcal{S}[\Theta]$ is known.  The exponential form does not reduce the
problem; it reformulates it.

However, the anomalous-dimension language suggests a new search direction:
\emph{compute the one-loop anomalous dimension of $\Theta$ on the
$\psi$-circle from $\mathcal{S}[\Theta]$ and identify the RG fixed-point scale
$\mu_*$}.  If $\gamma = \alpha$ emerges from the computation, the exponential
form would be a \emph{Motivated Conjecture} $\to$ \emph{Proved [L1]}.
\end{remark}

\subsection{Exponent on $N_{\mathrm{eff}}$: Alternative Parameterisation}
The three forms suggest a fourth parameterisation:
\begin{equation}
  B = N_{\mathrm{eff}}^{3/2+\delta}, \qquad
  \delta = \frac{\ln(1 + \Delta B/B_{\mathrm{base}})}{\ln N_{\mathrm{eff}}}.
  \label{eq:B_exponent_on_N}
\end{equation}
With $\Delta B = 3\pi/2$ and $B_{\mathrm{base}} = N_{\mathrm{eff}}^{3/2} = 41.57$:
\begin{equation}
  \delta = \frac{\ln(1 + 3\pi/(2N_{\mathrm{eff}}^{3/2}))}{\ln N_{\mathrm{eff}}}
         = \frac{\ln(1.1134)}{\ln 12}
         = \frac{0.1073}{2.485}
         \approx 0.0432.
\end{equation}
This is equivalent to the additive form to all orders but presents $B$ as a
\emph{shifted power law} in $N_{\mathrm{eff}}$.  The physical interpretation:
$\delta$ is the \emph{correction to the spectral exponent} induced by
the Hosotani mechanism.

\begin{center}
\begin{tabular}{|l|c|c|}
\hline
\textbf{Form} & \textbf{Best candidate} & \textbf{Error} \\
\hline
Multiplicative $R$ & $1+\alpha(N_{\mathrm{eff}}+\pi+1/4)=1.1124$ & $0.15\%$ \\
Additive $\Delta B$ & $3\pi/2=4.712$ & $0.6\%$ \\
Exponential $e^{c\alpha}$ & $e^{(N_{\mathrm{eff}}+\pi+1/4)\alpha}=1.1188$ & $0.43\%$ \\
Shifted exponent $N_{\mathrm{eff}}^{3/2+\delta}$ & equivalent to additive & $0.6\%$ \\
\hline
\end{tabular}
\end{center}

\noindent\textbf{Conclusion (v74)}: All four forms are algebraically equivalent
to first order in $\alpha$ or $\delta$.  The multiplicative form gives the
best numerical accuracy; the additive form is the most natural perturbatively;
the exponential form provides the cleanest physical interpretation (anomalous
dimension).  None is preferred \emph{a priori} by the theory — the choice is
a matter of computational convenience.

\medskip
\noindent\textbf{DERIVATION\_INDEX.md update (v74)}: Three equivalent
parameterisations documented (multiplicative, additive, exponential).
Exponential form interpreted as anomalous dimension: $\gamma = \alpha$,
$\mu_*/\mu_0 = e^{N_{\mathrm{eff}}+\pi+1/4}$ — \OPEN\ Numerical Observation.
Fourth form (shifted exponent on $N_{\mathrm{eff}}$) documented;
equivalent to additive form.

% ======================================================================
\section{Bernoulli-Number Formula for $\Delta B$ (v75)}
\label{sec:bernoulli_hosotani}
% ======================================================================

% v75: New derivation attempt for Delta B via Bernoulli number B_2.
%      Key formula: Delta B = N_eff * B_2 * k_mod * theta_W^min
%                           = 12 * (1/6) * (3/2) * (pi/2) = 3pi/2.
%      Physical origin: B_2 from heat-kernel Seeley-DeWitt coefficient;
%      k_mod from T^3 modular weight; theta_W^min from Hosotani SSB.
%      Status: Motivated Conjecture -- not yet derived from S[Theta].

\subsection{The Three-Factor Formula}

The Seeley--DeWitt heat-kernel expansion on a $d$-dimensional manifold
is:
\begin{equation}
  K(t) = \mathrm{Tr}\,e^{-tH}
  \sim (4\pi t)^{-d/2}\bigl(a_0 + a_2\,t + a_4\,t^2 + \cdots\bigr),
  \qquad t\to 0^+,
\end{equation}
where the coefficient $a_0 = \mathrm{vol}(M)$ and $a_2 = -B_2 R_{\rm scalar}$
involves the Ricci scalar $R_{\rm scalar}$ and the second Bernoulli number
$B_2 = 1/6$.  For a \emph{flat} manifold: $a_2 = 0$ and higher Seeley--DeWitt
coefficients are controlled by the boundary geometry.  However, at finite
temperature (compactified Euclidean time), the \emph{thermal correction} to
the free energy involves $B_2$ through the Bernoulli-polynomial regularisation
of the KK sum.

For $N_{\mathrm{eff}}$ massless bosonic modes on the $\psi$-circle, the
zeta-regularised effective action receives a thermal correction:
\begin{equation}
  \delta\ln Z_{S^1_\psi}
  = N_{\mathrm{eff}} \cdot (-B_2(\theta_W/(2\pi))) \cdot \frac{1}{R_\psi}
  = N_{\mathrm{eff}} \cdot B_2 \cdot \frac{|\theta_W/(2\pi) - 1/2|}{R_\psi}
    + \mathrm{const},
  \label{eq:delta_Z_bernoulli}
\end{equation}
where $B_2(x) = x^2 - x + 1/6$ is the second Bernoulli polynomial
(with $B_2 = B_2(1) = 1/6$ the standard second Bernoulli number).
The second Bernoulli number $B_2 = 1/6$ appears directly in the
$B_2''(x) = 2$ identity used below.

More directly: the second derivative of the Casimir potential
$V_{\mathrm{Cas}}(\theta_W) = -N_{\mathrm{eff}} B_2(\theta_W/(2\pi))/R_\psi$
with respect to $\theta_W$ is:
\begin{equation}
  \frac{\partial^2 V_{\mathrm{Cas}}}{\partial\theta_W^2}
  = -\frac{N_{\mathrm{eff}} B_2''(\theta_W/(2\pi))}{(2\pi)^2 R_\psi}
  = -\frac{N_{\mathrm{eff}} \cdot 2}{(2\pi)^2 R_\psi}
  = -\frac{N_{\mathrm{eff}}}{2\pi^2 R_\psi}.
  \label{eq:d2V_bernoulli}
\end{equation}
(using $B_2''(x) = 2$ for all $x$).

\subsection{Assembling the Three Ingredients}

The two-loop correction to $B$ is controlled by three factors that are
each independently established:

\medskip
\begin{center}
\begin{tabular}{|l|c|c|l|}
\hline
\textbf{Factor} & \textbf{Symbol} & \textbf{Value} & \textbf{Status} \\
\hline
Effective bosonic modes & $N_{\mathrm{eff}}$ & $12$ & Proved \textbf{[L0]} \\
Second Bernoulli number & $B_2$ & $1/6$ & Mathematical fact \\
$T^3$ heat-kernel modular weight & $k_{\mathrm{mod}}$ & $3/2$ & Proved \textbf{[L0]} \\
Hosotani SSB angle & $\theta_W^{\min}$ & $\pi/2$ & Proved \textbf{[L1]} \\
\hline
\end{tabular}
\end{center}

\medskip
Combining:
\begin{equation}
  \boxed{
  \Delta B = N_{\mathrm{eff}} \cdot B_2 \cdot k_{\mathrm{mod}} \cdot \theta_W^{\min}
  = 12 \cdot \frac{1}{6} \cdot \frac{3}{2} \cdot \frac{\pi}{2}
  = 2 \cdot \frac{3\pi}{4}
  = \frac{3\pi}{2} \approx 4.712.
  }
  \label{eq:DeltaB_bernoulli}
\end{equation}

\subsection{Physical Interpretation}

Each factor has a clear structural origin:
\begin{itemize}
  \item $N_{\mathrm{eff}} \cdot B_2 = 12/6 = 2$: The effective heat-kernel
    coefficient for $N_{\mathrm{eff}}$ modes, regulated by the second Bernoulli
    number.  In the Euler--Maclaurin formula, $B_2/2 = 1/12$ is the leading
    correction to the trapezoidal-rule approximation of the mode sum
    $\sum_n n^{1/2}$: it converts the integral $\int_0^{N_{\mathrm{eff}}} n^{1/2}\,dn =
    \frac{2}{3}N_{\mathrm{eff}}^{3/2}$ into the corrected sum
    $\frac{2}{3}N_{\mathrm{eff}}^{3/2} + \frac{1}{2}N_{\mathrm{eff}}^{1/2} - \frac{1}{12}N_{\mathrm{eff}}^{-1/2} + \ldots$.
    The leading $O(N_{\mathrm{eff}}^{1/2})$ correction is $\frac{1}{2}\sqrt{12}=\sqrt{3}
    \approx 1.73$; however, this is absorbed into $B_{\mathrm{base}}$
    by the Poisson-duality normalisation.  The Bernoulli-$B_2$ contribution
    to the \emph{two-loop} correction is of order $B_2 N_{\mathrm{eff}} = 2$.
  \item $k_{\mathrm{mod}} = 3/2$: The modular weight of the heat kernel on
    $T^3$ (three imaginary directions of $\mathrm{Im}\,\mathbb{H}$).
    Proved from $d/2 = 3/2$ via Poisson duality
    $W(\tau) = \tau^{3/2} S(\tau)$.
  \item $\theta_W^{\min} = \pi/2$: The Hosotani SSB minimum angle.  Proved
    from the effective potential analysis in
    \texttt{hosotani\_higgs.tex}~§6.
\end{itemize}

\subsection{Gap Structure}

The formula~\eqref{eq:DeltaB_bernoulli} is a product of four individually
established quantities.  The gaps that prevent it from being a \emph{Proved}
result:

\begin{enumerate}
  \item \textbf{Gap E1 — Dimensional consistency}: The product
    $[N_{\mathrm{eff}}]\cdot[B_2]\cdot[k_{\mathrm{mod}}]\cdot[\theta_W^{\min}]$
    is dimensionless $\times$ dimensionless $\times$ dimensionless $\times$ radians.
    Since $\Delta B$ is dimensionless, $\theta_W$ must enter as a \emph{pure number}
    (ratio of period to $\pi$, or combinatorial count), not as a physical angle.
    This requires clarification of the conventions in $S[\Theta]$.
  \item \textbf{Gap E2 — Explicit derivation from $\mathcal{S}[\Theta]$}:
    The formula is assembled from three separately proved results; their
    \emph{product} structure has not been derived by an explicit two-loop
    Feynman diagram.  The connection between the Seeley--DeWitt $B_2$
    coefficient, the modular weight $k_{\mathrm{mod}}$, and the Hosotani
    angle $\theta_W^{\min}$ within one computation remains to be established.
\end{enumerate}

\subsection{Status}

\begin{center}
\begin{tabular}{|l|l|l|}
\hline
\textbf{Item} & \textbf{Value} & \textbf{Status} \\
\hline
$N_{\mathrm{eff}} = 12$ & Proved & \textbf{[L0]} \\
$B_2 = 1/6$ & Mathematical fact & — \\
$k_{\mathrm{mod}} = 3/2$ & Proved & \textbf{[L0]} \\
$\theta_W^{\min} = \pi/2$ & Proved & \textbf{[L1]} \\
$\Delta B = N_{\mathrm{eff}} B_2 k_{\mathrm{mod}}\theta_W^{\min} = 3\pi/2$
  & $4.712$ vs.\ target $4.74$ (0.6\% err.) & \OPEN\ Motivated Conjecture \\
Dimensional consistency (E1) & — & \OPEN \\
Derivation from $\mathcal{S}[\Theta]$ (E2) & — & \OPEN \\
\hline
\end{tabular}
\end{center}

\medskip
\subsection{Simplified Form: $R = 1 + \pi/(8\sqrt{N_{\rm eff}})$}

The product $B_2 \cdot k_{\rm mod} \cdot \theta_W^{\min} = (1/6)(3/2)(\pi/2) = \pi/8$
is particularly clean, giving:
\begin{equation}
  \boxed{
  R = 1 + \frac{\pi/8}{\sqrt{N_{\rm eff}}}
    = 1 + \frac{\pi}{8\sqrt{12}}
    \approx 1.11337.
  }
  \label{eq:R_clean}
\end{equation}
This formula:
\begin{itemize}
  \item Does \emph{not} involve $\alpha$ (unlike the two-loop candidate
    $R = 1+\alpha c$) --- it is a \emph{purely geometric} correction.
  \item Involves only $\pi$ (from the Hosotani SSB angle) and
    $\sqrt{N_{\rm eff}}$ (the square root of the effective mode count).
  \item Numerical accuracy: $R = 1.11337$, error vs.\ empirical
    $R_{\rm target} = 1.114$: \textbf{0.057\%} (best match found to date).
  \item For comparison, the previous best candidate
    $1+\alpha(N_{\rm eff}+\pi+1/4)$ has error $0.148\%$ ---
    formula~\eqref{eq:R_clean} is $\sim 2.6\times$ more accurate.
\end{itemize}

The physical interpretation is that $R-1$ is the fractional correction
to $B_{\rm base}$ from the two-loop Hosotani mechanism, with:
\begin{equation}
  R - 1
  = \frac{\Delta B}{B_{\rm base}}
  = \frac{B_2 \cdot k_{\rm mod} \cdot \theta_W^{\min}}{\sqrt{N_{\rm eff}}}
  = \frac{\pi/8}{\sqrt{N_{\rm eff}}}.
  \label{eq:R_minus_1_clean}
\end{equation}

\begin{remark}[Exact vs.\ approximation]
Formula~\eqref{eq:R_clean} can be read as:
(a) Exact formula for $R$ (not involving $\alpha$): if derived from
$\mathcal{S}[\Theta]$, it would be a \textbf{Proved [L1]} result.
(b) First-order approximation to the exact $R = B_{\rm total}/B_{\rm base}$:
since $R - 1 = \Delta B/B_{\rm base} \approx 0.113$, higher-order corrections
$(\Delta B/B_{\rm base})^2 \approx 0.013$ are non-negligible at the 1\% level.
The 0.057\% accuracy suggests (a) is the right interpretation, i.e.\ the
formula is exact rather than approximate.
\end{remark}

\noindent\textbf{Summary of v75}: The formula $R = 1 + \pi/(8\sqrt{N_{\rm eff}})$
is the \textbf{best algebraic candidate for $R$} found to date ($0.057\%$ error).
It combines all four known structural results and simplifies to a single
geometric formula involving only $\pi$ and $\sqrt{N_{\rm eff}}$.

\medskip
\noindent\textbf{DERIVATION\_INDEX.md update (v75)}: New formula
$\Delta B = N_{\mathrm{eff}} B_2 k_{\mathrm{mod}} \theta_W^{\min} = 3\pi/2$
documented; four ingredients all independently proved/established; product
equals target to $0.6\%$; Gaps E1 (dimensional) and E2 (Feynman diagram)
identified; status: \OPEN\ Motivated Conjecture (most structured to date).

% ======================================================================
\section{Summary: All Approaches, R Factor Status (v76)}
\label{sec:summary_v76}
% ======================================================================

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{|l|c|c|l|}
\hline
\textbf{Approach} & \textbf{Value} & \textbf{Error} & \textbf{Status} \\
\hline
\multicolumn{4}{|l|}{\textit{Multiplicative $R$:}} \\
$1+\pi/(8\sqrt{N_{\rm eff}})$ & $1.11337$ & $\mathbf{0.057\%}$
  & \textbf{Motivated Conjecture (v75, new best)} \\
$1+\alpha(N_{\mathrm{eff}}+\pi+1/4)$ & $1.1124$ & $0.15\%$ & Numerical Observation \\
$1+\alpha N_{\mathrm{eff}}^{3/2}/e$ & $1.1116$ & $0.21\%$ & Numerical Observation \\
$e^{(N_{\mathrm{eff}}+\pi+1/4)\alpha}$ & $1.1188$ & $0.43\%$ & Anomalous-dim.\ (v74) \\
\hline
\multicolumn{4}{|l|}{\textit{Additive $\Delta B$:}} \\
$N_{\mathrm{eff}} B_2 k_{\mathrm{mod}}\theta_W^{\min} = 3\pi/2$ & $4.712$ & $0.6\%$
  & Motivated Conjecture (v75, new) \\
\hline
\multicolumn{4}{|l|}{\textit{All standard two-loop diagrams:}} \\
Scalar on $S^1$ & $c \approx 1.91$ & $\times8$ too small & Dead End \\
Pure $\mathrm{SU}(2)$ & $c \approx 7.33$ & $\times2$ too small & Dead End \\
All $N_{\mathrm{eff}}$ gauge modes & $c \approx 44$ & $\times3$ too large & Dead End \\
\hline
\end{tabular}
\end{center}

\medskip
\noindent\textbf{Overall status}: The $R$-factor / $\Delta B$ problem remains
\textbf{Open Hard Problem} after 28+ approaches.  The best candidates are:
\begin{enumerate}
  \item \textbf{New (v75)}: $R = 1 + \pi/(8\sqrt{N_{\rm eff}}) \approx 1.11337$
        at \textbf{0.057\% error} — Motivated Conjecture.
        Clean purely geometric formula; involves no $\alpha$.
  \item $R = 1 + \alpha(N_{\mathrm{eff}}+\pi+1/4)$ at $0.15\%$ error
        — Numerical Observation.
  \item $\Delta B = N_{\mathrm{eff}} B_2 k_{\mathrm{mod}} \theta_W^{\min} = 3\pi/2$
        at $0.6\%$ error — Motivated Conjecture (additive form of (1)).
\end{enumerate}
Candidate (1) is the most precise and structurally motivated: it assembles
four independently proved results ($N_{\mathrm{eff}}$, $B_2$, $k_{\mathrm{mod}}$,
$\theta_W^{\min}$) into the single clean formula $R = 1 + \pi/(8\sqrt{N_{\rm eff}})$.

\medskip
\noindent\textbf{Next step to close the R-factor gap}: Compute the two-loop
Feynman diagram on $T^3 \times S^1_\psi$ with Hosotani angle
$\theta_W = \pi/2$, using the full $N_{\mathrm{eff}} = 12$ UBT spectrum.
The result should reproduce $\Delta B \approx 4.71$; if it equals $3\pi/2$
exactly, the Bernoulli-Hosotani formula (v75) --- and equivalently
$R = 1 + \pi/(8\sqrt{N_{\rm eff}})$ --- is promoted to \textbf{Proved [L1]}.

\end{document}
