%% ================================================
%% SPECULATIVE / WIP — Not part of CORE claims.
%% ================================================
\maketitle

\input{../THEORY_STATUS_DISCLAIMER}
\SpeculativeContentWarning

\paragraph{Note on Canonical Complex Time in Dark Energy.}
\textbf{Complex time $\tau = t + i\psi$ is the canonical time formulation in UBT (AXIOM B)}: Dark energy phenomenology uses canonical complex time. Cosmological dynamics occur on very large scales where local commutators $[\Theta_i, \Theta_j] \approx 0$ are satisfied, the phase field $\psi$ varies slowly compared to the Hubble time, and the universe is approximately homogeneous and isotropic at large scales—all consistent with the canonical complex time framework. Extended biquaternionic time formalism (discussed in Appendix~\ref{sec:biquaternion_vs_complex_time}) represents a theoretical exploration for specialized contexts such as inhomogeneous or anisotropic cosmologies with strong primordial gravitational waves.

\section*{M.1 Motivation and Scope}
This appendix consolidates the UBT description of \emph{dark energy} based on the complex-time framework $\tau=t+i\psi$ and the biquaternionic master field $\Theta(q,\tau)$.
We derive an \emph{effective cosmological sector} sourced by the slow phase $\psi$ and show how $\Lambda$CDM is recovered for $\psi\to 0$.
Links to: Appendix F (psychons \& $\psi$-sector dynamics), Appendix J (metric deformations), Appendix K (field propagation in curved backgrounds).

\section*{M.2 UBT Action and Emergent Vacuum Sector}
Consider the effective gravitational action (signature $-,+,+,+$)
\begin{equation}
S_{\rm UBT}=\frac{c^3}{16\pi G}\int d^4x\,\sqrt{-g}\,\Big(R-2\Lambda_0\Big)+
S_\Theta[\Theta,g,\psi]+S_\psi[\psi,g]\,,
\end{equation}
where $\Theta$ carries internal (biquaternionic/spinor) structure and the slow phase $\psi$ is the imaginary part of the complex time $\tau=t+i\psi$.
At long wavelengths, integrating out fast $\Theta$-modes yields an \emph{effective vacuum energy density}
\begin{equation}
\rho_{\rm vac}^{\rm (UBT)}(\psi)\;=\;\rho_{\Lambda 0}\,\big(1+\kappa_\Lambda\,\psi\big)+\frac{1}{2}M_\psi^2\,\psi^2+\frac{\alpha_\psi}{2}(\nabla\psi)^2+\cdots,
\label{eq:rho_vac}
\end{equation}
so that the \emph{effective} cosmological term becomes
\begin{equation}
\Lambda_{\rm eff}(\psi)\;=\;\frac{8\pi G}{c^4}\,\rho_{\rm vac}^{\rm (UBT)}(\psi)\,.
\end{equation}
The coefficients $(\kappa_\Lambda,M_\psi^2,\alpha_\psi)$ are UBT couplings; $\kappa_\Lambda\!\to\!0$ restores $\Lambda$CDM with $\Lambda_{\rm eff}=\Lambda_0$.

\section*{M.3 Homogeneous and Isotropic Cosmology}
For a spatially flat FLRW metric,
\begin{equation}
ds^2=-c^2dt^2+a(t)^2\,d\vb{x}^2,\qquad H\equiv \dot{a}/a,
\end{equation}
the Friedmann equations with UBT dark-energy sector read
\begin{align}
H^2 &= \frac{8\pi G}{3}\left(\rho_m+\rho_r+\rho_{\rm vac}^{\rm (UBT)}(\psi)\right),\\
\dot{H} &= -4\pi G\left(\rho_m+\frac{4}{3}\rho_r+\rho_{\rm vac}^{\rm (UBT)}(\psi)+p_{\rm vac}^{\rm (UBT)}(\psi)\right)/c^2,
\end{align}
with effective equation of state
\begin{equation}
w_{\rm UBT}(\psi)\equiv \frac{p_{\rm vac}^{\rm (UBT)}}{\rho_{\rm vac}^{\rm (UBT)}}\;\approx\;-1+\frac{\alpha_\psi (\nabla\psi)^2 - M_\psi^2\psi^2}{2\,\rho_{\Lambda 0}}+\mathcal{O}(\psi^2).
\end{equation}
For a homogeneous slow phase ($\nabla\psi=0$) we obtain $w_{\rm UBT}\gtrsim -1$ for $M_\psi^2\psi^2\!\ll\!\rho_{\Lambda 0}$; phantom-like $w_{\rm UBT}<-1$ requires parity-odd or higher-derivative mixings (cf.\ Appendix F).

\section*{M.4 Linear Perturbations (Sketch)}
Writing $\psi=\bar{\psi}(t)+\delta\psi(t,\vb{x})$, the scalar sector gains an extra gauge-invariant mode coupled to metric potentials $\Phi,\Psi$.
At sub-horizon scales the effective dark-energy sound speed is
\begin{equation}
c_{s,\rm UBT}^2 \;\simeq\; \frac{\alpha_\psi}{\alpha_\psi + M_\psi^2/k^2}\in(0,1],
\end{equation}
limiting clustering of the vacuum sector; $\alpha_\psi\!\to\!0$ recovers an unclustered $\Lambda$.

\section*{M.5 Recovery of $\Lambda$CDM}
Setting $(\kappa_\Lambda,\alpha_\psi,M_\psi)\to 0$ freezes $\psi$ and yields $\rho_{\rm vac}^{\rm (UBT)}\to \rho_{\Lambda 0}$ with constant $w=-1$ and standard distances, growth, and CMB background. This ensures compatibility with precision cosmology when the UBT phase sector is inactive.

\section*{M.6 Illustrative Hubble Curves (No External Files)}
Below we plot $E(z)\equiv H(z)/H_0$ for three small UBT deformations parameterized by $\kappa_\Lambda\psi\equiv \epsilon\in\{-0.05,0,+0.05\}$, keeping $\Omega_{m0}=0.3$, $\Omega_{\Lambda 0}=0.7$ at $z=0$.
\begin{figure}[h!]
\centering
\begin{tikzpicture}
\begin{axis}[width=0.85\textwidth,height=0.45\textwidth,
xlabel={$z$}, ylabel={$E(z)$}, grid=both,
legend style={at={(0.02,0.98)},anchor=north west,fill=white,draw=none}, ticklabel style={font=\small}, label style={font=\small}]
\addplot+[domain=0:2.0,samples=400,thick] {sqrt(0.3*(1+x)^3 + 0.7*(1+0))};
\addlegendentry{$\epsilon=0$ (LCDM)}
\addplot+[domain=0:2.0,samples=400,thick,dashed] {sqrt(0.3*(1+x)^3 + 0.7*(1+0.05))};
\addlegendentry{$\epsilon=+0.05$}
\addplot+[domain=0:2.0,samples=400,thick,dotted] {sqrt(0.3*(1+x)^3 + 0.7*(1-0.05))};
\addlegendentry{$\epsilon=-0.05$}
\end{axis}
\end{tikzpicture}
\caption{Illustrative expansion histories with a small UBT dark-energy deformation $\epsilon=\kappa_\Lambda\psi$. For $\epsilon\to 0$ we recover $\Lambda$CDM.}
\label{fig:Hz}
\end{figure}

\section*{M.7 Observable Consequences (Qualitative)}
\begin{itemize}
\item Slight shifts in distance--redshift relations $D_L(z), D_A(z)$ and in the derived $H_0$ if $\epsilon\neq 0$ today.
\item Modified ISW effect and low-$\ell$ CMB for time-varying $\bar{\psi}(t)$.
\item Growth rate changes $f(z)\sigma_8$ suppressed by $c_{s,\rm UBT}^2\lesssim 1$; $\epsilon\to 0$ reproduces $\Lambda$CDM.
\end{itemize}

\section*{M.8 Relation to Psychon Sector and Local Tests}
The same $\psi$ that sources $\Lambda_{\rm eff}$ couples to local experiments (Appendix L/N). Constraints from cosmology (global $\bar{\psi}$) and laboratory (local $\psi$ modulations) are complementary; combined fits determine $(\kappa_\Lambda,M_\psi,\alpha_\psi)$ or bound them.

\section*{M.9 Summary}
UBT dark energy arises from a slow phase sector $\psi$ that perturbs the vacuum energy density and hence the effective cosmological constant. The framework recovers $\Lambda$CDM when the phase sector is inactive and predicts small, testable deviations otherwise. This ties cosmic acceleration to the same $\psi$ dynamics appearing in local UBT protocols.
