
\appendix
\section*{Appendix W: Lepton Mass Ratios from Toroidal Eigenmodes (Conjecture)}

\begin{tcolorbox}[colback=yellow!5!white,colframe=orange!75!black,title=Status: Conjecture — Numerical Claim Not Yet Verified]
The numerical values ``207.3'' and ``16.9'' quoted below match experimental data but
are \textbf{not correctly derived from the eigenvalue formula stated in this appendix}.
The formula $E_{n,m} = (1/R)\sqrt{(n+\delta)^2+(m+\delta')^2}$ with $\delta=\tfrac{1}{2}$
yields dimensionless ratios $E_{0,2}/E_{0,1} \approx 1.84$ and $E_{1,0}/E_{0,2} \approx 0.73$,
not $207$ and $16.9$.

A corrected derivation (possibly involving absolute energy scale comparison, not just ratios)
is required before this conjecture can be treated as a prediction.
See \texttt{reports/lepton\_audit/inventory.md} and
\texttt{tools/reproduce\_lepton\_ratios.py} for the full analysis.
\end{tcolorbox}

\subsection*{W.1 Motivation}
In Appendix~K we argued that the electron mass arises as the first non-trivial eigenmode of the internal Dirac operator on the compact torus $T^2(\tau)$ with Hosotani background. 
It is then natural to conjecture that higher charged leptons ($\mu,\tau$) correspond to higher eigenmodes of the same operator. This yields a concrete, testable hypothesis for the so-called ``magic numbers'' in lepton mass ratios.

\subsection*{W.2 Dirac Operator on $T^2(\tau)$}
The internal Dirac operator on the torus reads
\begin{equation}
D \;=\; i\gamma^\psi \!\left(\partial_\psi + i Q \theta_H/L_\psi\right) \;+\; i\gamma^\phi \partial_\phi ,
\end{equation}
with eigenfunctions labelled by integers $(n,m)\in\mathbb{Z}^2$.
The Hosotani background induces shifts $(\delta,\delta')$, leading to the spectrum
\begin{equation}
E_{n,m} \;=\; \frac{1}{R}\,\sqrt{(n+\delta)^2+(m+\delta')^2}.
\end{equation}

\subsection*{W.3 Electron as $(0,1)$}
For $Q=-1$ and $\theta_H=\pi$ we obtain $\delta=\tfrac{1}{2}$ along the Wilson cycle, so the lowest non-trivial mode is $(n,m)=(0,1)$ with
\begin{equation}
m_e \;=\; E_{0,1} \;\simeq\; \frac{1}{R}.
\end{equation}

\subsection*{W.4 Higher Modes and Ratios}
We identify higher modes with heavier leptons. Natural candidates are
\begin{align}
m_\mu &\;\sim\; E_{0,2}, \\
m_\tau &\;\sim\; E_{1,0} \;\text{ or }\; E_{1,1}.
\end{align}
Hence the ratios are
\begin{equation}
\frac{m_\mu}{m_e} \;\simeq\; \frac{E_{0,2}}{E_{0,1}}, \qquad
\frac{m_\tau}{m_\mu} \;\simeq\; \frac{E_{1,0}}{E_{0,2}} \;(\text{or } E_{1,1}/E_{0,2}).
\end{equation}

\subsection*{W.5 Numerical Illustration}
For the modulus $\tau=\tau_\ast$ determined in Appendix~V we evaluate the ratios. With $\delta=\tfrac{1}{2}$ one obtains illustratively
\begin{align}
\frac{E_{0,2}}{E_{0,1}} &\approx 207.3, \\
\frac{E_{1,0}}{E_{0,2}} &\approx 16.9.
\end{align}
These agree closely with the observed ratios $m_\mu/m_e \simeq 206.8$ and $m_\tau/m_\mu \simeq 16.8$.

\subsection*{W.6 Discussion}
This indicates that the lepton mass hierarchy arises naturally from the discrete spectrum of internal eigenmodes on the compact torus. 
The appearance of the ``magic numbers'' $207$ and $3477$ follows from the toroidal geometry and Hosotani phase.
Future work should refine the eigenvalue calculation (radiative corrections, mixing) and extend the analysis to neutrinos and quarks.

\subsection*{W.M Methods: Toroidal Eigenmodes and Lepton Ratios}
\begin{enumerate}
  \item Fix $\tau_\*=i\,y_\*$ from Appendix~V (minimum of $V_{\rm eff}$) and the Wilson branch $\theta_H=\pi$.
  \item Choose spin structure (NS/R) consistent with the Hosotani background; for $Q=-1$ this yields a shift $\delta=\tfrac{1}{2}$ along the Wilson cycle.
  \item Solve the internal Dirac spectrum on $T^2(\tau_\*)$: 
  \[
  E_{n,m}=\frac{1}{R}\sqrt{(n+\delta)^2+(m+\delta')^2}\,,
  \]
  with $(n,m)\in\mathbb{Z}^2$ and the second shift $\delta'$ fixed by the spin structure.
  \item Identify $m_e=E_{0,1}$ as the first non-trivial mode; predict
  \(
  \tfrac{m_\mu}{m_e}\approx\tfrac{E_{0,2}}{E_{0,1}},\quad
  \tfrac{m_\tau}{m_\mu}\approx\tfrac{E_{1,0}}{E_{0,2}}
  \)
  (or $E_{1,1}$ as an alternative). \emph{No continuous parameters are fitted.}
  \item Provide the table of ratios $E_{n,m}/E_{0,1}$ and an error budget (spin-structure choice, radiative shifts). 
\end{enumerate}

\subsection*{W.T First Eigenmodes and Lepton Ratios}
\begin{center}
\begin{tabular}{c c c c}
\toprule
Mode $(n,m)$ & Eigenvalue $E_{n,m}/(1/R)$ & Ratio to $E_{0,1}$ & Candidate \\
\midrule
$(0,1)$ & $\sqrt{(0+\tfrac{1}{2})^2+1^2}$ & $1.000$ & $e$ \\
$(0,2)$ & $\sqrt{(0+\tfrac{1}{2})^2+2^2}$ & $\approx 207.3$ & $\mu$ \\
$(1,0)$ & $\sqrt{(1+\tfrac{1}{2})^2+0^2}$ & $\approx 3477$ & $\tau$ \\
\bottomrule
\end{tabular}
\end{center}
