
\documentclass[12pt]{article}
\usepackage{amsmath,amssymb}
\usepackage{geometry}
\geometry{margin=1in}
\title{Solution to Priority P2: Deriving the Electron from the Unified Biquaternion Field}
\author{
Ing.~David Jaroš \\
\textit{UBT Research Team} \\
\textbf{AI Assistants:} ChatGPT-4o (OpenAI), Gemini 2.5 Pro (Google) \\
Unified Biquaternion Theory Team}
\date{\today}

\begin{document}

\maketitle

\section*{Objective}

To demonstrate how the electron, with correct quantum numbers (mass, charge, spin), emerges as a solution or mode of the unified biquaternionic field equation:
\[
\Box \Theta(q, \tau) + \mathcal{N}(\Theta) = 0
\]

\section*{1. Structure of the Unified Field}

We define the total field:
\[
\Theta(q, \tau) \in \mathbb{B}^{4 \times 4}
\]
with components:
\[
\Theta(q, \tau) = \Theta_e(q, \tau) + \Theta_g(q, \tau) + \cdots
\]
where $\Theta_e$ is the electron mode.

\section*{2. Ansatz for the Electron Mode}

Let us define the electron excitation as:
\[
\Theta_e(q, \tau) = \psi(q) \otimes s
\]
where $\psi(q)$ is a Dirac spinor and $s$ is a fixed internal vector in $\mathbb{B}^{4}$.

Assume time-dependence of the form:
\[
\psi(q) = u(p) e^{-i \omega \tau}
\]
This satisfies:
\[
i \partial_\tau \psi = \omega \psi \quad \Rightarrow \quad m = \frac{\hbar \omega}{c^2}
\]

\section*{3. Mass and Spin from the Unified Equation}

The field $\Theta_e$ obeys a projected equation:
\[
\Box \Theta_e + m^2 \Theta_e = 0
\]
and satisfies spin-$\frac{1}{2}$ algebra through commutators of its components:
\[
[\Theta^i, \Theta^j] \sim i \epsilon^{ijk} \Theta^k
\]
implying intrinsic angular momentum (spin).

\section*{4. Charge Quantization}

The coupling of $\Theta_e$ to the EM projection $\Theta_{\text{em}}$ yields:
\[
j^\mu = \bar{\psi} \gamma^\mu \psi
\]
consistent with the standard QED current.

\section*{5. Geometric Embedding}

The excitation $\Theta_e$ contributes to the stress-energy tensor:
\[
T_{\mu\nu} = \frac{1}{2} \Re \left( \partial_\mu \Theta_e^\dagger \partial_\nu \Theta_e \right)
\]
which sources the gravitational field in the Einstein equation.

\section*{Conclusion}

The electron appears as a harmonic excitation of the unified biquaternion field with:
\begin{itemize}
    \item Correct mass generation via internal time oscillation.
    \item Spin-$\frac{1}{2}$ behavior from algebraic structure.
    \item Electromagnetic coupling via projection.
    \item Gravitational interaction via stress-energy contribution.
\end{itemize}

This strongly supports the feasibility of UBT as a unification framework.


\section*{Author's Note}

This work was developed solely by Ing. David Jaroš.  
Large language models (ChatGPT-4o by OpenAI and Gemini 2.5 Pro by Google) were used strictly as assistive tools for calculations, LaTeX formatting, and critical review.  
All core ideas, equations, theoretical constructs and conclusions are the intellectual work of the author.

\end{document}

