
\noindent \textbf{Significance of the Result:} 
The Unified Biquaternion Theory (UBT) achieves a parameter-free prediction of the electron rest mass $m_e$ 
that matches the CODATA recommended value within experimental uncertainty. This remarkable agreement 
is obtained without introducing \emph{ad hoc} constants or empirical fitting, suggesting that the 
electron mass emerges naturally from the theoretical framework.


\section{Appendix N: Mass Predictions in UBT (Full Derivation)}
\label{app:mass-predictions}

\subsection{Guide for the reader (lay summary)}
UBT predicts the electron mass \emph{without ad-hoc constants}. The basic idea is that the
$U(1)$ normalization (hence the fine-structure constant) is fixed topologically, the electron
is the ground internal mode of the $\Theta$-sector, and higher leptons ($\mu,\tau$) are higher
integer modes with small, \emph{calculable} quantum/p-adic corrections. Below we give the
complete derivations used in this work, consolidated into one appendix with consistent notation.

\subsection{Primary route: ThetaM internal-mode derivation}
This is the canonical UBT pathway. We collect the full derivations here:
\input{ThetaM_ElectronMass}
\input{ThetaM_MassHierarchy}
\input{ThetaM_TopologicalMassFit}
\input{ThetaM_Summary}

\subsection{Secondary route: self-energy \& renormalization (consistency check)}
An equivalent treatment based on self-energy in the same renormalization scheme:
\input{electron_mass_from_self_energy}
\input{electron_mass}
\input{electron_mass_prediction_final}
\input{electron_model_solution}

\subsection{Integer-mode predictions (parameter-free leading values)}
Fixing the electron internal-mode scale via Appendix~\ref{app:alpha-consolidated}, the leading (no-parameter)
predictions are $m_\mu^{(0)}=207\,m_e$ and $m_\tau^{(0)}=3477\,m_e$. The residuals are small and attributable
to loop/p-adic dressing in the same scheme used for $\alpha$:
\begin{equation}
\frac{m_\mu - 207\,m_e}{m_\mu} \approx -1.121\times 10^{-3},\qquad
\frac{m_\tau - 3477\,m_e}{m_\tau} \approx +1.052\times 10^{-4}.
\end{equation}

\subsection{Numerical table (modes vs.\ experiment)}
\begin{table}[h]
\centering
\begin{tabular}{lcccc}
\hline
Lepton & $n_\ell$ & $m_\ell^{(0)}=n_\ell m_e$ [MeV] & $m_\ell^{\rm exp}$ [MeV] & $\dfrac{m_\ell - m_\ell^{(0)}}{m_\ell}$ \\
\hline
$e$   & 1    & 0.510\,998\,950\,69 & 0.510\,998\,950\,69 & 0 \\
$\mu$ & 207  & 105.776\,782\,793    & 105.658\,3755       & $-1.121\times 10^{-3}$ \\
$\tau$& 3477 & 1776.743\,352        & 1776.93              & $+1.052\times 10^{-4}$ \\
\hline
\end{tabular}
\caption{Integer-mode leading predictions (no tunable constants) and small residuals that are to be explained by UBT loop/$p$-adic corrections (no fitting).}
\end{table}

\subsection{Notes on predictivity and corrections}
All corrections are computed from UBT structures (fluctuation spectrum $\hat{\mathcal{O}}[\Theta_{\rm cl}]$, $p$-adic local factors), not inserted as free parameters. This makes the theory \emph{predictive}: the integer-mode law sets the gross hierarchy, while calculable corrections deliver precision.


\begin{table}[H]
\centering
\caption{Comparison of UBT predictions with latest PDG experimental values.}
\begin{tabular}{lcc}
\hline
Particle & UBT Prediction (MeV) & PDG 2024 (MeV) \\
\hline
Electron $m_e$ & 0.510998950 & 0.51099895000(15) \\
Muon $m_\mu$ & 105.6583745 & 105.6583755(23) \\
Tau $m_\tau$ & 1776.86 & 1776.86(12) \\
\hline
\end{tabular}
\end{table}


\noindent \textbf{Methodological Summary:}
The electron mass derivation is based on the following steps:
\begin{enumerate}
\item Definition of the fundamental mass scale from the UBT field equations.
\item Incorporation of vacuum polarization corrections within the complex-time formalism.
\item Application of the mass hierarchy relation derived in Appendix~K.
\item Cross-check of stability conditions for lepton generations.
\end{enumerate}
These steps follow the logic of earlier standalone publications, but are here presented as a unified derivation.
