\appendix
\section*{Appendix 20: Unified Bibliography and Reference Corpus}
\addcontentsline{toc}{section}{Appendix 20: Unified Bibliography and Reference Corpus}

This appendix collects all bibliographic references relevant to the Unified Biquaternion Theory (UBT), its theoretical underpinnings, physical extensions, and experimental hypotheses. It serves as a centralized source for further reading, cross-referencing, and citation.

\subsection*{Primary Works and Foundational Papers}
\begin{itemize}
  \item Jaroš, D. (2025). \textit{A Complex-Time Theory of Consciousness: Drift, Diffusion, and Phase Collapse in Toroidal Cognitive Space}. OSF Preprint.
  \item Jaroš, D. (2025). \textit{Unified Biquaternion Theory: Field Equations, Metric Geometry, and Psychonic Extensions}. GitHub Repository.
  \item Misner, C. W., Thorne, K. S., & Wheeler, J. A. (1973). \textit{Gravitation}. W. H. Freeman.
  \item Penrose, R. (2004). \textit{The Road to Reality}. Jonathan Cape.
  \item Wald, R. M. (1984). \textit{General Relativity}. University of Chicago Press.
\end{itemize}

\subsection*{Mathematical and Topological Background}
\begin{itemize}
  \item Conway, J. B., \textit{Functions of One Complex Variable}. Springer.
  \item Joshi, A. W. (1997). \textit{Elements of Group Theory for Physicists}. New Age International.
  \item Frankel, T. (2011). \textit{The Geometry of Physics: An Introduction}. Cambridge University Press.
  \item Berest, Y., Wilson, G. (2000). \textit{Vector Bundles and the Geometric Langlands Program}.
\end{itemize}

\subsection*{Quantum Field Theory and Extensions}
\begin{itemize}
  \item Peskin, M. E., & Schroeder, D. V. (1995). \textit{An Introduction to Quantum Field Theory}. Addison-Wesley.
  \item Zee, A. (2010). \textit{Quantum Field Theory in a Nutshell}. Princeton University Press.
  \item Weinberg, S. (1995). \textit{The Quantum Theory of Fields (Vol. 1–3)}. Cambridge University Press.
\end{itemize}

\subsection*{Consciousness and Psychophysical Theories}
\begin{itemize}
  \item Hameroff, S., & Penrose, R. (1996). \textit{Orchestrated reduction of quantum coherence in brain microtubules: A model for consciousness}. J. Conscious. Stud.
  \item Tegmark, M. (2000). \textit{Importance of quantum decoherence in brain processes}. Phys. Rev. E.
  \item Tononi, G. (2004). \textit{An information integration theory of consciousness}. BMC Neuroscience.
\end{itemize}

\subsection*{Specialized Topics and Theoretical Extensions}
\begin{itemize}
  \item Barbour, J. (1999). \textit{The End of Time: The Next Revolution in Physics}.
  \item Rovelli, C. (2004). \textit{Quantum Gravity}. Cambridge University Press.
  \item Kiefer, C. (2012). \textit{Quantum Gravity}. Oxford University Press.
  \item Bousso, R. (2002). \textit{The Holographic Principle}. Rev. Mod. Phys.
\end{itemize}

\subsection*{Online Resources and Preprints}
\begin{itemize}
  \item OSF.io (Open Science Framework): \url{https://osf.io}
  \item arXiv.org: \url{https://arxiv.org}
  \item GitHub Repository (Unified Biquaternion Theory): \url{https://github.com/yourprofile/unified-biquaternion-theory}
\end{itemize}

\vspace{1cm}
\noindent Last updated: 2025-06-28
