% ===============================================================
% Appendix K.5 — QCD Scale from Geometric/Holonomy Data (UBT)
% (to be merged into Appendix K after review)
% ===============================================================
\appendix
\section*{Appendix K.5: Deriving \texorpdfstring{$\Lambda_{\mathrm{QCD}}$}{Lambda\_QCD} from the UBT Geometry}

\subsection*{Setup and Scheme}
We evaluate the QCD coupling $g_3(\mu)$ at the geometric scale $\mu=R^{-1}$ fixed by the
Casimir/Hosotani minimum (Appendix~V) and run it down using the standard QCD beta function.
We present results both in a physical scheme and in $\overline{\mathrm{MS}}$ for scheme-independence checks (Appendix~K).

\subsection*{One- and Two-Loop Running}
For $n_f$ active flavors and beta coefficients $b_0,b_1$, the implied confinement scale reads
\begin{equation}
  \Lambda_{\mathrm{QCD}} = \mu \ \exp\!\Big(-\tfrac{8\pi^2}{b_0\,g_3^2(\mu)}\Big)\ [b_0\,g_3^2(\mu)]^{-b_1/b_0^2}\ \Theta_{\rm th} ,
  \label{eq:K5_lambdaQCD}
\end{equation}
where $\Theta_{\rm th}$ encodes discrete threshold factors from holonomies/spin structures (same discrete data as Appendix~V).

\subsection*{Discrete Thresholds from Holonomies}
We compute $\Theta_{\rm th}(\tau,h)$ from KK-mode spectra on $\mathbb{T}^2$;
these are exponentially suppressed for warped/weakly-coupled directions in accordance with roadmap KK phenomenology.

\subsection*{Deliverables}
\begin{enumerate}
  \item Numerical estimate of $\Lambda_{\mathrm{QCD}}$ at $\mu=R^{-1}$ with uncertainty from discrete choices.
  \item Cross-check vs.\ hadronic phenomenology, $n_f$ thresholds, and lattice-informed ranges.
  \item Comparison physical vs.\ $\overline{\mathrm{MS}}$; demonstration of scheme robustness.
\end{enumerate}

\section*{Summary}
This section provides the quantitative bridge between the geometric scale $R$ and the observed $\Lambda_{\mathrm{QCD}}$,
completing the quark-sector consistency with the UBT mechanism.
