% VERSION: v17 Stable Release

\section{Biquaternion Gravity}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Biquaternionic Geometry Lock-In]
\textbf{Gravitational field in UBT is fundamentally biquaternionic.}

The fundamental objects are:
\begin{itemize}
\item \textbf{Metric}: $\mathcal{G}_{\mu\nu}(x) \in \mathbb{B}$, classical metric is $g_{\mu\nu} := \text{Re}(\mathcal{G}_{\mu\nu})$
\item \textbf{Connection}: $\Omega_\mu(x) \in \mathbb{B}$, Christoffel symbols are $\Gamma^\lambda_{\mu\nu} := \text{Re}(\Omega^\lambda_{\mu\nu})$
\item \textbf{Curvature}: $\mathcal{R}_{\mu\nu} \in \mathbb{B}$, Ricci tensor is $R_{\mu\nu} := \text{Re}(\mathcal{R}_{\mu\nu})$
\item \textbf{Stress-energy}: $\mathcal{T}_{\mu\nu} \in \mathbb{B}$, classical stress-energy is $T_{\mu\nu} := \text{Re}(\mathcal{T}_{\mu\nu})$
\end{itemize}

Field equation: $\mathcal{E}_{\mu\nu} = \kappa \mathcal{T}_{\mu\nu}$ (biquaternionic). Einstein equations arise only after $\text{Re}(\cdot)$ projection.
\end{tcolorbox}

This appendix presents the gravitational sector of the Unified Biquaternion Theory (UBT).
It combines the core theoretical framework from the original Biquaternion Gravity appendix
with the detailed analysis from the Quantum Gravity solution files, reformulated for
clarity and coherence.

\subsection{Introduction}

The gravitational field in UBT emerges naturally from the covariant formulation of the
biquaternionic tensor-spinor field equations. UBT generalizes Einstein's General Relativity (GR) 
by embedding the metric tensor in a biquaternionic field $\Theta(q,\tau)$ defined over 
complex time $\tau = t + i\psi$. 

\textbf{Note on canonical complex time}: For the gravitational sector, canonical complex time $\tau = t + i\psi$ (AXIOM B) is the standard formulation in UBT. Gravity is effectively an Abelian field in this formulation, satisfying $[\Theta_i, \Theta_j] \approx 0$, consistent with the canonical complex time framework. Extended biquaternionic time formalism (with quaternionic vector components) represents a theoretical exploration for specialized contexts such as extreme rotating spacetimes or coupled non-Abelian gauge-gravity systems.

The real part of this field corresponds to the classical 
metric tensor, ensuring that \textbf{GR is fully contained within UBT as a special case}.

In the real-valued limit (or equivalently, when the imaginary time component $\psi \to 0$), 
the biquaternionic field equations reduce exactly to Einstein's field equations, including 
all cases where the scalar curvature $R \neq 0$. The extended biquaternionic structure 
introduces additional degrees of freedom that represent phase-like and nonlocal components 
of spacetime, which have no signature in classical observations but may be relevant for 
dark sector physics and quantum gravitational effects.

In this formulation, spacetime is represented by a projection from a higher-dimensional
complex manifold, and curvature is encoded in the covariant derivatives of the
biquaternion field $\Theta(q,\tau)$. The gravitational interaction is therefore not an
independent postulate, but a manifestation of the underlying field geometry.

\subsection{Core Equations}

The classical line element (real projection):
\begin{equation}
  ds^2 = g_{\mu\nu} \, dx^\mu dx^\nu ,
\end{equation}
where the metric tensor $g_{\mu\nu}$ is obtained as the real projection of the biquaternionic metric:
\begin{equation}
  g_{\mu\nu} := \text{Re}(\mathcal{G}_{\mu\nu}) = \Re\left[ \frac{\partial_\mu \Theta \cdot \partial_\nu \Theta^\dagger}{\mathcal{N}} \right],
\end{equation}
with $\mathcal{N}$ a normalisation factor ensuring the correct signature.

The classical Einstein tensor (real projection of biquaternionic Einstein tensor):
\begin{equation}
  G_{\mu\nu} := \text{Re}(\mathcal{E}_{\mu\nu}) = R_{\mu\nu} - \frac{1}{2} g_{\mu\nu} R ,
\end{equation}
where curvature tensors $R_{\mu\nu}$ and $R$ propagate from the biquaternionic connection
coefficients via real projection: $\Gamma^\rho_{\mu\nu} := \text{Re}(\Omega^\rho_{\mu\nu})$.

The gravitational field equations (GR limit after real projection):
\begin{equation}
  G_{\mu\nu} = 8\pi G \, T_{\mu\nu}[\Theta] .
\end{equation}
where $T_{\mu\nu} := \text{Re}(\mathcal{T}_{\mu\nu})$ is the real projection of the biquaternionic stress-energy tensor.

\subsection{analysis Summary}

The original analysis proceeds by first defining the biquaternionic connection
compatible with the metric propose from $\Theta(q,\tau)$. The connection coefficients
are computed as:
\begin{equation}
  \Gamma^\rho_{\mu\nu} =
  \frac{1}{2} g^{\rho\sigma} \left( \partial_\mu g_{\nu\sigma}
  + \partial_\nu g_{\mu\sigma}
  - \partial_\sigma g_{\mu\nu} \right),
\end{equation}
where $g_{\mu\nu}$ is substituted from the field definition above.

The curvature tensor $R^\rho_{\ \sigma\mu\nu}$ is then obtained from:
\begin{equation}
  R^\rho_{\ \sigma\mu\nu} =
  \partial_\mu \Gamma^\rho_{\nu\sigma} -
  \partial_\nu \Gamma^\rho_{\mu\sigma} +
  \Gamma^\rho_{\mu\lambda} \Gamma^\lambda_{\nu\sigma} -
  \Gamma^\rho_{\nu\lambda} \Gamma^\lambda_{\mu\sigma} .
\end{equation}

Contracting appropriately yields $R_{\mu\nu}$ and the scalar curvature $R$.

\paragraph{Phase Curvature and Invisibility.}
The biquaternionic field $\Theta(q,\tau)$ has both real and imaginary components. While the 
real part $g_{\mu\nu} = \Re[\Theta_{\mu\nu}]$ corresponds to the classical metric tensor that 
couples to ordinary matter and radiation, the imaginary components $\Im[\Theta_{\mu\nu}]$ represent 
what we call \textbf{phase curvature}.

These phase curvature components satisfy their own field equations within the biquaternionic 
algebra and can carry energy-momentum in configurations where:
\begin{equation}
\Re[G_{\mu\nu}] = 0, \quad \text{but} \quad \Im[G_{\mu\nu}] \neq 0.
\end{equation}

Such configurations are mathematically consistent solutions but remain \textbf{invisible} to 
classical observations because ordinary matter and electromagnetic radiation couple only to the 
real-valued metric $g_{\mu\nu}$. They have no real-valued curvature signature ($\Re[R_{\mu\nu}] = 0$) 
yet possess nonlocal energy structure in the imaginary components. This invisibility is a 
mathematical property arising from the complex structure of the biquaternionic algebra, not a 
contradiction with General Relativity.

In the quantum gravity extension, fluctuations of the field $\Theta$ are quantised,
leading to corrections to the classical curvature in the form of effective stress-energy
terms arising from vacuum polarisation effects. The semiclassical approximation shows that
these corrections become significant near the Planck scale, modifying the black hole
horizon structure and potentially allowing for stable micro-horizon configurations.

\subsection{Summary}

Biquaternion Gravity generalizes Einstein's General Relativity by embedding it within a
richer biquaternionic algebraic framework. In the real-valued limit, UBT reproduces 
Einstein's field equations exactly, ensuring full compatibility with all experimental 
confirmations of GR—from the perihelion precession of Mercury to gravitational wave 
detection. The extended structure unifies gravity with other interactions at the
geometric level and introduces additional degrees of freedom corresponding to phase 
curvature and nonlocal energy configurations.

These imaginary components of the biquaternionic metric may represent dark sector 
phenomena or quantum gravitational corrections, but they remain invisible to classical 
matter and electromagnetic radiation that couple only to the real metric $g_{\mu\nu}$. 
The connection to Quantum Gravity arises from treating $\Theta$ as a quantum field, 
where gravitational effects emerge from its covariant structure. 

This approach may suggest possible deviations from classical GR at very small scales 
(near the Planck scale), while reducing exactly to Einstein's equations in all 
macroscopic regimes where GR has been tested. UBT does not contradict or replace 
General Relativity; it extends and embeds it within a broader mathematical framework. 
For detailed derivations of the GR recovery, see Appendix~R.
