\documentclass[11pt,a4paper]{article}
\usepackage{slashed}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{calc,decorations.pathmorphing,decorations.markings,positioning,arrows.meta,plotmarks}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\geometry{margin=1in}

\title{Unified Biquaternion Theory\\[0.5em]
\large Core Framework: Gravity, Electromagnetism, and Gauge Symmetries}
\author{Ing. David Jaroš}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
We present the Unified Biquaternion Theory (UBT), a theoretical framework that generalizes and embeds Einstein's General Relativity within a biquaternionic field structure defined over complex time $\tau = t + i\psi$. The theory recovers the Einstein--Maxwell--Dirac system through a variational action and demonstrates correct classical limits. In the real-valued limit, UBT exactly reproduces Einstein's field equations for all curvature regimes, including cases where the Ricci scalar $R \neq 0$. We derive Standard Model gauge symmetries $\mathrm{SU}(3) \times \mathrm{SU}(2) \times \mathrm{U}(1)$ as emergent geometric structures from the internal phase space of the fundamental biquaternionic field $\Theta(q,\tau)$. The biquaternionic degrees of freedom extend classical spacetime with phase-like components that remain invisible to ordinary observations but may be relevant for quantum gravitational corrections and dark sector physics.
\end{abstract}

\tableofcontents
\newpage

\section*{Core Scope and Claims}
\addcontentsline{toc}{section}{Core Scope and Claims}

This manuscript presents a biquaternion formulation that \textbf{generalizes and embeds Einstein's General Relativity} while recovering the Einstein--Maxwell--Dirac system under a standard variational action. The theory demonstrates the correct limits (Minkowski and weak-field), and \textbf{in the real-valued limit, exactly reproduces Einstein's field equations} for all curvature regimes, including cases where the Ricci scalar $R \neq 0$.

UBT extends GR by introducing biquaternionic degrees of freedom that represent phase-like and nonlocal components of spacetime. These additional components remain invisible to classical observations because ordinary matter couples only to the real metric $g_{\mu\nu} = \mathrm{Re}(\Theta^\dagger\Theta)$. The imaginary components may be relevant for:
\begin{itemize}
\item Dark sector physics (dark matter and dark energy)
\item Quantum gravitational corrections at Planck scales
\item Nonlocal phase correlations in quantum field theory
\end{itemize}

\subsection*{Important Clarifications}

\paragraph{Fine-structure constant.}
\textbf{We do not claim an ab-initio derivation of the fine-structure constant}~$\alpha$. In this core manuscript, $\alpha(\mu)$ is treated as an empirical input parameter consistent with QED running coupling measurements. Speculative approaches linking $\alpha$ to topological quantization or $p$-adic structures are excluded from the core claims.

\paragraph{Consciousness and $\psi$.}
Any interpretations connecting the auxiliary phase coordinate $\psi$ (the imaginary part of complex time $\tau = t + i\psi$) to consciousness are considered \emph{interpretive and speculative} and are \textbf{not part of the core results}. In this manuscript, $\psi$ is treated purely as a mathematical phase variable with potential information-theoretic significance.

\paragraph{Testable predictions.}
Quantitative, testable predictions are stated explicitly with their assumptions and orders of magnitude. All claims are formulated to be falsifiable through experimental tests or observational data.

\appendix
\tableofcontents

% ---- CORE APPENDICES ----

\input{appendix_A_biquaternion_gravity_consolidated}
\input{appendix_R_GR_equivalence}
\input{appendix_C_electromagnetism_gauge_consolidated}
\input{appendix_D_qed_consolidated}
\input{appendix_E_SM_QCD_embedding}
\input{appendix_G_internal_color_symmetry}
\input{appendix_K_maxwell_curved_space}

% ---- MATHEMATICAL FOUNDATIONS ----

\input{appendix_P1_biquaternion_inner_product}
\input{appendix_P3_hilbert_space}
\input{appendix_P5_integration_measure}

% ---- TESTABILITY ----

\input{appendix_W_testable_predictions}

% ---- BIBLIOGRAPHY ----

\input{appendix_Z_bibliography}

\section*{Acknowledgments}
\addcontentsline{toc}{section}{Acknowledgments}

The author acknowledges the use of AI assistants (ChatGPT, Claude, Gemini) as computational and editorial tools in the development of this theoretical framework. All theoretical concepts, derivations, and scientific claims originate from the author.

\section*{License}
\addcontentsline{toc}{section}{License}

This work is licensed under a Creative Commons Attribution 4.0 International License (CC BY 4.0). You are free to share and adapt this work with appropriate attribution.

\end{document}
