\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{slashed}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{tikz-3dplot}
\usetikzlibrary{calc,decorations.pathmorphing,decorations.markings,positioning,arrows.meta}
\pgfplotsset{compat=1.18}
\geometry{a4paper, margin=1in}

% Define theorem environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\title{Unified Biquaternion Theory 2.0 (Consolidated)}
\author{UBT Team}
\date{\today}

% Define flag for included files to skip their preambles
\def\INCLUDEMODE{}

\begin{document}
\maketitle
\tableofcontents

\section*{CORE Scope and Claims}
This CORE manuscript presents a biquaternion formulation that \textbf{generalizes and embeds Einstein's General Relativity} while recovering the Einstein--Maxwell--Dirac system under a standard variational action. The theory demonstrates the correct limits (Minkowski and weak-field), and \textbf{in the real-valued limit, exactly reproduces Einstein's field equations} for all curvature regimes, including cases where the Ricci scalar $R \neq 0$.

UBT extends GR by introducing biquaternionic degrees of freedom that represent phase-like and nonlocal components of spacetime. These additional components remain invisible to classical observations but may be relevant for dark sector physics and quantum gravitational corrections. \textbf{We do not claim an ab-initio derivation of the fine-structure constant}~$\alpha$; in the CORE track $\alpha(\mu)$ is treated as an empirical input consistent with QED running. Any links between the auxiliary phase coordinate~$\psi$ and consciousness are considered \emph{interpretive and speculative} and are not part of the CORE results. Quantitative, testable predictions are stated with their assumptions and orders of magnitude.

\subsection*{Contextual Assessment of Originality}

While UBT introduces original constructs—complex time, biquaternionic unification, and consciousness–field coupling—it builds upon several well-established mathematical frameworks including quaternionic and octonionic algebras, gauge-group embeddings, and holographic correspondences. 

The unified derivation of the fine-structure constant ($\alpha$) represents an innovative synthesis but currently depends on an empirically fitted renormalization factor. 

Acknowledging previous quaternionic unification efforts (e.g., Lanczos 1929 \cite{Lanczos1929}, Gürsey 1956 \cite{Gursey1956}, Finkelstein 1962 \cite{Finkelstein1962}, De Leo 2000 \cite{DeLeo2000}, and recent 2022 arXiv complex-quaternion Dirac models \cite{ComplexQuaternion2022a,ComplexQuaternion2022b}) strengthens the theory's credibility by clarifying its novel contributions: the incorporation of complex time, drift–diffusion of consciousness, and p-adic multiverse sectors. See Appendix~\ref{sec:originality} for detailed comparative analysis.

\subsection*{Rigorous vs. Speculative Content}

UBT distinguishes between:
\begin{itemize}
\item \textbf{Rigorously derived results}: General Relativity compatibility, gauge field structure, Einstein-Maxwell-Dirac system recovery, dimensional reduction to 4D spacetime
\item \textbf{Conceptual/exploratory frameworks}: Consciousness integration via psychon dynamics, p-adic multiverse interpretation, closed timelike curves, ab initio $\alpha$ derivation
\end{itemize}

Speculative sections are clearly marked throughout this document (see Appendix~\ref{app:speculative_notes} for comprehensive list).

\subsection*{Acknowledgement of Prior Work}

UBT stands on the shoulders of a rich tradition of quaternionic physics, extending from Lanczos's pioneering application of quaternions to electromagnetism (1929) through Gürsey's conform-invariant spinor equations (1956), Finkelstein et al.'s rigorous foundations (1962), Adler's comprehensive quaternionic quantum field theory (1995), and De Leo's quaternionic electroweak theory (2000). Recent work on complex-quaternionic Dirac models (2022) has explored similar algebraic structures in parallel.

UBT's originality lies not in the use of quaternions per se, but in the specific synthesis of: (1) complex time as geometric extension, (2) theta-function attractors, (3) emergent gauge structure from phase topology, and (4) speculative integration of consciousness. This positions UBT as a continuation and extension of historical quaternionic unification programs rather than a replacement.


\appendix

% ---- CONTEXT AND ORIGINALITY ----
\input{appendix_originality_context}

% ---- MATHEMATICAL FOUNDATIONS (Priority 1) ----
\input{appendix_P1_biquaternion_inner_product}
% NOTE: appendix_P2_multiverse_projection.tex moved to speculative_extensions/appendices/ (Nov 2025)
\input{appendix_P3_hilbert_space}
\input{appendix_P4_alpha_status}
\input{appendix_P5_integration_measure}
\input{appendix_P6_lorentz_in_HC}
\input{appendix_RH_riemann_zeta_connection}

% ---- TESTABILITY AND TSVF (Priority 2 & 4) ----
\input{appendix_W_testable_predictions}
\input{appendix_CERN_BSM_predictions}
\input{appendix_X_TSVF_integration}

% ---- CORE ----
\input{appendix_A_biquaternion_gravity_consolidated}
\input{appendix_QG_quantum_gravity_unification}
\input{appendix_R_GR_equivalence}
\input{appendix_N_holographic_verlinde_desitter}
\input{appendix_C_electromagnetism_gauge_consolidated}
\input{appendix_D_qed_consolidated}
\input{appendix_E_SM_QCD_embedding}
\input{appendix_K_maxwell_curved_space}
\input{appendix_G_internal_color_symmetry}
\input{appendix_G_hamiltonian_theta_exponent}  % NEW: Hamiltonian-in-exponent formulation (2025)
\input{appendix_G6_neutrino_mass_biquaternionic_time}  % NEW: Neutrino mass from biquaternionic time (2025)

% ---- PHILOSOPHICAL AND STRUCTURAL COHERENCE ----
% NOTE: appendix_I_philosophical_coherence.tex moved to speculative_extensions/appendices/ (Nov 2025)

% ---- SPECULATIVE / WIP ----
\input{appendix_U_dark_matter_unified_padic}
\input{appendix_K2_fundamental_constants_consolidated}
\input{appendix_O_padic_overview}
\input{appendix_ALPHA_padic_derivation}

% ---- GLOSSARY AND BIBLIOGRAPHY ----
\input{appendix_glossary_symbols}
% NOTE: appendix_S_speculative_notes.tex moved to speculative_extensions/appendices/ (Nov 2025)
\input{appendix_Z_bibliography}

\section*{License}
This work is licensed under a Creative Commons Attribution 4.0 International License (CC BY 4.0).

\end{document}
