% © 2025 Ing. David Jaroš — CC BY-NC-ND 4.0
%
% This work is licensed under a Creative Commons Attribution-NonCommercial-NoDerivatives 
% 4.0 International License (CC BY-NC-ND 4.0).
%
% License History: Earlier drafts (up to v0.3) were released under CC BY 4.0. 
% From v0.4 onward, all material is released under CC BY-NC-ND 4.0 to protect 
% the integrity of the theoretical work during ongoing academic development.
%
% See LICENSE.md for full license text.

\documentclass[11pt,a4paper]{article}
\usepackage{slashed}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{calc,decorations.pathmorphing,decorations.markings,positioning,arrows.meta,plotmarks}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\geometry{margin=1in}

% Define theorem environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\title{Unified Biquaternion Theory (Consolidation) --- CORE Manuscript}
\author{UBT Team}
\date{\today}

% Define flag for included files to skip their preambles
\def\INCLUDEMODE{}

\begin{document}
\maketitle

% License Notice - Visible in PDF
\noindent
\textbf{License:} © 2025 Ing. David Jaroš. This work is licensed under a Creative Commons Attribution-NonCommercial-NoDerivatives 4.0 International License (CC BY-NC-ND 4.0). Earlier drafts (up to v0.3) were released under CC BY 4.0. From v0.4 onward, all material is released under CC BY-NC-ND 4.0 to protect the integrity of the theoretical work during ongoing academic development. See \url{https://creativecommons.org/licenses/by-nc-nd/4.0/} for details.

\vspace{1em}

\noindent\textbf{Lock-in statement:} Throughout this work, all geometric and dynamical structures are defined at the level of biquaternionic fields. Any real-valued spacetime metric, curvature, or stress--energy tensor represents a Hermitian projection corresponding to an observer-restricted sector. No physical conclusion should be interpreted at the level of the real projection alone.

\vspace{0.5em}

\noindent\textbf{Future-proofing rule:} Any future extension of UBT must (i) define new dynamics at the biquaternionic level, (ii) state explicitly how the GR sector is obtained as $\Re(\,\cdot\,)$, and (iii) avoid introducing classical GR objects (metrics, Levi--Civita symbols, or stress--energy tensors) as axioms without such a projection. This applies to all new appendices, phenomenological discussions, and experimental proposals.

\vspace{1em}

\appendix
\tableofcontents

\section*{CORE Scope and Claims}
This CORE manuscript presents a biquaternion formulation that \textbf{generalizes and embeds Einstein's General Relativity} while recovering the Einstein--Maxwell--Dirac system under a standard variational action. The theory demonstrates the correct limits (Minkowski and weak-field), and \textbf{in the real-valued limit, exactly reproduces Einstein's field equations} for all curvature regimes, including cases where the Ricci scalar $R \neq 0$.

UBT extends GR by introducing biquaternionic degrees of freedom that represent phase-like and nonlocal components of spacetime. These additional components remain invisible to classical observations but may be relevant for dark sector physics and quantum gravitational corrections. \textbf{We do not claim an ab-initio derivation of the fine-structure constant}~$\alpha$; in the CORE track $\alpha(\mu)$ is treated as an empirical input consistent with QED running. Any links between the auxiliary phase coordinate~$\psi$ and consciousness are considered \emph{interpretive and speculative} and are not part of the CORE results. Quantitative, testable predictions are stated with their assumptions and orders of magnitude.

\subsection*{Contextual Assessment of Originality}

While UBT introduces original constructs—complex time, biquaternionic unification, and consciousness–field coupling—it builds upon several well-established mathematical frameworks including quaternionic and octonionic algebras, gauge-group embeddings, and holographic correspondences. 

The unified derivation of the fine-structure constant ($\alpha$) represents an innovative synthesis but currently depends on an empirically fitted renormalization factor. 

Acknowledging previous quaternionic unification efforts (e.g., Lanczos 1929, Gürsey 1956, Finkelstein 1962, De Leo 2000, and recent 2022 arXiv complex-quaternion Dirac models) strengthens the theory's credibility by clarifying its novel contributions: the incorporation of complex time, drift–diffusion of consciousness, and p-adic multiverse sectors.


% ---- CORE APPENDICES (auto-included) ----
% NOTE: do not include speculative/WIP files here.

\input{appendix_originality_context}
\input{appendix_A_biquaternion_gravity_consolidated}
\def\INCLUDEMODE{}  % Signal to appendices that we're including them, not compiling standalone
\input{appendix_QG_quantum_gravity_unification}
\input{appendix_R_GR_equivalence}
\input{appendix_C_electromagnetism_gauge_consolidated}
\input{appendix_D_qed_consolidated}
\input{appendix_E_SM_QCD_embedding}
\input{appendix_K_maxwell_curved_space}
\input{appendix_G_internal_color_symmetry}
\input{appendix_G_hamiltonian_theta_exponent}  % NEW: Hamiltonian-in-exponent formulation (2025)
\input{appendix_G5_biquaternionic_fokker_planck}  % NEW: Biquaternionic Fokker-Planck equation (2025)
% NOTE: appendix_I_philosophical_coherence.tex moved to speculative_extensions/appendices/ (Nov 2025)
\input{appendix_glossary_symbols}
\input{appendix_Z_bibliography}

\input{appendix_alpha_statement}

\section*{License}

© 2025 Ing. David Jaroš — CC BY-NC-ND 4.0

This work is licensed under a Creative Commons Attribution-NonCommercial-NoDerivatives 4.0 International License (CC BY-NC-ND 4.0).

\textbf{License History:} Earlier drafts (up to v0.3) were released under CC BY 4.0. From v0.4 onward, all material is released under CC BY-NC-ND 4.0 to protect the integrity of the theoretical work during ongoing academic development.

\end{document}
