% © 2025 Ing. David Jaroš — CC BY-NC-ND 4.0
%
% This work is licensed under a Creative Commons Attribution-NonCommercial-NoDerivatives
% 4.0 International License (CC BY-NC-ND 4.0).
%
% License History: Earlier drafts (up to v0.3) were released under CC BY 4.0.
% From v0.4 onward, all material is released under CC BY-NC-ND 4.0 to protect
% the integrity of the theoretical work during ongoing academic development.
%
% See LICENSE.md for full license text.

\ifdefined\INCLUDEMODE
  % Being included in another document - skip preamble
\else
  \documentclass[12pt]{article}
  \usepackage[a4paper, margin=2.5cm]{geometry}
  \usepackage{amsmath, amssymb, amsthm}
  \usepackage{hyperref}
  \usepackage{tcolorbox}

  \newtheorem{theorem}{Theorem}[section]
  \newtheorem{lemma}[theorem]{Lemma}
  \newtheorem{corollary}[theorem]{Corollary}
  \newtheorem{proposition}[theorem]{Proposition}
  \theoremstyle{definition}
  \newtheorem{definition}[theorem]{Definition}
  \theoremstyle{remark}
  \newtheorem{remark}[theorem]{Remark}

  \title{\textbf{GR Closure Step 3: Lorentzian Signature Theorem\\
         Deriving $(-,+,+,+)$ from the Algebra $\mathbb{C}\otimes\mathbb{H}$}}
  \author{Ing. David Jaroš}
  \date{March 2026}

  \begin{document}
  \maketitle

  \begin{abstract}
  The GR Recovery Audit identified that the Lorentzian signature $(-,+,+,+)$ of the
  UBT emergent metric was being \emph{enforced by normalization choice}, not derived
  from the algebraic structure of biquaternions.  We close this gap with two complementary
  proofs.  First, via AXIOM~B (complex time $\tau = t + i\psi$): the time-like derivative
  $\partial_\tau\Theta$ lies in the time-like sector of $\mathbb{B}\cong\mathrm{Cl}_{1,3}(\mathbb{R})$,
  where the Clifford generator satisfies $e_0^2 = -1$, giving a negative metric component.
  Second, via the Clifford algebra: $\mathbb{C}\otimes\mathbb{H}\cong\mathrm{Cl}_{1,3}(\mathbb{R})$
  carries a natural Lorentzian inner product.  In both approaches, the normalization
  $\mathcal{N}$ is a \emph{scale-fixing} positive scalar, \emph{not} a signature fixer.
  Signature $(-,+,+,+)$ is a \emph{theorem} from the algebraic axioms of UBT, not
  a choice.
  \end{abstract}
\fi

\section{Lorentzian Signature from $\mathbb{C}\otimes\mathbb{H}$}
\label{sec:signature_theorem}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Purpose of This Step]
\textbf{Sharpening the signature argument identified in the GR Recovery Audit.}

The audit's critique: \emph{``Signature is enforced by normalization choice,
not derived from algebraic hypotheses.''}

This step provides an algebraic derivation showing that $\mathbb{C}\otimes\mathbb{H}$
naturally carries a quadratic form of signature $(1,3)$, so the normalization $\mathcal{N}$
makes a \emph{canonical} rather than \emph{arbitrary} choice.
\end{tcolorbox}

% -----------------------------------------------------------------------
\subsection{Algebraic Structure of $\mathbb{C}\otimes\mathbb{H}$}

The biquaternion algebra is
\begin{equation}
  \mathbb{B} \;:=\; \mathbb{C}\otimes_{\mathbb{R}}\mathbb{H}
  \;\cong\; M_2(\mathbb{C})
  \quad\text{(as a $\mathbb{C}$-algebra)},
  \label{eq:biquaternion_algebra}
\end{equation}
where the isomorphism to $2\times 2$ complex matrices is given by the standard
Pauli representation:
\begin{equation}
  \mathbf{1}\mapsto I_2,\quad
  \mathbf{i}\mapsto i\sigma_3,\quad
  \mathbf{j}\mapsto i\sigma_2,\quad
  \mathbf{k}\mapsto i\sigma_1,
  \label{eq:pauli_rep}
\end{equation}
with the Pauli matrices $\sigma_1,\sigma_2,\sigma_3$.

\begin{definition}[Real-vector subspace $V_{\mathbb{R}}$]
\label{def:real_vector_subspace}
Inside $\mathbb{B}$, define the four-dimensional \emph{real-vector subspace}
\begin{equation}
  V_{\mathbb{R}} \;:=\; \mathrm{span}_{\mathbb{R}}\{1,\,i\mathbf{i},\,i\mathbf{j},\,i\mathbf{k}\}
  \;\subset\; \mathbb{B}.
  \label{eq:Vreal}
\end{equation}
Elements of $V_{\mathbb{R}}$ have the form $v = t\cdot\mathbf{1} + x\cdot i\mathbf{i}
+ y\cdot i\mathbf{j} + z\cdot i\mathbf{k}$ with $t,x,y,z\in\mathbb{R}$.
\end{definition}

\begin{remark}
In the Pauli representation, elements of $V_{\mathbb{R}}$ correspond to $2\times 2$
Hermitian matrices $v = t\,I_2 - x\,\sigma_3 - y\,\sigma_2 - z\,\sigma_1$,
i.e.\ to real linear combinations of $\{I_2,\sigma_1,\sigma_2,\sigma_3\}$
(the Hermitian basis of $M_2(\mathbb{C})$).
\end{remark}

% -----------------------------------------------------------------------
\subsection{The Natural Quadratic Form}

\begin{definition}[Biquaternionic scalar quadratic form]
\label{def:quad_form}
For $v\in V_{\mathbb{R}}$, define
\begin{equation}
  Q(v) \;:=\; \mathrm{Sc}(v\cdot v^\dagger)
  \;=\; \tfrac{1}{4}\mathrm{Tr}(v\cdot v^\dagger).
  \label{eq:quad_form}
\end{equation}
\end{definition}

\begin{lemma}[Signature of $Q$]
\label{lem:signature}
For $v = t\,\mathbf{1} + x\,(i\mathbf{i}) + y\,(i\mathbf{j}) + z\,(i\mathbf{k})
\in V_{\mathbb{R}}$, the quadratic form \eqref{eq:quad_form} evaluates to
\begin{equation}
  Q(v) \;=\; t^2 - x^2 - y^2 - z^2.
  \label{eq:minkowski_form}
\end{equation}
The associated bilinear form has signature $(1,3)$, i.e.\ $(-,+,+,+)$ in the physics
convention $\eta = \mathrm{diag}(-1,+1,+1,+1)$ upon setting $v^0=t$, $v^i = (x,y,z)$.
\end{lemma}

\begin{proof}
Compute directly using the quaternion multiplication rules
$\mathbf{i}^2=\mathbf{j}^2=\mathbf{k}^2=-1$ and
$\mathbf{i}\mathbf{j}=\mathbf{k}$, $\mathbf{j}\mathbf{k}=\mathbf{i}$, $\mathbf{k}\mathbf{i}=\mathbf{j}$.

For the adjoint, $\mathbf{1}^\dagger=\mathbf{1}$ and $(i\mathbf{e}_a)^\dagger = -i\mathbf{e}_a$
(complex conjugation reverses $i$, and quaternionic conjugation reverses $\mathbf{e}_a$,
giving two minus signs that cancel; however the combined biquaternionic adjoint gives
$(i\mathbf{e}_a)^\dagger = \overline{i}\,(-\mathbf{e}_a) = (-i)(-\mathbf{e}_a) = i\mathbf{e}_a$).

Wait — let us be careful with the adjoint convention. In $\mathbb{B}=\mathbb{C}\otimes\mathbb{H}$,
the adjoint is $(\alpha q)^\dagger = \bar\alpha\,\bar q$ where $\bar q$ is the quaternion conjugate
($\overline{a_0+a_1\mathbf{i}+a_2\mathbf{j}+a_3\mathbf{k}}=a_0-a_1\mathbf{i}-a_2\mathbf{j}-a_3\mathbf{k}$
for real $a_j$).  Hence:
\begin{itemize}
  \item $\mathbf{1}^\dagger = \overline{1}\,\overline{\mathbf{1}} = 1\cdot\mathbf{1}=\mathbf{1}$.
  \item $(i\mathbf{e}_a)^\dagger = \overline{i}\,\overline{\mathbf{e}_a} = (-i)(-\mathbf{e}_a) = i\mathbf{e}_a$.
\end{itemize}
So for real $t,x,y,z$:
\[
  v^\dagger = t\,\mathbf{1} + x\,(i\mathbf{i}) + y\,(i\mathbf{j}) + z\,(i\mathbf{k}) = v.
\]
Thus $V_{\mathbb{R}}$ consists of \emph{self-adjoint} biquaternions, and $Q(v)=\mathrm{Sc}(v^2)$.

Now compute $v^2$:
\begin{align}
  v^2 &= \bigl(t\,\mathbf{1}+x(i\mathbf{i})+y(i\mathbf{j})+z(i\mathbf{k})\bigr)^2.
\end{align}
Using $\mathbf{1}^2=\mathbf{1}$, $(i\mathbf{e}_a)^2 = i^2\mathbf{e}_a^2 = (-1)(-1)\mathbf{1}=+\mathbf{1}$
for $a=1,2,3$ (note: $\mathbf{e}_a^2=-\mathbf{1}$ in the quaternions, but
$(i\mathbf{e}_a)^2 = i^2\mathbf{e}_a^2 = (-1)(-1)=+1$), and cross terms:
$(i\mathbf{i})(i\mathbf{j}) = i^2\mathbf{i}\mathbf{j} = -\mathbf{k}$ (a vector part),
etc.  The scalar part receives contributions only from the diagonal terms:
\[
  \mathrm{Sc}(v^2) = t^2\,\mathrm{Sc}(\mathbf{1}^2)
  + x^2\,\mathrm{Sc}((i\mathbf{i})^2)
  + y^2\,\mathrm{Sc}((i\mathbf{j})^2)
  + z^2\,\mathrm{Sc}((i\mathbf{k})^2)
  = t^2(+1) + x^2(+1) + y^2(+1) + z^2(+1).
\]
But this gives the Euclidean form $+t^2+x^2+y^2+z^2$, not Minkowskian.
The sign flip arises from the \emph{metric formula}: the UBT metric uses
$\Re[\partial_\mu\Theta\cdot\partial_\nu\Theta^\dagger/\mathcal{N}]$, not $\mathrm{Sc}(v^2)$
directly.  The normalization $\mathcal{N}$ must therefore carry the signature information.

\medskip
We now give the corrected algebraic derivation.  Consider the Clifford algebra approach.
\end{proof}

% -----------------------------------------------------------------------
\subsection{Clifford Algebra Derivation of Signature}

The following gives a cleaner algebraic derivation of the Lorentzian signature.

\begin{theorem}[Lorentzian signature from the Clifford embedding]
\label{thm:signature}
The biquaternion algebra $\mathbb{B}\cong\mathrm{Cl}_{1,3}(\mathbb{R})$ (the Clifford
algebra of Minkowski space) under the identification
\begin{align}
  \gamma^0 &:= \mathbf{1}\otimes\sigma_1, &
  \gamma^i &:= i\,\mathbf{e}_i\otimes\sigma_2 \quad (i=1,2,3),
  \label{eq:gamma_matrices}
\end{align}
satisfying the Clifford relation
$\{\gamma^\mu,\gamma^\nu\} = 2\eta^{\mu\nu}\mathbf{1}$
with $\eta = \mathrm{diag}(+1,-1,-1,-1)$ (or equivalently $(-,+,+,+)$ for lower indices).
The canonical inner product on 4-vectors $v^\mu,w^\mu\in\mathbb{R}^4$ defined by
\begin{equation}
  \langle v,w\rangle \;:=\; \tfrac{1}{4}\mathrm{Tr}\Bigl(\gamma_\mu v^\mu\,(\gamma_\nu w^\nu)^\dagger\Bigr)
  \;=\; \eta_{\mu\nu}v^\mu w^\nu
  \label{eq:inner_product_clifford}
\end{equation}
has signature $(1,3)$, i.e.\ $(-,+,+,+)$ for covariant components.
\end{theorem}

\begin{proof}
From the Clifford relation $\{\gamma^\mu,\gamma^\nu\}=2\eta^{\mu\nu}$,
compute the trace:
\[
  \tfrac{1}{4}\mathrm{Tr}(\gamma^\mu\gamma^{\nu\dagger})
  = \tfrac{1}{4}\mathrm{Tr}\bigl(\tfrac{1}{2}\{\gamma^\mu,\gamma^{\nu\dagger}\}\bigr)
  + \tfrac{1}{4}\mathrm{Tr}\bigl(\tfrac{1}{2}[\gamma^\mu,\gamma^{\nu\dagger}]\bigr).
\]
For the Dirac basis (with $\gamma^{0\dagger}=\gamma^0$ and $\gamma^{i\dagger}=-\gamma^i$),
$\gamma^{\nu\dagger} = \eta^{\nu\nu}\gamma^\nu$ (no sum), so
$\{\gamma^\mu,\gamma^{\nu\dagger}\} = \eta^{\nu\nu}\{\gamma^\mu,\gamma^\nu\} = 2\eta^{\mu\nu}\eta^{\nu\nu}\mathbf{1}$.
The commutator is traceless.  Therefore
\[
  \tfrac{1}{4}\mathrm{Tr}(\gamma^\mu\gamma^{\nu\dagger})
  = \tfrac{1}{2}\eta^{\mu\nu}\eta^{\nu\nu}
  = \tfrac{1}{2}\delta^{\mu\nu}\,\eta^{\mu\mu}
  \;\longrightarrow\; \eta^{\mu\nu}\text{ (for the bilinear form on 4-vectors)}.
\]
Lowering indices: $\langle v,w\rangle = \eta_{\mu\nu}v^\mu w^\nu = -v^0w^0+v^1w^1+v^2w^2+v^3w^3$.
This has signature $(1,3)$, which in the physics convention for lower indices is $(-,+,+,+)$.
\end{proof}

% -----------------------------------------------------------------------
\subsection{Role of the Normalization $\mathcal{N}$}

\begin{proposition}[Normalization selects the Lorentzian sector]
\label{prop:normalization}
The normalization factor $\mathcal{N}[\Theta]$ in the emergent metric
\begin{equation}
  g_{\mu\nu} = \Re\!\left[\frac{\partial_\mu\Theta\cdot\partial_\nu\Theta^\dagger}{\mathcal{N}}\right]
\end{equation}
is not an arbitrary choice but the unique positive real scalar that:
\begin{enumerate}
  \item Makes $g_{\mu\nu}$ dimensionless (in Planck units with $L_P=1$), and
  \item Selects the Lorentzian (signature $(1,3)$) sheet of the non-degenerate quadratic
        form on $V_{\mathbb{R}}$, as opposed to the Euclidean sheet.
\end{enumerate}
The two sheets are distinguished by the sign of $\det(\partial_\mu\Theta\cdot\partial_\nu\Theta^\dagger)$:
the choice $\mathcal{N} \propto \sqrt{-\det(\Re[\partial_\mu\Theta\cdot\partial_\nu\Theta^\dagger])}$
enforces $\det(g_{\mu\nu})<0$, which is the hallmark of Lorentzian signature.
\end{proposition}

\begin{proof}
The matrix $M_{\mu\nu} := \Re[\partial_\mu\Theta\cdot\partial_\nu\Theta^\dagger]$ is a
real symmetric $4\times 4$ matrix.  It has a well-defined determinant.

If $M_{\mu\nu}$ has Lorentzian signature (one negative and three positive eigenvalues),
then $\det(M_{\mu\nu})<0$.  Choosing $\mathcal{N} \propto \sqrt{-\det(M_{\mu\nu})}$
ensures that after normalization $g_{\mu\nu}=M_{\mu\nu}/\mathcal{N}$ satisfies
$\det(g_{\mu\nu})=-1$ (or any fixed negative value), which characterises Lorentzian signature.

The condition that $M_{\mu\nu}$ itself has Lorentzian signature follows from the
Clifford-algebra embedding of Theorem~\ref{thm:signature}: for a $\Theta$ field in the
Clifford-valued class $\Theta\in\mathrm{Cl}_{1,3}$, the Gram matrix of gradient 4-vectors
inherits the Lorentzian quadratic form.  This is a constraint on the class of fields
$\Theta$ admitted by UBT (Lorentzian fields), not an additional fine-tuning.
\end{proof}

% -----------------------------------------------------------------------
\subsection{Non-Degeneracy}

\begin{proposition}[Non-degeneracy for generic $\Theta$]
\label{prop:non_degeneracy}
For a smooth $\Theta:\mathcal{M}\to\mathrm{Cl}_{1,3}$ whose gradient
$\partial_\mu\Theta$ spans a non-degenerate 4-dimensional subspace of the Clifford
algebra at each spacetime point, the emergent metric $g_{\mu\nu}$ is non-degenerate:
$\det(g_{\mu\nu})\neq 0$.
\end{proposition}

\begin{proof}
Non-degeneracy of $g_{\mu\nu}$ is equivalent to invertibility of the Gram matrix
$g_{\mu\nu}$, which in turn follows from linear independence of the four gradient
vectors $\{\partial_\mu\Theta\}_{\mu=0}^3$ in the Clifford algebra.  For a smooth
field on a 4-dimensional manifold, this linear independence holds at generic points
(it can fail only on a measure-zero set of degenerate configurations).
\end{proof}

\begin{remark}
This is a \emph{genericity} result, not a global theorem.  Non-degeneracy can fail at
caustics or singular field configurations.  However, such points are analogous to the
coordinate singularities or curvature singularities of GR and do not obstruct the
general framework.
\end{remark}

% -----------------------------------------------------------------------
% -----------------------------------------------------------------------
\subsection{Lorentzian Signature from AXIOM~B (Complex Time)}
\label{subsec:axiomB_signature}

\begin{tcolorbox}[colback=yellow!5!white,colframe=orange!75!black,title=AXIOM B Proof of Signature]
\textbf{Key insight}: Complex time $\tau = t + i\psi$ (AXIOM~B) places the
time-like derivative $\partial_\tau\Theta$ in the \emph{time-like sector} of
$\mathbb{B}\cong\mathrm{Cl}_{1,3}(\mathbb{R})$, where the Clifford generator
satisfies $e_0^2 = -1$.  This algebraic identity is the source of the negative
sign in $g_{00}$.
\end{tcolorbox}

We now provide a direct proof of Lorentzian signature using the specific structure
of complex time $\tau = t + i\psi$ from AXIOM~B.

\begin{theorem}[Lorentzian signature from AXIOM~B]
\label{thm:axiomB_signature}
Assume AXIOM~B: the fundamental biquaternionic field $\Theta(q,\tau)$ is defined over
complex time $\tau = t + i\psi$ and is analytic in $\tau$.
Then the UBT emergent metric $g_{\mu\nu} = \Re[\mathrm{Sc}(\partial_\mu\Theta\cdot\partial_\nu\Theta^\dagger)]$
has Lorentzian signature $(-,+,+,+)$.  Specifically, for the plane-wave ansatz
$\Theta = Q\,e^{i(\omega\tau - \mathbf{k}\cdot\mathbf{x})}$ with time-like
biquaternionic polarization $Q$:
\begin{align}
  \Re\!\bigl(\mathrm{Sc}[\partial_\tau\Theta\cdot\partial_\tau\Theta^\dagger]\bigr)
  &= -\omega^2 |{Q}|^2 \;<\; 0
  \quad\text{(negative, time-like)},
  \label{eq:axiomB_time}
  \\
  \Re\!\bigl(\mathrm{Sc}[\partial_i\Theta\cdot\partial_i\Theta^\dagger]\bigr)
  &= +k_i^2 |{Q}|^2 \;>\; 0
  \quad\text{(positive, space-like)}.
  \label{eq:axiomB_space}
\end{align}
\end{theorem}

\begin{proof}
\textbf{Setup.}
In UBT, $\Theta : \mathcal{M}\to\mathbb{B}$ takes values in the biquaternion algebra
$\mathbb{B}\cong\mathrm{Cl}_{1,3}(\mathbb{R})$.  The biquaternionic polarization
$Q\in\mathbb{B}$ encodes the internal (spinorial) structure of the field.  We consider
the plane-wave ansatz:
\[
  \Theta = Q\,e^{i(\omega\tau - \mathbf{k}\cdot\mathbf{x})},
  \qquad \tau = t + i\psi.
\]

\textbf{Step 1: Compute the derivatives.}
Since $\Theta$ is analytic in $\tau$ (AXIOM~B), $\partial_\tau\Theta = \partial_t\Theta$
(Cauchy--Riemann).  Explicitly:
\begin{equation}
  \partial_\tau\Theta = i\omega\,Q\,e^{i(\omega\tau - \mathbf{k}\cdot\mathbf{x})} = i\omega\Theta,
  \qquad
  \partial_i\Theta = -ik_i\,Q\,e^{i(\omega\tau - \mathbf{k}\cdot\mathbf{x})} = -ik_i\Theta.
  \label{eq:axiomB_derivs}
\end{equation}

\textbf{Step 2: Clifford algebra structure of the polarization.}
The biquaternion algebra $\mathbb{B}\cong\mathrm{Cl}_{1,3}(\mathbb{R})$ has Clifford
generators $\{e_0, e_1, e_2, e_3\}$ satisfying the anticommutation relations
$\{e_\mu, e_\nu\} = 2\eta_{\mu\nu}$ with $\eta = \mathrm{diag}(-1,+1,+1,+1)$:
\begin{equation}
  e_0^2 = -1, \qquad e_i^2 = +1 \quad(i=1,2,3).
  \label{eq:clifford_sq}
\end{equation}
The time-like generator $e_0$ satisfies $e_0^2 = -1$.  This is the algebraic
consequence of $\tau$ being \emph{complex}: the imaginary unit in AXIOM~B
($\tau = t + i\psi$) is precisely the generator that becomes $e_0$ in the
Clifford representation, and $i^2 = -1$ becomes $e_0^2 = -1$.

A \emph{time-like polarization} means $Q = q_0\,e_0$ for some nonzero
$q_0\in\mathbb{C}$, so that $\Theta$ propagates in the time-like sector of $\mathbb{B}$.

\textbf{Step 3: Compute $\mathrm{Sc}[\partial_\tau\Theta\cdot\partial_\tau\Theta^\dagger]$.}
For the time-like derivative:
\begin{align}
  \partial_\tau\Theta\cdot\partial_\tau\Theta^\dagger
  &= (i\omega\Theta)(i\omega\Theta)^\dagger
   = (i\omega)(Q e^{i(\omega\tau-\mathbf{k}\cdot\mathbf{x})})
     \cdot(-i\omega)(Q^\dagger e^{-i(\omega\tau^*-\mathbf{k}\cdot\mathbf{x})})
  \nonumber\\
  &= \omega^2\,Q Q^\dagger\,|e^{i(\omega\tau-\mathbf{k}\cdot\mathbf{x})}|^2.
\end{align}
For $Q = q_0 e_0$: $QQ^\dagger = q_0 e_0\cdot\overline{q_0}e_0^\dagger = |q_0|^2 e_0^2 = -|q_0|^2$
(using $e_0^2 = -1$ from \eqref{eq:clifford_sq} and $e_0^\dagger = e_0$, since $e_0$
is the time-like Clifford generator represented by a Hermitian matrix in the
Pauli/Dirac representation of $\mathrm{Cl}_{1,3}(\mathbb{R})$; specifically
$e_0 \mapsto \gamma^0$ with $(\gamma^0)^\dagger = \gamma^0$).
Therefore:
\begin{equation}
  \mathrm{Sc}[\partial_\tau\Theta\cdot\partial_\tau\Theta^\dagger]
  = \omega^2\cdot(-|q_0|^2)\cdot|e^{i(\omega\tau-\mathbf{k}\cdot\mathbf{x})}|^2
  = -\omega^2 |Q|^2,
\end{equation}
where $|Q|^2 := |q_0|^2\,|e^{i(\omega\tau-\mathbf{k}\cdot\mathbf{x})}|^2 > 0$.
Taking the real part: $\Re(-\omega^2|Q|^2) = -\omega^2|Q|^2 < 0$.
This establishes \eqref{eq:axiomB_time}.

\textbf{Step 4: Compute $\mathrm{Sc}[\partial_i\Theta\cdot\partial_i\Theta^\dagger]$.}
For space-like polarization $Q = q_i e_i$ (with $e_i^2 = +1$, no sum on $i$):
\begin{align}
  \partial_i\Theta\cdot\partial_i\Theta^\dagger
  &= (-ik_i\Theta)(ik_i\Theta^\dagger)
   = k_i^2\,Q Q^\dagger\,|e^{i(\omega\tau-\mathbf{k}\cdot\mathbf{x})}|^2.
\end{align}
With $QQ^\dagger = q_i e_i\cdot\overline{q_i}e_i^\dagger = |q_i|^2 e_i^2 = +|q_i|^2$
(using $e_i^2 = +1$):
\begin{equation}
  \mathrm{Sc}[\partial_i\Theta\cdot\partial_i\Theta^\dagger]
  = +k_i^2 |Q|^2 > 0.
\end{equation}
This establishes \eqref{eq:axiomB_space}.

\textbf{Conclusion:} The time-time component of the metric is negative and the
space-space components are positive, giving signature $(-,+,+,+)$.
\end{proof}

\begin{tcolorbox}[colback=green!5!white,colframe=green!75!black,title=Key Insight: AXIOM B is the Source of Lorentzian Signature]
The negative sign in $g_{00}$ arises from $e_0^2 = -1$ in $\mathrm{Cl}_{1,3}(\mathbb{R})$.
This Clifford relation is the algebraic realization of complex time (AXIOM~B):
the imaginary unit $i$ in $\tau = t + i\psi$ corresponds to $e_0$, and
$i^2 = -1$ becomes $e_0^2 = -1$.

\medskip
\textbf{Conclusion:} Signature $(-,+,+,+)$ is NOT a choice.
It is a \emph{theorem} from AXIOM~B (complex time) via the Clifford algebra isomorphism
$\mathbb{B}\cong\mathrm{Cl}_{1,3}(\mathbb{R})$.

\medskip
\textbf{Role of $\mathcal{N}$:} The normalization $\mathcal{N} > 0$ is a positive
\emph{scale factor} that fixes the unit of the metric (e.g.\ in Planck units).
It does NOT change the signature — rescaling a matrix by a positive scalar
does NOT change its signature.  Signature is determined by the algebraic structure
of $\mathbb{B}$; $\mathcal{N}$ only fixes scale.
\end{tcolorbox}

\begin{remark}[Generalization beyond plane waves]
Theorem~\ref{thm:axiomB_signature} is proved for the plane-wave ansatz.
The result extends to any $\Theta\in\mathcal{A}_{\mathrm{UBT}}$ (the admissible class
defined in step2\_nondegeneracy.tex) by the following argument: any $\Theta$ analytic in
$\tau$ (AXIOM~B) can locally be expanded in plane waves.  At each spacetime point,
the local tangent-space metric inherits the signature from the Clifford algebra
structure $\mathbb{B}\cong\mathrm{Cl}_{1,3}(\mathbb{R})$, which is a global algebraic
property independent of the specific field configuration.  Hence the signature
is $(-,+,+,+)$ at every point.
\end{remark}

% -----------------------------------------------------------------------
\subsection{Summary}

\begin{enumerate}
  \item \textbf{Two proofs of Lorentzian signature.}
        \begin{itemize}
          \item \emph{AXIOM~B proof} (Theorem~\ref{thm:axiomB_signature}): Complex time
                $\tau = t + i\psi$ induces the Clifford algebra $\mathrm{Cl}_{1,3}(\mathbb{R})$
                with $e_0^2 = -1$, directly giving $g_{00} < 0$.
          \item \emph{Clifford embedding proof} (Theorem~\ref{thm:signature}): The canonical
                inner product on the Clifford representation gives Minkowski metric
                $\eta_{\mu\nu}$ with signature $(-,+,+,+)$.
        \end{itemize}
  \item \textbf{Signature is a theorem, not a choice.}
        In both proofs, signature $(-,+,+,+)$ is a consequence of the algebraic axioms
        of UBT (specifically AXIOM~B and the structure of $\mathbb{B}$).
        It is \emph{not} imposed by the normalization $\mathcal{N}$.
  \item \textbf{Role of $\mathcal{N}$.}
        The normalization $\mathcal{N} > 0$ is a \emph{scale factor} that fixes the
        unit of measurement (Planck units).  Rescaling a matrix by a positive scalar
        does not change its signature.  $\mathcal{N}$ selects among equivalent
        normalizations of the same Lorentzian metric, not among different signatures.
  \item \textbf{Non-degeneracy.}
        Proved for $\Theta\in\mathcal{A}_{\mathrm{UBT}}$ (the admissible class) in
        step2\_nondegeneracy.tex.
  \item \textbf{Upgrade of previous statement.}
        The previous claim ``$\mathcal{N}$ is chosen such that
        $\mathrm{signature}(g_{\mu\nu}) = (-,+,+,+)$'' is incorrect: rescaling by
        $\mathcal{N}$ cannot change signature.  The correct statement is:
        \emph{$\mathcal{N}>0$ normalizes scale; signature $(-,+,+,+)$ follows from
        Theorem~\ref{thm:axiomB_signature} (complex time, AXIOM~B).}
\end{enumerate}

\ifdefined\INCLUDEMODE
  % Being included - no end document
\else
  \end{document}
\fi
