
\appendix{G: Dark Matter as Topological Hopfions in Unified Biquaternion Theory}

\section*{Introduction}
In the Unified Biquaternion Theory (UBT), dark matter is modeled as topologically stable field configurations – \emph{hopfions} – arising as non-trivial mappings between compactified spatial slices \(S^3\) and internal field space \(S^2\). These configurations are solutions to the UBT field equations for the tensor–spinor field \(\Theta(q,\tau)\) with non-zero Hopf invariant \(Q_H \in \mathbb{Z}\).

Unlike conventional particle dark matter candidates (e.g., WIMPs, axions), hopfions do not require direct coupling to the Standard Model fields; their stability is protected by topological conservation laws.

\section*{Mathematical Formulation}
Let the UBT field \(\Theta: S^3 \to S^2\) be expressed in normalized complex components \(\Theta = (\phi_1, \phi_2) \in \mathbb{C}^2\), satisfying \(|\phi_1|^2 + |\phi_2|^2 = 1\). The Hopf invariant is defined as:

\begin{equation}
Q_H = \frac{1}{(4\pi)^2} \int_{S^3} \mathbf{A} \wedge \mathbf{F},
\end{equation}
where \(\mathbf{F} = d\mathbf{A}\) is the field strength associated with the pullback of the area form on \(S^2\).

The energy functional in the biquaternionic Lagrangian takes the form:
\begin{equation}
E[\Theta] = \int_{\mathbb{R}^3} \left( \alpha |\nabla \Theta|^2 + \beta |\mathbf{F}|^2 \right) \, d^3x,
\end{equation}
with \(\alpha, \beta\) determined by the underlying UBT coupling constants.

\section*{Hopfion Solutions and Stability}
Minimization of \(E[\Theta]\) under fixed \(Q_H\) yields knotted soliton solutions. In the UBT framework, these solutions are generalized to include the complex-time coordinate \(\tau = t + i\psi\), giving rise to \emph{phase–linked hopfions} where the imaginary time \(\psi\) modulates their internal twist.

Stability is guaranteed by the integer-valued Hopf invariant; any continuous deformation preserving field smoothness cannot change \(Q_H\).

\section*{Dark Matter Interpretation}
Hopfions interact only gravitationally and through topology–mediated phase effects with normal matter. Their contribution to the energy–momentum tensor in the UBT metric is given by:
\begin{equation}
T_{\mu\nu}^{\text{hopfion}} = 2\alpha \partial_\mu \Theta \cdot \partial_\nu \Theta^* + 2\beta F_{\mu\lambda} F_{\nu}{}^\lambda - g_{\mu\nu} \mathcal{L}_{\text{hopfion}}.
\end{equation}

Cosmologically, a relic density of hopfions could arise naturally during symmetry-breaking phase transitions in the early Universe, analogously to cosmic strings or monopoles, but without leading to observational inconsistencies thanks to their low interaction cross-section.

\section*{Observational Signatures}
Possible observational consequences include:
\begin{itemize}
\item Microlensing events without electromagnetic counterparts.
\item Non-Gaussian features in the CMB due to large-scale hopfion structures.
\item Anomalous gravitational wave dispersion from hopfion–induced metric perturbations.
\end{itemize}

\section*{Conclusion}
The UBT hopfion model provides a purely topological candidate for dark matter, naturally arising from the field \(\Theta(q,\tau)\) without introducing ad hoc particles. This aligns with the theory’s aim to unify physical phenomena through a single geometric–algebraic framework.

\section*{Appendix G: Dark Matter, Hopfions, and Topological Excitations in the $\Theta$-Field}
\addcontentsline{toc}{section}{Appendix G: Dark Matter, Hopfions, and Topological Excitations in the $\Theta$-Field}
\subsection*{G.1 Motivation}
Within the Unified Biquaternion Theory (UBT), the master field $\Theta(q,\tau)$ allows for nontrivial topological configurations. 
A particularly relevant class are \emph{Hopfions}---localized, finite-energy field configurations characterized by a nonzero Hopf invariant. 
We outline their mathematical structure, dynamics in complex time $\tau=t+i\psi$, energetic properties, and gravitational phenomenology as candidates for (a component of) dark matter.
\subsection*{G.2 Hopf Map and Topological Index}
The canonical Hopf fibration maps $S^3 \to S^2$. 
Introduce a complex doublet $z=(z_1,z_2)^T\in\mathbb{C}^2$ with $z^\dagger z = 1$ (so that $z$ parametrizes $S^3$). 
The unit vector $n:\mathbb{R}^3\cup\{\infty\}\simeq S^3 \to S^2$ is
n^a = z^\dagger \sigma^a z,\qquad a=1,2,3,
where $\sigma^a$ are the Pauli matrices. The $U(1)$ connection and curvature are
A_i = -i\, z^\dagger \partial_i z,\qquad F_{ij} = \partial_i A_j - \partial_j A_i.
The \emph{Hopf invariant} is
Q_H \;=\; \frac{1}{4\pi^2}\int_{\mathbb{R}^3} \mathrm{d}^3x\; \epsilon^{ijk}\, A_i F_{jk} \;\in\; \mathbb{Z},
which counts the linking number of preimages in $n(\mathbf{x})$. Finite-energy boundary conditions compactify space to $S^3$, ensuring $Q_H$ is topological.
\subsection*{G.3 Biquaternionic Representation}
In UBT, the $\Theta$-field takes values in the biquaternion algebra $\mathbb{B}$. 
A minimal embedding of a Hopfion uses a normalized doublet $z(q,\tau)$ extracted from $\Theta(q,\tau)$ through a projection $\Pi:\mathbb{B}\to\mathbb{C}^2$:
z(q,\tau) \;=\; \Pi\!\left[\Theta(q,\tau)\right],\qquad z^\dagger z = 1.
The physical $n(q,\tau)$ and the $U(1)$ potential $A_i(q,\tau)$ then follow from the above constructions, with possible $\psi$-dependence via $\tau=t+i\psi$.
\subsection*{G.4 Energetics and the Faddeev--Skyrme Functional}
A standard energy functional supporting stable Hopfions is the Faddeev--Skyrme model:
E[n] \;=\; \int\! \mathrm{d}^3x \;\Big\{ \alpha\, (\partial_i n)^2 \;+\; \beta\, (F_{ij})^2 \Big\},\qquad n\cdot n=1.
In our setting, $\alpha,\beta$ can acquire a weak $\psi$-dependence from the complex-time sector,
\alpha \to \alpha(\psi),\qquad \beta \to \beta(\psi),
and additional small corrections may arise from couplings to electromagnetic and matter fields (Appendix D, F). 
The topological lower bound scales as $E \gtrsim c\, |Q_H|^{3/4}$ for some constant $c>0$, consistent with known Hopfion energy--charge relations.
\subsection*{G.5 Dynamics in Complex Time}
Allowing slow $\psi$-evolution, the effective Lagrangian density reads
\mathcal{L}_{\mathrm{Hopf}} \;=\; \frac{1}{2} \kappa_\psi \, (\partial_\psi n)^2 \;+\; \frac{1}{2}\rho \, (\partial_t n)^2 \;-\; \alpha(\psi)\, (\partial_i n)^2 \;-\; \beta(\psi)\, (F_{ij})^2 \;-\; V(n;\psi),
with $V$ encoding weak explicit symmetry breaking or environmental pinning. 
Variations give a generalized Landau--Lifshitz-type equation with a Skyrme term and small $\psi$-derivative corrections, stabilizing knotted textures under perturbations:
\rho\, \partial_t^2 n + \kappa_\psi\, \partial_\psi^2 n \;-\; 2\,\alpha \,\Delta n \;+\; \cdots \;=\; \lambda(x,\psi,t)\, n,\qquad n\cdot n=1.
\subsection*{G.6 Gravitational Coupling and Dark-Matter Phenomenology}
Coupling to gravity proceeds via the stress--energy tensor derived from $\mathcal{L}_{\mathrm{Hopf}}$:
T_{\mu\nu} \;=\; \frac{\partial \mathcal{L}}{\partial(\partial^\mu n)}\cdot \partial_\nu n \;-\; g_{\mu\nu}\,\mathcal{L},
supplemented by contributions from the $\psi$-sector. 
In galactic halos, an ensemble of dilute, long-lived Hopfions with typical size $R_H$ and charge $Q_H=\mathcal{O}(1\text{--}10)$ yields an effective mass density
\rho_{\mathrm{DM}} \;\approx\; \frac{N_H\, E(Q_H)}{V_{\mathrm{halo}}} \;\sim\; \frac{N_H}{V_{\mathrm{halo}}}\, c\, |Q_H|^{3/4},
where $N_H$ is the number of Hopfions in volume $V_{\mathrm{halo}}$. 
Since Hopfions are neutral in the $U(1)$ electromagnetic sector (projection to $A_\mu$ vanishes at infinity) and interact mainly through gravity (and possibly very weak $\psi$-mediated channels), they are natural dark-matter candidates in UBT.
\subsection*{G.7 Couplings to Electromagnetism and Matter}
At low energies, allowed gauge-invariant couplings are suppressed by small parameters:
\mathcal{L}_{\mathrm{int}} \;=\; - g_{nA}\, ( \partial_i n \cdot \partial_j n )\, F^{ij} \;-\; g_{nF}\, (F_{ij})^2\, ( \partial_k n )^2 \;-\; g_{n\psi}\, (\partial_\psi n)^2\, \bar{\Psi}\Psi \;+\; \cdots,
leading to minute birefringence-like effects or dispersion shifts (Appendix D) only in extreme conditions. 
Psychon-induced $\psi$-fluctuations (Appendix F) may modulate Hopfion stability via $\kappa_\psi$ and $\beta(\psi)$, offering a route to small, structured signatures without spoiling standard-model tests.
\subsection*{G.8 3D Schematic: Linked Flux Surfaces (Hopf Link)}
\begin{figure}[h!]
\centering
\tdplotsetmaincoords{70}{120}
\begin{tikzpicture}[tdplot_main_coords,scale=1.0, line cap=round,line join=round]
  % Parameters for two linked circles (Hopf link)
  \def\R{2.2}
  \def\r{0.9}
  % Circle 1 in x-y plane
  \draw[very thick, blue!70!black]
    plot[domain=0:360, samples=200, variable=\t]
    ({\R*cos(\t)}, {\R*sin(\t)}, {0});
  % Circle 2 in x-z plane, shifted to link with circle 1
  \draw[very thick, red!70!black]
    plot[domain=0:360, samples=200, variable=\u]
    ({0}, {\r*cos(\u)}, {\r*sin(\u)});
  % Soft field-line hints
  \foreach \k in {0,30,...,330}{
    \draw[blue!40!black, opacity=0.35]
      plot[domain=0:360, samples=80, variable=\t]
      ({(\R+0.3*cos(\k))*cos(\t)}, {(\R+0.3*cos(\k))*sin(\t)}, {0.2*sin(\k)});
  }
  \node[anchor=south] at (0,-2.9,0) {\small Schematic Hopf link representing linked preimages of $n(\mathbf{x})$ (topological charge $Q_H=1$)};
\end{tikzpicture}
\caption{Conceptual 3D schematic of a Hopfion as a linked flux structure (Hopf link).}
\label{fig:hopf_link}
\end{figure}
\subsection*{G.9 Observational and Experimental Windows}
\item \textbf{Astrophysics:} core/halo profiles modified by a population of stable knotted excitations; lensing and rotation curve fits with effective pressure terms from Skyrme corrections.
\item \textbf{Cosmology:} early-universe production of Hopfions during symmetry-breaking transitions; relic abundance depending on $\alpha(\psi),\beta(\psi)$.
\item \textbf{Laboratory analogues:} knotted light and superfluid vortices as table-top proxies to test relaxation/stability scaling.
\subsection*{G.10 Summary and Open Questions}
Hopfions provide a topologically protected, electromagnetically neutral sector consistent with UBT and compatible with precision QED. 
Key open directions include: quantitative halo modeling with $\psi$-dependent parameters, production mechanisms in cosmology, and refined bounds on weak couplings to standard fields.
\subsection*{G.11 Explicit $Q_H=1$ Parametrization (Stereographic Construction)}
Let $r^2=x^2+y^2+z^2$. A standard explicit $S^3\!\to\!S^2$ Hopf-map representative (unit charge $Q_H=1$) can be written via a normalized doublet $z=(z_1,z_2)^T$,
z(\mathbf{x}) \;=\; \frac{1}{r^2+1}\begin{pmatrix}
(r^2-1) + 2 i z \\[4pt]
2(x+i y)
\end{pmatrix},\qquad z^\dagger z = 1.
Then
n^a(\mathbf{x}) = z^\dagger(\mathbf{x})\,\sigma^a\, z(\mathbf{x}),\qquad
A_i(\mathbf{x})=-i\, z^\dagger \partial_i z,\qquad F_{ij}=\partial_i A_j-\partial_j A_i,
and the Hopf invariant
Q_H \;=\; \frac{1}{4\pi^2}\int\!\mathrm{d}^3x\;\epsilon^{ijk} A_i F_{jk} \;=\; 1.
A size (core-radius) parameter $R_0$ is introduced by the replacement $\mathbf{x}\mapsto \mathbf{x}/R_0$ in $z(\mathbf{x})$, which scales the configuration without changing $Q_H$.
\paragraph{Biquaternion Embedding.} In UBT one may take $z=\Pi[\Theta]$ for a projection $\Pi:\mathbb{B}\!\to\!\mathbb{C}^2$; the above $z(\mathbf{x})$ fixes the $\Theta$ phase-texture up to internal rotations, preserving $Q_H=1$ in the projected $U(1)$ sector.
\subsection*{G.12 Order-of-Magnitude Estimates (Size, Energy, Density)}
For the Faddeev--Skyrme energy functional
E[n] \;=\; \int\!\mathrm{d}^3x\;\Big\{\alpha\, (\partial_i n)^2 + \beta\, (F_{ij})^2\Big\},
Derrick scaling with $n(\mathbf{x})\to n(\lambda \mathbf{x})$ yields
E(\lambda)= \alpha\, \lambda\, I_2 + \beta\, \lambda^{-1} I_4,\qquad
I_2=\!\int\! (\partial_i n)^2,\ \ I_4=\!\int\! (F_{ij})^2,
minimized at $\lambda_\star=\sqrt{\dfrac{\beta I_4}{\alpha I_2}}$ with
E_{\min}=2\sqrt{\alpha\beta\, I_2 I_4},\qquad
R_H \;\propto\; \lambda_\star^{-1} \;\sim\; \sqrt{\frac{\beta}{\alpha}}\, \Big(\frac{I_2}{I_4}\Big)^{1/4}.
For $Q_H=1$ one typically has $I_2,I_4=\mathcal{O}(1)$ (dimensionless after extracting a core scale), hence the \emph{size--energy} scalings
R_H \sim \sqrt{\frac{\beta}{\alpha}},\qquad E_{Q=1} \sim \kappa\, \sqrt{\alpha\beta},
with $\kappa=\mathcal{O}(1)$ depending on the specific profile. For general charge $Q_H$, numerical studies give $E \gtrsim c\,|Q_H|^{3/4}$; a practical interpolation is
E(Q_H)\ \approx\ \kappa\, \sqrt{\alpha\beta}\, |Q_H|^{3/4}.
\paragraph{Illustrative Density.} For an ensemble of $N_H$ Hopfions of typical size $R_H$ (non-overlapping) in a halo volume $V_{\rm halo}$,
\rho_{\rm DM} \;\approx\; \frac{N_H\, E(Q_H)}{V_{\rm halo}} \ \sim\  \frac{N_H}{V_{\rm halo}}\ \kappa\, \sqrt{\alpha\beta}\, |Q_H|^{3/4},
while kinematics and stability are controlled by $R_H\!\sim\!\sqrt{\beta/\alpha}$. The $\psi$-sector induces mild running $\alpha(\psi),\beta(\psi)$, enabling environmental dependence without spoiling topological protection.
\subsection*{G.13 Numerical Illustration (Order-of-Magnitude)}
To make the scaling relations concrete, consider illustrative (dimensionally normalized) parameters
\alpha = 2.0,\qquad \beta = 0.5,\qquad Q_H=1,\qquad \kappa \approx 1.
Then
R_H \sim \sqrt{\frac{\beta}{\alpha}} \;=\; \sqrt{\frac{0.5}{2.0}} \;\approx\; 0.5,\qquad
E_{Q=1} \sim \kappa\,\sqrt{\alpha\beta} \;=\; \sqrt{2.0\times 0.5} \;\approx\; 1.0.
For a sparse ensemble in a fiducial volume $V_{\rm halo}$ with number $N_H$,
\rho_{\rm DM} \;\approx\; \frac{N_H\,E_{Q=1}}{V_{\rm halo}} \;\approx\; \frac{N_H}{V_{\rm halo}}.
These values serve only as a \emph{scaling} guide; in physical units, $\alpha,\beta$ inherit dimensions from the microscopic UBT action and may acquire weak $\psi$-dependence, $\alpha(\psi),\beta(\psi)$, which shifts $R_H$ and $E$.