\documentclass[11pt,a4paper]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\newtheorem{theorem}{Theorem}
\newtheorem{assumption}{Assumption}

\title{Unified Biquaternion Theory: Fit-Free Alpha Baseline\\
\large arXiv Preprint}
\author{UBT Team\\
Principal Investigator: Ing. David Jaroš}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
We present the fit-free baseline result for the fine-structure constant $\alpha$ derivation within the Unified Biquaternion Theory (UBT). Under three standard assumptions—geometric locking of mode parameters (A1), standard renormalization scheme (A2), and Thomson-limit extraction (A3)—we rigorously establish that the two-loop counterterm renormalization factor equals unity: $\mathcal{R}_{\mathrm{UBT}} = 1$. This yields $B = \frac{2\pi N_{\mathrm{eff}}}{3R_\psi}$ and $\alpha^{-1} = F(B)$ with no fitted parameters. We provide complete derivations, reproducibility checks, and a replication protocol for independent verification.
\end{abstract}

\tableofcontents

\section{Introduction}

The fine-structure constant $\alpha \approx 1/137.036$ is one of the fundamental dimensionless constants of nature. In standard quantum field theory, it enters as a free parameter. The Unified Biquaternion Theory (UBT) provides a geometric framework where $\alpha$ emerges from the topology of complex time.

This manuscript establishes the \textbf{fit-free baseline}: at two-loop order in perturbation theory, under well-defined assumptions A1--A3, the renormalization factor $\mathcal{R}_{\mathrm{UBT}}$ equals unity. This result eliminates free parameters from the alpha derivation within this framework.

\paragraph{Note on references.} This manuscript is excerpted from the full UBT consolidation document. Some internal references point to sections in the main document (labeled \texttt{app:ct-baseline-R1} and \texttt{thm:RUBT-equals-one}), which correspond to the CT baseline theorem presented here as Theorem~\ref{thm:ct-baseline}. For the complete derivation, see the full repository at \url{https://github.com/DavJ/unified-biquaternion-theory}.

% Define labels for compatibility with included sections
\label{app:ct-baseline-R1}
\label{thm:RUBT-equals-one}

\subsection{Main Result}

\begin{theorem}[CT Baseline, Informal]
Under assumptions A1 (geometric locking), A2 (standard CT scheme with Ward identities), and A3 (Thomson-limit extraction), the finite renormalized ratio at two loops equals:
\[
\boxed{\mathcal{R}_{\mathrm{UBT}} = 1}
\]
\end{theorem}

This implies:
\[
B = \frac{2\pi N_{\mathrm{eff}}}{3R_\psi} \times \mathcal{R}_{\mathrm{UBT}} = \frac{2\pi N_{\mathrm{eff}}}{3R_\psi}
\]
with no adjustable parameters.

\subsection{Scope and Limitations}

\textbf{What this result establishes:}
\begin{itemize}
\item A rigorous baseline for $\mathcal{R}_{\mathrm{UBT}}$ at two-loop order
\item A fit-free framework under explicit assumptions
\item Compatibility with standard QED in the real-time limit
\end{itemize}

\textbf{What this result does NOT claim:}
\begin{itemize}
\item Complete derivation of $\alpha$ from first principles (requires validating A1--A3 and computing $F(B)$)
\item Three-loop or higher-order results
\item Explicit complex-time effects beyond the stated assumptions
\end{itemize}

\section{Assumptions}

We state the three core assumptions explicitly:

\begin{assumption}[A1: Geometric Locking]
\label{ass:A1}
The effective number of modes $N_{\mathrm{eff}}$ and the imaginary-time radius $R_\psi$ are uniquely determined by:
\begin{itemize}
\item Mode counting on the Hermitian slice (Appendix P6)
\item Periodicity condition $\psi \sim \psi + 2\pi$
\item Lorentz invariance on the Hermitian slice
\item Thomson-limit normalization
\end{itemize}
No alternative values satisfy all four conditions simultaneously.
\end{assumption}

\begin{assumption}[A2: CT Renormalization Scheme]
\label{ass:A2}
The complex-time (CT) renormalization scheme satisfies:
\begin{itemize}
\item Dimensional regularization in $d = 4 - 2\epsilon$ dimensions
\item Ward identities: $Z_1 = Z_2$ (vertex $=$ field renormalization)
\item Reduction to QED in the real-time limit ($\psi \to 0$)
\item Standard counterterm structure up to scheme ambiguities
\end{itemize}
\end{assumption}

\begin{assumption}[A3: Observable Extraction]
\label{ass:A3}
The coupling constant is extracted via:
\begin{itemize}
\item Thomson scattering limit at $q^2 = 0$
\item Gauge-invariant observable definition
\item Transverse photon polarization (longitudinal parts vanish)
\end{itemize}
\end{assumption}

\section{CT Baseline Theorem}

The full proof is provided in the accompanying appendices. We state the theorem precisely:

\begin{theorem}[CT Baseline]
\label{thm:ct-baseline}
Under Assumptions~\ref{ass:A1}, \ref{ass:A2}, and \ref{ass:A3}, the finite renormalized two-loop ratio
\[
\mathcal{R}_{\mathrm{UBT}} := \frac{\Pi^{(2)}_{\text{CT, finite}}(q^2=0;\mu)}{\Pi^{(2)}_{\text{QED, finite}}(q^2=0;\mu)} \times \mathcal{N}_{\text{CT}\to\text{QED}}
\]
equals unity at two-loop order:
\[
\mathcal{R}_{\mathrm{UBT}} = 1 + \mathcal{O}(\alpha^3).
\]
\end{theorem}

\subsection{Proof Sketch}

The proof proceeds in three steps:

\textbf{Step 1: CT $\to$ QED reduction.} In the limit $\psi \to 0$, the complex-time scheme reduces to standard QED. All complex-time-dependent propagators and vertices reduce to their real-time counterparts.

\textbf{Step 2: Ward identities.} Gauge invariance enforces $Z_1 = Z_2$ up to $\mathcal{O}(\alpha^3)$. This eliminates scheme-dependent finite parts in the counterterms.

\textbf{Step 3: Thomson normalization.} Fixing the coupling at the physical observable $q^2 = 0$ via Thomson scattering provides a gauge-invariant extraction. Combined with Steps 1--2, this uniquely determines the finite ratio to be unity.

For complete details, see the included appendices from the main UBT consolidation document.

\section{Geometric Inputs}

\input{../../consolidation_project/alpha_two_loop/tex/geometric_inputs_proof.tex}

\section{CT Scheme Definition}

\input{../../consolidation_project/alpha_two_loop/tex/ct_scheme_definition.tex}

\section{$\mathcal{R}_{\mathrm{UBT}}$ Extraction}

\input{../../consolidation_project/alpha_two_loop/tex/R_UBT_extraction.tex}

\section{Reproducibility Checklist}

\input{../reproducibility_checklist.tex}

\section{Fermion Masses: Program Sketch}

We outline a first-principles approach within the biquaternionic framework for fermion mass derivation. The goal is to extend the fit-free methodology (A1--A3) from the gauge sector ($\alpha$ derivation) to the Yukawa sector.

\subsection{Program Structure}

The program consists of four stages:

\paragraph{(i) Algebraic constraints on Yukawas.}
On the Hermitian slice $\mathcal{H} \subset \mathbb{H}_{\mathbb{C}}$, fermions are represented as left ideals and the Higgs field emerges from the biquaternionic structure. The three involutions (complex conjugation, quaternionic conjugation, Hermitian adjoint) implement discrete symmetries (P, T, C) and constrain the allowed Yukawa bilinears.

Under these constraints, the charged-lepton Yukawa matrix admits a decomposition (details in included appendices):
\[
Y_\ell = a \cdot \mathbb{1}_{3\times 3} + b \cdot \Sigma + c \cdot T + \ldots,
\]
where $a, b, c$ are coefficients determined by overlap integrals on $\mathcal{H}$ (not free parameters), and $\Sigma, T$ are fixed matrices from geometric structure.

\paragraph{(ii) Two-loop CT renormalization.}
Extending the CT renormalization framework from Appendix CT to the Yukawa sector, Ward identities and gauge invariance constrain Yukawa counterterms. The key insight: certain combinations of Yukawa couplings become functions of the same geometric invariants ($N_{\mathrm{eff}}, R_\psi$) used in the $\alpha$ derivation, providing a ``locking'' mechanism analogous to $\mathcal{R}_{\mathrm{UBT}} = 1$.

\paragraph{(iii) Sum rules and ratios.}
Before fixing the absolute mass scale, we aim for falsifiable predictions of mass ratios (e.g., $m_\mu/m_e$, $m_\tau/m_\mu$) independent of the overall scale. These ratios emerge from eigenvalues of $Y_\ell$ and can be compared to experimental PDG values with full error budgeting.

\textbf{Pre-registration requirement}: All mass ratio predictions must be published before comparison to data, preventing hindsight bias.

\paragraph{(iv) Absolute scale anchor.}
Two candidate mechanisms for fixing the overall fermion mass scale without empirical input:
\begin{enumerate}
  \item \textbf{Invariant measure normalization}: Relate the Yukawa scale to the same geometric normalization used for $N_{\mathrm{eff}}$ and $R_\psi$.
  \item \textbf{Topological constraint}: Tie fermion masses to soliton energies and winding numbers derived from UBT field equations.
\end{enumerate}

Any anchor must satisfy: (a) no fitted parameters, (b) scheme independence, (c) consistency with mass ratios, (d) falsifiability.

\subsection{Current Status and Timeline}

\textbf{Status}: Framework established; detailed two-loop calculations in progress.

\textbf{Milestones}:
\begin{itemize}
  \item Lepton mass ratio predictions: 12--18 months
  \item Absolute scale anchor: 24--36 months
  \item Extension to quark sector: 36+ months
\end{itemize}

\textbf{Falsification}: If predicted ratios disagree with experimental values beyond error budget, the Yukawa texture hypothesis is falsified.

\subsection{Relation to $\alpha$ Baseline}

The fermion mass program mirrors the fit-free $\alpha$ derivation:
\begin{itemize}
  \item \textbf{Geometric locking} (A1): Same $N_{\mathrm{eff}}, R_\psi$ constrain Yukawa coefficients
  \item \textbf{CT renormalization} (A2): Ward identities enforce consistency
  \item \textbf{Observable extraction} (A3): Mass ratios provide gauge-invariant tests
\end{itemize}

Both programs uphold the principle: \textbf{no fitted parameters, only geometric constraints and renormalization consistency}.

For detailed technical content, see the masses module appendices in the full UBT repository.

\section{Conclusions}

We have established a fit-free baseline for the fine-structure constant derivation in UBT:
\begin{itemize}
\item \textbf{Result}: $\mathcal{R}_{\mathrm{UBT}} = 1$ at two loops under A1--A3
\item \textbf{Implication}: $B = \frac{2\pi N_{\mathrm{eff}}}{3R_\psi}$ with no free parameters
\item \textbf{Requirement}: Any $\mathcal{R}_{\mathrm{UBT}} \neq 1$ must be derived by explicit CT calculation
\end{itemize}

This baseline provides a rigorous foundation for further development of UBT's predictive framework.

\subsection{Future Work}

\begin{enumerate}
\item Complete three-loop $\mathcal{R}_{\mathrm{UBT}}$ calculation
\item Finalize the function $F(B)$ for numeric prediction
\item Extend to weak mixing angle $\theta_W$ and strong coupling $\alpha_s$
\item Develop experimental tests of complex-time signatures
\end{enumerate}

\section*{Acknowledgments}

We thank the open-source community for code review and replication efforts. This work benefits from public scrutiny and collaborative verification.

\section*{Data Availability}

All source code, LaTeX documents, test suites, and build artifacts are publicly available at:

\url{https://github.com/DavJ/unified-biquaternion-theory}

Replication protocol: \texttt{docs/REPLICATION\_PROTOCOL.md}

\bibliographystyle{plain}
\bibliography{../../consolidation_project/references}

\end{document}
