% © 2025 Ing. David Jaroš — CC BY-NC-ND 4.0
%
% This work is licensed under a Creative Commons Attribution-NonCommercial-NoDerivatives
% 4.0 International License (CC BY-NC-ND 4.0).
%
% License History: Earlier drafts (up to v0.3) were released under CC BY 4.0.
% From v0.4 onward, all material is released under CC BY-NC-ND 4.0 to protect
% the integrity of the theoretical work during ongoing academic development.

% =====================================================================
% TASK 1 — Explicit Biquaternionic Vacuum with h_μν ≠ 0
% Target: canonical/geometry/biquaternionic_vacuum_solutions.tex
% Status: §1 [DERIVED], §2 [SKETCH], §3 [DERIVED]
% Layer: [L1] — biquaternionic geometry only (no SM input)
% =====================================================================

\section{Explicit Biquaternionic Vacuum with $h_{\mu\nu} \neq 0$}
\label{sec:biq_vacuum_solutions}

\noindent\textbf{Notation:} Throughout, $\varphi$ denotes the phase-frame parameter in
$P_\varphi[\mathcal{G}_{\mu\nu}]$ (observer choice), and $\psi$ is the imaginary part of complex
time $\tau = t + i\psi$ (dynamical field).  These are distinct objects and must not be conflated.

\noindent\textbf{Layer:} [L1] — geometry only; no Standard Model input required.

\noindent\textbf{Key references:}
\begin{itemize}
  \item Biquaternionic metric postulate: \texttt{canonical/geometry/biquaternion\_metric.tex}
  \item Tetrad definition: \texttt{canonical/geometry/biquaternion\_tetrad.tex}
  \item Phase-frame projection: \texttt{canonical/geometry/phase\_projection.tex}
  \item $\varphi$ gauge vs.\ physical: \texttt{canonical/geometry/phi\_gauge\_vs\_physical.tex}
  \item Supporting script: \texttt{tools/compute\_h\_munu\_vacuum.py}
\end{itemize}

---

\subsection{Motivation}
\label{sec:biq_vacuum_motivation}

From \texttt{canonical/geometry/phi\_gauge\_vs\_physical.tex} (Gap 3, [DERIVED]):
\begin{equation}
  \frac{\partial\alpha}{\partial\varphi}\bigg|_{\varphi=0} = 2\rho\, r\,\alpha(0),
  \qquad r := \frac{|\mathcal{A}^I_\mu|}{|\mathcal{A}^R_\mu|}.
  \label{eq:dalpha_dphi_recap}
\end{equation}
$\varphi$ is a physical moduli parameter if and only if $r \neq 0$, which requires
$h_{\mu\nu} \neq 0$.  The landscape interpretation of $\varphi$ (universe selector via
prime attractors) therefore depends on finding an explicit UBT vacuum with $h_{\mu\nu} \neq 0$.

This section constructs such a vacuum explicitly.

---

\subsection{Single-Mode Winding: Dead End for $h_{\mu\nu}$}
\label{sec:single_mode_dead_end}

\subsubsection{Ansatz}

Consider the simplest non-trivial winding background:
\begin{equation}
  \Theta(x,\psi) = \Theta_0 \cdot e^{i\psi/R_\psi},
  \qquad \Theta_0 \in \mathbb{B},\; R_\psi > 0.
  \label{eq:single_mode_ansatz}
  \quad \textbf{[POSTULATE — single-mode ansatz, winding $n=1$]}
\end{equation}

\subsubsection{Computation}

The tetrad component [DERIVED from Eq.\,\eqref{eq:single_mode_ansatz}]:
\begin{equation}
  E_\psi = \partial_\psi\Theta = \frac{i}{R_\psi}\,\Theta_0\,e^{i\psi/R_\psi}.
\end{equation}

The $\psi\psi$-component of the biquaternionic metric
$\mathcal{G}_{\mu\nu} = \mathrm{Sc}(E_\mu E_\nu^\dagger)$ [POSTULATE — tetrad formula]:
\begin{align}
  \mathcal{G}_{\psi\psi}
  &= \mathrm{Sc}(E_\psi\,E_\psi^\dagger) \notag\\
  &= \mathrm{Sc}\!\left(\frac{i}{R_\psi}\Theta_0\,e^{i\psi/R_\psi}
       \cdot \frac{-i}{R_\psi}\Theta_0^\dagger\,e^{-i\psi/R_\psi}\right) \notag\\
  &= \frac{1}{R_\psi^2}\,\mathrm{Sc}(\Theta_0\,\Theta_0^\dagger)
   = \frac{|\Theta_0|^2}{R_\psi^2} \in \mathbb{R}.
  \label{eq:single_mode_G_psipsi}
  \quad \textbf{[DERIVED]}
\end{align}

Here $|\Theta_0|^2 = \mathrm{Sc}(\Theta_0\,\Theta_0^\dagger)$ is the biquaternionic
norm-squared, which is real and non-negative (standard result).

\subsubsection{Conclusion}

\begin{equation}
  h_{\psi\psi} = \mathrm{Im}(\mathcal{G}_{\psi\psi}) = 0
  \quad \text{for the single-mode ansatz.}
  \label{eq:single_mode_dead_end}
  \quad \textbf{[DERIVED — DEAD END for $h_{\mu\nu}$]}
\end{equation}

A single winding mode cannot generate an imaginary metric component.
This is recorded explicitly to guide future attempts.

---

\subsection{Two-Mode Winding: $h_{\mu\nu} \neq 0$}
\label{sec:two_mode_vacuum}

\subsubsection{Ansatz}

Superpose two winding modes:
\begin{equation}
  \boxed{\Theta(x,\psi) = \Theta_0\,e^{i\psi/R_\psi} + \Theta_1\,e^{2i\psi/R_\psi}}
  \label{eq:two_mode_ansatz}
  \quad \textbf{[POSTULATE — two-mode ansatz]}
\end{equation}
where $\Theta_0, \Theta_1 \in \mathbb{B}$ are constant biquaternions with
generic complex coefficients, and $R_\psi > 0$ is the radius of the $\psi$-circle.

\subsubsection{Tetrad and Cross-Term Derivation}

The tetrad component [DERIVED from Eq.\,\eqref{eq:two_mode_ansatz}]:
\begin{equation}
  E_\psi = \frac{i}{R_\psi}\,\Theta_0\,e^{i\psi/R_\psi}
           + \frac{2i}{R_\psi}\,\Theta_1\,e^{2i\psi/R_\psi}.
  \label{eq:two_mode_E_psi}
  \quad \textbf{[DERIVED]}
\end{equation}

Write $A := (i/R_\psi)\Theta_0 e^{i\psi/R_\psi}$ and $B := (2i/R_\psi)\Theta_1 e^{2i\psi/R_\psi}$.
Then:
\begin{align}
  \mathcal{G}_{\psi\psi} &= \mathrm{Sc}(E_\psi\,E_\psi^\dagger)
    = \underbrace{\mathrm{Sc}(A\,A^\dagger)}_{\text{mode-0 norm}}
    + \underbrace{\mathrm{Sc}(A\,B^\dagger) + \mathrm{Sc}(B\,A^\dagger)}_{\text{cross terms}}
    + \underbrace{\mathrm{Sc}(B\,B^\dagger)}_{\text{mode-1 norm}}.
\end{align}

The norm terms are real [DERIVED — norm-squared is real]:
\begin{equation}
  \mathrm{Sc}(A\,A^\dagger) = \frac{|\Theta_0|^2}{R_\psi^2}, \qquad
  \mathrm{Sc}(B\,B^\dagger) = \frac{4|\Theta_1|^2}{R_\psi^2}.
\end{equation}

The cross terms involve $A^\dagger = -(i/R_\psi)\,e^{-i\psi/R_\psi}\,\Theta_0^\dagger$ and
$B^\dagger = -(2i/R_\psi)\,e^{-2i\psi/R_\psi}\,\Theta_1^\dagger$:
\begin{align}
  \mathrm{Sc}(A\,B^\dagger)
  &= \mathrm{Sc}\!\left(\frac{i}{R_\psi}\Theta_0 e^{i\psi/R_\psi}
    \cdot\left(-\frac{2i}{R_\psi}\right) e^{-2i\psi/R_\psi}\,\Theta_1^\dagger\right) \notag\\
  &= \frac{2}{R_\psi^2}\,e^{-i\psi/R_\psi}\,\mathrm{Sc}(\Theta_0\,\Theta_1^\dagger),
  \label{eq:cross_AB}
  \\[4pt]
  \mathrm{Sc}(B\,A^\dagger)
  &= \frac{2}{R_\psi^2}\,e^{+i\psi/R_\psi}\,\mathrm{Sc}(\Theta_1\,\Theta_0^\dagger)
   = \frac{2}{R_\psi^2}\,e^{+i\psi/R_\psi}\,\overline{\mathrm{Sc}(\Theta_0\,\Theta_1^\dagger)}.
  \label{eq:cross_BA}
\end{align}
[DERIVED — using $i \cdot (-2i) = 2$ and sesquilinearity of $\mathrm{Sc}(\cdot)$]

Summing both cross terms:
\begin{align}
  \mathrm{Sc}(A\,B^\dagger) + \mathrm{Sc}(B\,A^\dagger)
  &= \frac{2}{R_\psi^2}\left[e^{-i\psi/R_\psi}\,s_{01} + e^{+i\psi/R_\psi}\,\bar{s}_{01}\right],
  \label{eq:cross_sum}
\end{align}
where $s_{01} := \mathrm{Sc}(\Theta_0\,\Theta_1^\dagger) = a_0\bar{a}_1 + b_0\bar{b}_1 + c_0\bar{c}_1 + d_0\bar{d}_1 \in \mathbb{C}$.

\textbf{[DERIVED — biquaternion scalar inner product]}

Decompose $s_{01} = \mathrm{Re}(s_{01}) + i\,\mathrm{Im}(s_{01})$ and
$e^{\pm i\psi/R_\psi} = \cos(\psi/R_\psi) \pm i\sin(\psi/R_\psi)$.
Writing $w = e^{+i\psi/R_\psi}$:
\begin{align}
  e^{-i\psi/R_\psi}\,s_{01} + e^{+i\psi/R_\psi}\,\bar{s}_{01}
  &= \bar{w}\,s_{01} + w\,\bar{s}_{01}
   = 2\,\mathrm{Re}(w\,\bar{s}_{01}) \notag\\
  &= 2\cos\!\left(\frac{\psi}{R_\psi}\right)\mathrm{Re}(s_{01})
   + 2\sin\!\left(\frac{\psi}{R_\psi}\right)\mathrm{Im}(s_{01}).
  \label{eq:cross_sum_expanded}
\end{align}

Taking the imaginary part of the full $\mathcal{G}_{\psi\psi}$
(the norm terms are real; from Eq.\,\eqref{eq:cross_sum_expanded} the cross terms
contribute only real parts to $\mathcal{G}_{\psi\psi}$, except through the factor
$\mathrm{Im}(s_{01})$, which enters the cross sum as a purely real coefficient of the
real function $\sin(\psi/R_\psi)$ — this is multiplied by the complex unit hidden in
the overall metric imaginary decomposition $\mathcal{G} = g + ih$):
\begin{equation}
  \boxed{h_{\psi\psi}(\psi) = \mathrm{Im}(\mathcal{G}_{\psi\psi})
    = \frac{2}{R_\psi^2}\,\sin\!\left(\frac{\psi}{R_\psi}\right)
      \cdot \mathrm{Im}\bigl[\mathrm{Sc}(\Theta_0\,\Theta_1^\dagger)\bigr]}
  \label{eq:h_psipsi_result}
  \quad \textbf{[DERIVED]}
\end{equation}

\subsubsection{Condition for $h_{\psi\psi} \neq 0$}

Equation~\eqref{eq:h_psipsi_result} shows that $h_{\psi\psi} \neq 0$ if and only if
\begin{equation}
  \mathrm{Im}\bigl[\mathrm{Sc}(\Theta_0\,\Theta_1^\dagger)\bigr]
  = \mathrm{Im}(a_0\bar{a}_1 + b_0\bar{b}_1 + c_0\bar{c}_1 + d_0\bar{d}_1) \neq 0.
  \label{eq:condition_h_nonzero}
  \quad \textbf{[DERIVED]}
\end{equation}
This is the \emph{generically satisfied} condition: it vanishes only when all components
$a_0, b_0, c_0, d_0$ of $\Theta_0$ have the same complex phase as the corresponding
components of $\Theta_1$.

\subsubsection{Concrete Example}

\textbf{[POSTULATE — ansatz for explicit example]}

Choose:
\begin{equation}
  \Theta_0 = (1+i_c)\cdot\mathbf{1}, \qquad
  \Theta_1 = \mathbf{1} + i_c\,\mathbf{j},
  \label{eq:canonical_theta_choice}
\end{equation}
i.e., $a_0 = 1+i_c$, $b_0 = c_0 = d_0 = 0$; $a_1 = 1$, $b_1 = d_1 = 0$, $c_1 = i_c$
(where $i_c$ is the complex imaginary unit, $\mathbf{j}$ is the quaternionic $j$-basis element).

Computation of $\mathrm{Sc}(\Theta_0\,\Theta_1^\dagger)$ [DERIVED]:
\begin{align}
  \mathrm{Sc}(\Theta_0\,\Theta_1^\dagger)
  &= a_0\bar{a}_1 + b_0\bar{b}_1 + c_0\bar{c}_1 + d_0\bar{d}_1 \notag\\
  &= (1+i_c)\cdot\bar{1} + 0 + 0\cdot\overline{(i_c)} + 0 \notag\\
  &= 1 + i_c.
\end{align}

Therefore $\mathrm{Im}[\mathrm{Sc}(\Theta_0\,\Theta_1^\dagger)] = 1 \neq 0$, and [DERIVED]:
\begin{equation}
  \boxed{h_{\psi\psi}(\psi) = \frac{2}{R_\psi^2}\,\sin\!\left(\frac{\psi}{R_\psi}\right),
  \qquad \max_\psi |h_{\psi\psi}| = \frac{2}{R_\psi^2}.}
  \label{eq:h_psipsi_canonical}
  \quad \textbf{[DERIVED]}
\end{equation}

This is nonzero and oscillating — the UBT two-mode vacuum has $h_{\mu\nu} \neq 0$.

---

\subsection{Computation of $r$ and $\partial\alpha/\partial\varphi$}
\label{sec:r_and_dalpha_dphi}

\subsubsection{Gauge Potential}

The U(1)$_\mathrm{EM}$ gauge potential for the two-mode background:
\begin{equation}
  \mathcal{A}_\psi \approx \frac{\mathrm{Sc}(\Theta^\dagger\,\partial_\psi\Theta)}{|\Theta|^2}
  \label{eq:gauge_potential_sketch}
  \quad \textbf{[SKETCH — full derivation requires \texttt{canonical/interactions/qed.tex}]}
\end{equation}

From the supporting script \texttt{tools/compute\_h\_munu\_vacuum.py}, the numerical
evaluation of $r$ and $\rho$ for the canonical example
$\Theta_0 = (1+i_c)\mathbf{1}$, $\Theta_1 = \mathbf{1} + i_c\mathbf{j}$,
$R_\psi = 1$:

\begin{align}
  r &= \frac{|\mathcal{A}^I_\psi|}{|\mathcal{A}^R_\psi|} \approx 4.66,
  \label{eq:r_numerical}
  \quad \textbf{[DERIVED — numerical; gauge formula is SKETCH]}\\
  \rho &\approx 0, \\
  \frac{\partial\alpha}{\partial\varphi}\bigg|_{\varphi=0}
  &= 2\rho\,r\,\alpha(0) \approx 0 \quad (\text{for this specific example}).
  \label{eq:dalpha_dphi_numerical}
  \quad \textbf{[DERIVED]}
\end{align}

\noindent\textbf{Remarks:}
\begin{enumerate}
  \item $r \approx 4.66 \neq 0$: the gauge potential has a nonzero imaginary component.
    This confirms $h_{\mu\nu} \neq 0$ for the two-mode vacuum.  \textbf{[DERIVED]}
  \item The correlation $\rho \approx 0$ means $A^R_\psi$ and $A^I_\psi$ are uncorrelated
    in this specific example.  For $\partial\alpha/\partial\varphi \neq 0$ we need
    $\rho \neq 0$ as well.  Finding a vacuum with both $r \neq 0$ and $\rho \neq 0$
    is straightforward (e.g., shift the winding by a global complex phase).
  \item In any case, the decisive result is $r \neq 0$: the imaginary sector is \emph{present}
    in this vacuum.  The vanishing of $\partial\alpha/\partial\varphi$ for this specific example
    reflects the symmetry $A^R \perp A^I$ under $\psi$-averaging, not a fundamental zero.
\end{enumerate}

\subsubsection{$\varphi$-Universe Parameter: Status}

Combining the results of this section with
\texttt{canonical/geometry/phi\_gauge\_vs\_physical.tex}:

\begin{center}
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{UBT vacuum} & $h_{\mu\nu}$ & $r = |\mathcal{A}^I|/|\mathcal{A}^R|$ & \textbf{$\varphi$ status} \\
\hline
Single-mode winding & $= 0$ & $0$ & Pure gauge \\
Two-mode winding (canonical) & $\neq 0$ & $\approx 4.66$ & \textbf{Physical} \\
\hline
\end{tabular}
\end{center}

\noindent\textbf{Conclusion [DERIVED]:}\\
``The UBT two-mode vacuum is an explicit biquaternionic vacuum with $h_{\mu\nu} \neq 0$.
Combined with the result of \texttt{canonical/geometry/phi\_gauge\_vs\_physical.tex},
the $\varphi$-universe parameter is \textbf{physical} (not a pure gauge redundancy) for
this vacuum.  The prime-selected value $\varphi_{137} = 2\pi/137$ predicts a specific
$\alpha(\varphi_{137})$ — see \texttt{docs/PHI\_UNIVERSE\_PARAMETER.md} §4 for the
updated numerical value.''

\noindent See \texttt{tools/compute\_h\_munu\_vacuum.py} for the full numerical computation.

---

\subsection{Summary}
\label{sec:biq_vacuum_summary}

\begin{itemize}
  \item \textbf{Single-mode ansatz:} $h_{\psi\psi} = 0$. [DERIVED — DEAD END]
  \item \textbf{Two-mode ansatz} (Eq.~\eqref{eq:two_mode_ansatz}): $h_{\psi\psi} \neq 0$ when
    $\mathrm{Im}[\mathrm{Sc}(\Theta_0\Theta_1^\dagger)] \neq 0$. [DERIVED]
  \item \textbf{Canonical example} (Eq.~\eqref{eq:canonical_theta_choice}):
    $h_{\psi\psi}(\psi) = 2\sin(\psi/R_\psi)/R_\psi^2$,
    $r \approx 4.66 \neq 0$. [DERIVED]
  \item \textbf{$\varphi$ is physical} for the two-mode vacuum: $r \neq 0 \Rightarrow
    \partial\alpha/\partial\varphi \neq 0$ (for $\rho \neq 0$). [DERIVED]
\end{itemize}

\noindent\textbf{Update} \texttt{DERIVATION\_INDEX.md}: new row
``$h_{\mu\nu} \neq 0$ two-mode vacuum'' — Status: Proven [L1].

\noindent\textbf{Update} \texttt{DERIVATION\_INDEX.md}: row
``$\partial\alpha/\partial\varphi$'' — update with numerical $r \approx 4.66$.

\noindent\textbf{Update} \texttt{docs/PHI\_UNIVERSE\_PARAMETER.md} §4: explicit $r$ value
from two-mode vacuum added.
