
\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{authblk}
\geometry{margin=1in}
\title{A Rigorous Unified Field Theory on Biquaternionic Manifold \( \mathbb{B}^4 \): Consciousness, Quantum Fields, and Emergent Space-Time}
\author{David Jaroš}
\date{2025}

\begin{document}

\maketitle

\begin{abstract}
We present a novel and rigorous unified field theory constructed on a four-dimensional biquaternionic manifold \( \mathbb{B}^4 \), which incorporates complexified spacetime and internal phase dimensions naturally into its geometric structure. The core field \( \Theta(q) \) is a tensor-spinor-gauge-valued section over \( \mathbb{B}^4 \), encoding all known fundamental interactions, emergent quantum behavior, and phenomenology of consciousness. The theory eliminates the need for external projections ...

We develop a rigorous unified field theory on a four-dimensional biquaternionic manifold \( \mathbb{B}^4 = (\mathbb{C} \otimes \mathbb{H})^4 \), where each coordinate is a biquaternion representing complexified internal and external degrees of freedom. The primary field \( \Theta(q) \) is a tensor-spinor-valued section over \( \mathbb{B}^4 \), encoding gravitational, gauge, quantum, and cognitive structure. We formulate the covariant dynamics, metric geometry, low-energy quantum limits, emergent conscious par...
\end{abstract}

\section{Biquaternionic Manifold and Field Definition}

We define the manifold \( \mathcal{M} = \mathbb{B}^4 \), where each point is a 4-tuple of biquaternions:
\[
q^\mu = x^\mu + i y^\mu + \mathbf{j} z^\mu + i \mathbf{j} w^\mu, \quad \mu = 0,1,2,3,
\]
with components \( x^\mu, y^\mu, z^\mu, w^\mu \in \mathbb{R} \). The basis elements \( \{1, \mathbf{i}, \mathbf{j}, \mathbf{k}\} \) satisfy the quaternionic algebra.

The unified field is defined as:
\[
\Theta(q) \in \Gamma \left( T^{(1,1)}(\mathbb{B}^4) \otimes \mathbb{S} \otimes \mathbb{G} \right),
\]
where \( T^{(1,1)} \) denotes the (1,1) tensor bundle over \( \mathbb{B}^4 \), \( \mathbb{S} \) is a spinor bundle, and \( \mathbb{G} \) is an internal gauge fiber.

\section{Metric Geometry and Covariant Derivatives}

We define a complexified metric tensor on \( \mathbb{B}^4 \):
\[
G_{\mu\nu}(q) = \langle dq^\mu, dq^\nu \rangle,
\]
where \( \langle \cdot, \cdot \rangle \) is a biquaternionic inner product. The affine and spin connections \( \Gamma^\rho_{\mu\nu} \), \( \Omega_\mu \) are derived accordingly.

Covariant derivative:
\[
\mathcal{D}_\mu \Theta = \partial_\mu \Theta + \Omega_\mu \cdot \Theta + i g A_\mu^a T^a \Theta,
\]
with gauge coupling and spin connection terms included.


\section{General Relativity and Emergent Geometry}

The metric tensor \( G_{\mu\nu}(q) \) defined over \( \mathbb{B}^4 \) generalizes the Lorentzian metric of General Relativity to a biquaternionic setting. The curvature tensors are constructed from the complexified affine connection:
\[
R^\rho_{\ \sigma\mu\nu} = \partial_\mu \Gamma^\rho_{\nu\sigma} - \partial_\nu \Gamma^\rho_{\mu\sigma} + \Gamma^\rho_{\mu\lambda} \Gamma^\lambda_{\nu\sigma} - \Gamma^\rho_{\nu\lambda} \Gamma^\lambda_{\mu\sigma}.
\]
The Einstein field equations emerge as a projection of the variation of the geometric Lagrangian:
\[
\delta \mathcal{L}_{\text{geom}} = \frac{1}{2\kappa} \left( R_{\mu\nu} - \frac{1}{2} G_{\mu\nu} R \right) \delta G^{\mu\nu}.
\]
This shows that classical spacetime curvature is a low-energy limit of the intrinsic geometry of the biquaternionic manifold. The extended degrees of freedom in \( y^\mu, z^\mu \) give rise to higher-order corrections and dynamical compactification in early-universe cosmology.
\section{Lagrangian and Field Equations}

Total action on \( \mathbb{B}^4 \) reads:
\[
S = \int_{\mathbb{B}^4} d^4q \, \sqrt{|\det G|} \left( \mathcal{L}_{\Theta} + \mathcal{L}_{\text{geom}} + \mathcal{L}_{\text{gauge}} \right),
\]
where:
\[
\mathcal{L}_{\Theta} = \Re\left[ \bar{\Theta} \left( i \Gamma^\mu \mathcal{D}_\mu - M(q) \right) \Theta \right],
\]
with \( M(q) \) potentially depending on internal phase coordinates \( y^\mu, z^\mu \), representing cognitive or entropic symmetry breaking.

\section{Tensor Decomposition of \(\Theta\)}

The field can be locally decomposed as:
\[
\Theta(q) = \phi(q) + i \psi^\mu(q) \gamma_\mu + \eta(q) \mathbb{I} + i \chi(q) \mathbb{J},
\]
where \( \chi(q) \) represents internal oscillations related to consciousness and subjective time.

\section{Quantum Limits and Classical Reductions}

In the real spacetime limit \( y^\mu, z^\mu, w^\mu \to 0 \), the action reduces to:
\begin{itemize}
  \item Dirac equation for spinor sector:
  \[
  (i \gamma^\mu \partial_\mu - m) \psi = 0,
  \]
  \item Schrödinger equation in nonrelativistic limit:
  \[
  i \hbar \partial_t \psi = \left( -\frac{\hbar^2}{2m} \nabla^2 + V \right) \psi,
  \]
  \item Effective quantum field theory in flat \( \mathbb{R}^4 \subset \mathbb{B}^4 \).
\end{itemize}


\section{Gauge Theory and Standard Model Embedding}

The internal gauge structure of \( \Theta \) accommodates:
\[
\mathbb{G} \cong SU(3) \times SU(2) \times U(1),
\]
and the covariant derivative generalizes to the Yang-Mills form. Higgs-like mass terms can arise via internal symmetry oscillations in the \( z^\mu, w^\mu \) directions.

\section{Gauge Symmetries: QED, QCD and the Standard Model Embedding}

We extend the field \(\Theta(q)\) to carry internal gauge indices under the symmetry group \( \mathcal{G} = U(1) \times SU(2) \times SU(3) \), corresponding to the Standard Model gauge groups for electromagnetism, weak, and strong interactions.

Let \(A_\mu = A_\mu^a T^a\) denote the full gauge connection, with generators \(T^a\) acting on \(\Theta(q)\) in the appropriate internal representation.

The full covariant derivative becomes:
\[
\mathcal{D}_\mu \Theta = \partial_\mu \Theta + \Omega_\mu \cdot \Theta + i g_1 B_\mu Y \Theta + i g_2 W_\mu^i \tau^i \Theta + i g_3 G_\mu^a \lambda^a \Theta,
\]
where:
- \(B_\mu\) is the \(U(1)_Y\) hypercharge field,
- \(W_\mu^i\) are the weak \(SU(2)_L\) fields,
- \(G_\mu^a\) are the gluon fields of \(SU(3)_C\),
- \(Y\), \(\tau^i\), and \(\lambda^a\) are the corresponding generators.

The gauge-invariant kinetic Lagrangian reads:
\[
\mathcal{L}_{\text{gauge}} = -\frac{1}{4} \sum_a F^{a}_{\mu\nu} F^{a\,\mu\nu},
\]
with:
\[
F^a_{\mu\nu} = \partial_\mu A^a_\nu - \partial_\nu A^a_\mu + f^{abc} A^b_\mu A^c_\nu,
\]
where \(f^{abc}\) are the structure constants of the respective Lie algebras.

The interaction term in the unified Lagrangian:
\[
\mathcal{L}_{\Theta, \text{int}} = \Re \left[ \bar{\Theta} i \Gamma^\mu \mathcal{D}_\mu \Theta \right],
\]
ensures that the dynamics of \(\Theta\) is fully coupled to the standard model gauge fields.

:\section{Low-Energy Limits: Dirac and Schrödinger Equations}

To demonstrate compatibility with established quantum mechanics, we derive the Dirac and Schrödinger equations as limiting cases of the unified biquaternionic field equation for \(\Theta(q)\).

Starting from the unified covariant derivative:
\[
\mathcal{D}_A \Theta = \partial_A \Theta + \Omega_A \cdot \Theta + i g A_A^a T^a \Theta,
\]
we consider the case where curvature effects are small, gauge fields are static or slowly varying, and the manifold \(\mathbb{C}^5\) can be approximated by a local inertial frame. In this approximation, the equation of motion simplifies to:
\[
\Gamma^A \partial_A \Theta = m \Theta,
\]
which reduces to the **Dirac equation** in the limit where only 4 spacetime dimensions are active and \(\Gamma^A\) matrices correspond to standard Dirac matrices \(\gamma^\mu\). This confirms that the spinor nature of \(\Theta\) is consistent with relativistic quantum field theory.

In the non-relativistic limit (\(v \ll c\)), a standard Foldy–Wouthuysen decomposition leads to the Schrödinger equation:
\[
i \hbar \frac{\partial \psi}{\partial t} = \left( -\frac{\hbar^2}{2m} \nabla^2 + V \right) \psi,
\]
where \(\psi\) is a projection of \(\Theta\) onto low-energy modes.

Thus, our unified theory contains both quantum mechanics and relativistic field theory as natural limits, while offering a richer structure due to its biquaternionic and spin-tensor form defined on \(\mathbb{C}^5\).

\section{Conscious Oscillations and Emergent Mind}

Internal modes \( \chi(q) \sim \sin(\omega \cdot y^\mu) \) describe periodic subjective phase. Collapse of these oscillations projects classical reality. We interpret eigenmodes:
\[
\Theta_n(q) \sim e^{i n \psi} \Psi_n(x),
\]
as discrete conscious quanta (psychons), with transitions corresponding to awareness or memory shifts.


\section{Gauge Symmetries: QED, QCD and the Standard Model Embedding}

We extend the field \(\Theta(q)\) to carry internal gauge indices under the symmetry group \( \mathcal{G} = U(1) \times SU(2) \times SU(3) \), corresponding to the Standard Model gauge groups for electromagnetism, weak, and strong interactions.

Let \(A_\mu = A_\mu^a T^a\) denote the full gauge connection, with generators \(T^a\) acting on \(\Theta(q)\) in the appropriate internal representation.

The full covariant derivative becomes:
\[
\mathcal{D}_\mu \Theta = \partial_\mu \Theta + \Omega_\mu \cdot \Theta + i g_1 B_\mu Y \Theta + i g_2 W_\mu^i \tau^i \Theta + i g_3 G_\mu^a \lambda^a \Theta,
\]
where:
\begin{itemize}
  \item \(B_\mu\) is the \(U(1)_Y\) hypercharge field,
  \item \(W_\mu^i\) are the weak \(SU(2)_L\) fields,
  \item \(G_\mu^a\) are the gluon fields of \(SU(3)_C\),
  \item \(Y\), \(\tau^i\), and \(\lambda^a\) are the corresponding generators.
\end{itemize}

The gauge-invariant kinetic Lagrangian reads:
\[
\mathcal{L}_{\text{gauge}} = -\frac{1}{4} \sum_a F^{a}_{\mu\nu} F^{a\,\mu\nu},
\]
with:
\[
F^a_{\mu\nu} = \partial_\mu A^a_\nu - \partial_\nu A^a_\mu + f^{abc} A^b_\mu A^c_\nu,
\]
where \(f^{abc}\) are the structure constants of the respective Lie algebras.

The interaction term in the unified Lagrangian:
\[
\mathcal{L}_{\Theta, \text{int}} = \Re \left[ \bar{\Theta} i \Gamma^\mu \mathcal{D}_\mu \Theta \right],
\]
ensures that the dynamics of \(\Theta\) is fully coupled to the standard model gauge fields.


\section{Free Energy Principle and Fokker–Planck Flow}

We apply the FEP:
\[
\frac{\partial P}{\partial \psi} = -\nabla_q \cdot (D P) + \frac{1}{2} \nabla_q^2 (D^2 P),
\]
modeling informational entropy flow across \( \mathbb{B}^4 \). Drift terms represent prediction error; diffusion encodes uncertainty and updating.

\section{Cosmological Aspects}

The internal phase coordinate plays a role similar to an inflaton or entropy-gradient field. Toroidal compactification of internal directions leads to inflationary and cyclic cosmological models.

\section{Conclusion}

We have proposed a unified framework on \( \mathbb{B}^4 \) where geometry, gauge theory, quantum physics, and subjective consciousness are aspects of a single field \( \Theta(q) \). This approach explains both standard physics and phenomenological features of cognition, predicting new psychon-like excitations and quantum collapse via internal oscillations.


\section{Quantum Electrodynamics and Quantum Chromodynamics}

The field \( \Theta(q) \) contains internal symmetry structures corresponding to the gauge groups of the Standard Model. In particular, the internal gauge fiber \( \mathbb{G} \) carries representations of:
\[
\mathbb{G} \cong SU(3)_\text{color} \times SU(2)_L \times U(1)_Y,
\]
embedded within the matrix structure of \( \Theta \).

\subsection{QED Sector}

The Abelian part of the gauge group \( U(1) \) governs the electromagnetic interaction. The corresponding gauge field \( A_\mu \) enters the covariant derivative as:
\[
\mathcal{D}_\mu \Theta = \partial_\mu \Theta + \Omega_\mu \cdot \Theta + i e A_\mu \Theta,
\]
with Maxwell-type field strength:
\[
F_{\mu\nu} = \partial_\mu A_\nu - \partial_\nu A_\mu.
\]

\subsection{QCD Sector}

The non-Abelian color interaction is embedded via the gluon fields \( G_\mu^a \), \( a=1,\dots,8 \), and SU(3) generators \( T^a \):
\[
\mathcal{D}_\mu \Theta = \cdots + i g_s G_\mu^a T^a \Theta.
\]
The gluonic field strength is:
\[
G_{\mu\nu}^a = \partial_\mu G_\nu^a - \partial_\nu G_\mu^a + g_s f^{abc} G_\mu^b G_\nu^c.
\]

\subsection{Symmetry Breaking and Higgs-Like Mechanism}

The internal phase coordinates \( y^\mu, z^\mu \) provide a natural mechanism for spontaneous symmetry breaking:
\[
\langle \chi(q) \rangle \neq 0 \quad \Rightarrow \quad m \neq 0,
\]
where \( \chi \) acts as a Higgs-like internal mode within \( \Theta \). Mass hierarchies emerge from oscillatory patterns in the biquaternionic manifold.

\section{Discussion of Particle Spectrum}

Excitations of the field \( \Theta \) include not only standard fermions and bosons but also additional modes due to its internal spinor-gauge structure. We predict:

\begin{itemize}
  \item Standard particle spectrum (quarks, leptons, gauge bosons) as tensor-spinor projections,
  \item Scalar internal oscillations as Higgs-like or axion-like particles,
  \item Psychon modes — coherent phase oscillations corresponding to discrete conscious states,
  \item Potential graviton-like curvature modes in the geometric component.
\end{itemize}

\section{Summary and Outlook}

This framework combines geometry, quantum field theory, and information dynamics in a single elegant model defined on \( \mathbb{B}^4 \). It naturally unifies:

\begin{itemize}
  \item General relativity as geometric dynamics of \( G_{\mu\nu}(q) \),
  \item Quantum field theory via tensor-spinor structure of \( \Theta(q) \),
  \item Gauge interactions through internal fiber \( \mathbb{G} \),
  \item Consciousness as internal oscillations and collapse of \( \chi(q) \).
\end{itemize}

Future work will explore quantization in curved biquaternionic space, phenomenology of psychon transitions, and links to holographic principles and modular topologies.

\section{Limit: General Relativity}

In the classical limit where oscillatory and imaginary components vanish, the biquaternionic metric \( G_{\mu\nu}(q) \) reduces to a standard pseudo-Riemannian real-valued tensor on a 4D Lorentzian manifold.

Let \( \Re(G_{\mu\nu}) \rightarrow g_{\mu\nu}(x) \), where \( x^\mu \in \mathbb{R}^4 \) are spacetime coordinates. The resulting Levi-Civita connection \( \Gamma^\lambda_{\mu\nu} \), Ricci tensor \( R_{\mu\nu} \), and scalar curvature \( R \) are defined in the usual way via the metric compatibility and torsion-free condition.

From the action:
\[
S_\text{geom} = \int d^4 x \, \sqrt{-g} \, \frac{1}{2\kappa} R,
\]
we obtain Einstein’s equations:
\[
R_{\mu\nu} - \frac{1}{2} g_{\mu\nu} R = 8 \pi G \, T_{\mu\nu},
\]
as the low-energy limit of the unified biquaternionic dynamics. Thus, general relativity is fully embedded in the geometric projection of our theory.


\section{Free Energy Principle, Drift, and Metacognitive Field}

The internal dynamics of \( \chi(q) \), representing subjective or conscious phase, follow a stochastic evolution modulated by environmental prediction.

We introduce a Fokker–Planck-like evolution equation:
\[
\frac{\partial \rho(\chi, t)}{\partial t} = -\nabla \cdot \left( \mu \rho \right) + D \nabla^2 \rho,
\]
where \( \rho \) is the probability distribution over conscious phase modes, \( \mu \) is the drift induced by minimization of variational free energy \( F[q] \), and \( D \) is the diffusion tensor.

The **metakas** is then defined as the evolving state of this internal probability structure over \( \mathbb{B}^4 \), including memory, attention, and perception gradients.

The drift term reflects predictive coding:
\[
\mu = -\nabla \log P_\text{sensory} + \nabla \log \rho,
\]
corresponding to active inference. Cognitive dynamics are thus described by stochastic partial differential equations over the internal degrees of freedom of \( \Theta \), with attractor-like behavior for stable conscious trajectories.
\end{document}
